### 2017-10-11 Response_3.tex
Add files via upload

\documentclass[11pt]{article}
\usepackage{color}
\usepackage{amsthm,amsmath,amsfonts}
\usepackage[utf8x]{inputenc}
\title{Revision notes for paper\\``The Conjunctive Disjunctive Graph Node Kernel''}
\author{Dinh Tran Van, Alessandro Sperduti and Fabrizio Costa }
\begin{document}
\maketitle

We thank the reviewer for the reviews and helpful comments on our manuscript ``The Conjunctive Disjunctive Graph Node Kernel''. 
In the following, we provide a description of how each comment (in \emph{italic}) has been addressed in this version of the paper.

\section{Reviewer 1}
\subsection*{1}
\begin{itemize}
\item \textit{lines 10-15: if the first claim (bias toward high degree nodes for the heat kernel) is understandable, the others are not and I don't get how the diffusion kernel (for instance) solves this issue, since it is very similar to the previous one. An evaluation of how the kernels address this property should be included in the experimental section or the authors should cite a paper already studying this property on different kernels;}

\underline{Answer}:

\item \textit{lines 20-25: what does "process in an additive and independent fashion" mean in this context. Similarly as before, please justify your claim about the poor discriminative power of these kernels;
}

\underline{Answer}:

\item \textit{lines 146-150: The first claim is misleading. The problem is not that the degree of some nodes is high it is that the degree distribution is highly skewed (and you must include a reference stating this fact for biological networks, as well as describe the own networks on which you are working for this property). Also, the rest of the argument is very strange: how do you get that the probability to find matches decreases exponentially. If there is a mathematical argument behind it, please, detail it;}

\underline{Answer}:

\end{itemize}

\subsection*{2} \textit{The paper is oriented exclusively toward an given application in biology. However, it is not visible from the title of the article and methodological choices are seldom justified in this context. Gene labelling should be evaluated: for instance, it would have been helpful to have a descriptive evaluation of the results of those labels in the experiment section. For the topological labelling, what is the biological meaning of nodes with similar degree being more similar? (give a reference) I don't understand if the topological labels are considered as a discrete or numerical label? An example of the way it partitions the network would be useful. For the functional labelling, evaluate your clusters in a way. Do they somehow relate to the graph structure or not? Do they correspond to clusters enriched in some functions? How robust is this clustering to very incomplete information as found in GO databases? In a similar fashion, can you provide an evaluation of your
decomposition on the networks you are working with?}

\underline{Answer}:

\subsection*{3} \textit{The paper is extremely badly organized, with some notions (like conjunctive and disjunctive edges) used before they are defined. I suggest the following re-organization of the paper:}

\begin{enumerate}
\item \textit{Introduction (that must include a discussion about the problem at stake, gene prioritization)}

\underline{Answer}:

\item \textit{A new kernel (or whatever title you want more informative than "Material and methods" that is a title better suited for bioinformatics journals) with A. Notations B. Node graph kernels + their problems C. Network decomposition D. Your kernel (with the case of discrete and numerical labels separated)}

\underline{Answer}:

\item \textit{Examples of node labelling in your context (to my opinion, Section 2.2.2 does not need the labels to be known. Just explain the kernel in Section 2 and then, in a separate section afterwards, explain your proposals for labelling with a better illustration of their usefulness in the biological framework)}

\underline{Answer}:

\item \textit{Experiments (since the discussion section is very short, I don't think that it is necessary or even relevant to have it separated from the rest of the paper)}
\end{enumerate}

\underline{Answer}:

\subsection*{4} \textit{The experimental part is very light. I would have expected to see an smart evaluation on a simulated dataset illustrating in which theoretical cases the approach is relevant. I think that an evaluation of the sensitivity of the method to some parameters is missing. For instance, why use a gene co-expression network with a correlation threshold of 0.5? (To my own experience, this is a very low threshold that should give a very dense network and is not the standard choices in this area). Also, why have twice less negative examples than positive ones (this is very far from being realistic since positive genes are usually very rare in this kind of real-life applications)?}

\underline{Answer}:

\section{Reviewer 2}
\begin{itemize}
\item \textit{I have found the paper to be hard to follow. This is mostly due to the large number of "sub-methods" that are involved in the proposed apparach, and to the too simplified description of the "global picture" of what the authors are proposing. It would be better to have an initial section or a table or a figure to summarize and easily introduce what the authors are going to propose/describe in the rest of the paper.}

\underline{Answer}:

\item \textit{Related to the previous comment, the notion of conjunctive and disjuctive edges seems central, however they are shortly mentioned in Section 2.2.2 and then shorlty re-defined (?) in Section 2.4. In the first case, the description of these edges must be expanced and clarified, togheter with some more details about Figure 1 (caption and text). Maybe it can be anticipated in the "Definitions" paragraph (page 3), or in the "initial decription" that I introduced in the previous comment.}

\underline{Answer}:

\item \textit{Overall, there is very large number of parameters involved in the proposed approach. The authors describe their model validation procedure in Page 9/10, however the space of possible combinations of parameters is huge + the choices over other thresholds described in the paper. This seems to be an important drawback of the method, and some comments on this point should be added.}

\underline{Answer}: The large number of parameters is a drawback of our propoased graph node kernel. However, some of them are suggested to have smale range of values to be effective.

\item \textit{Page 4, Functional label and real valued vector information encoding: The K-means algorithm is exploited, using the classical Euclidean distance, and the same distance is used in the vec.info.encoding. The authors should motivate this choice.}

\underline{Answer}: We have added a more detailed explanation in the new version of the paper.

\item \textit{Page 6, Section 2.3. A few words about the hash code computation can be helpful. }

\underline{Answer}: We have added a more detailed explanation in the new version of the paper.
\end{itemize}
\end{document}

### 2017-10-11 Scuba_bmc_revision_3.tex
Add files via upload

%% BioMed_Central_Tex_Template_v1.06
%%                                      %
%  bmc_article.tex            ver: 1.06 %
%                                       %

%%IMPORTANT: do not delete the first line of this template
%%It must be present to enable the BMC Submission system to
%%recognise this template!!

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                     %%
%%  LaTeX template for BioMed Central  %%
%%     journal article submissions     %%
%%                                     %%
%%          <8 June 2012>              %%
%%                                     %%
%%                                     %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                                                 %%
%% For instructions on how to fill out this Tex template           %%
%% document please refer to Readme.html and the instructions for   %%
%% authors page on the biomed central website                      %%
%% http://www.biomedcentral.com/info/authors/                      %%
%%                                                                 %%
%% Please do not use \input{...} to include other tex files.       %%
%% Submit your LaTeX manuscript as one .tex document.              %%
%%                                                                 %%
%% All additional figures and files should be attached             %%
%% separately and not embedded in the \TeX\ document itself.       %%
%%                                                                 %%
%% BioMed Central currently use the MikTex distribution of         %%
%% TeX for Windows) of TeX and LaTeX.  This is available from      %%
%% http://www.miktex.org                                           %%
%%                                                                 %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%% additional documentclass options:
%  [doublespacing]
%  [linenumbers]   - put the line numbers on margins

%%% loading packages, author definitions

\documentclass[twocolumn]{bmcart}% uncomment this for twocolumn layout and comment line below
%\documentclass{bmcart}

%%% Load packages
\usepackage{amsthm,amsmath,amsfonts}
%\RequirePackage{natbib}
%\RequirePackage[authoryear]{natbib}% uncomment this for author-year bibliography
%\RequirePackage{hyperref}
\usepackage[utf8]{inputenc} %unicode support
%\usepackage[applemac]{inputenc} %applemac support if unicode package fails
%\usepackage[latin1]{inputenc} %UNIX support if unicode package fails
\usepackage{color}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                             %%
%%  If you wish to display your graphics for   %%
%%  your own use using includegraphic or       %%
%%  includegraphics, then comment out the      %%
%%  following two lines of code.               %%
%%  NB: These line *must* be included when     %%
%%  submitting to BMC.                         %%
%%  All figure files must be submitted as      %%
%%  separate graphics through the BMC          %%
%%  submission process, not included in the    %%
%%  submitted article.                         %%
%%                                             %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\def\includegraphic{}
\def\includegraphics{}

%%% Put your definitions there:
\startlocaldefs
\endlocaldefs

%%% Begin ...
\begin{document}

%%% Start of article front matter
\begin{frontmatter}

\begin{fmbox}
\dochead{Research}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% Enter the title of your article here     %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\title{Scuba: scalable kernel-based gene prioritization}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% Enter the authors here                   %%
%%                                          %%
%% Specify information, if available,       %%
%% in the form:                             %%
%%   <key>={<id1>,<id2>}                    %%
%%   <key>=                                 %%
%% Comment or delete the keys which are     %%
%% not used. Repeat \author command as much %%
%% as required.                             %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\author[
   addressref={cribi,sdb},                   % id's of addresses, e.g. {aff1,aff2}
   noteref={n1},                        % id's of article notes, if any
   email={guido.zampieri@phd.unipd.it}   % email address
]{\inits{G}\fnm{Guido} \snm{Zampieri}}
\author[
   addressref={math},
   noteref={n1},
   email={dinh@math.unipd.it}
]{\inits{DV}\fnm{Dinh Van} \snm{Tran}}
\author[
   addressref={iit},
   email={michele.donini@iit.}
]{\inits{M}\fnm{Michele} \snm{Donini}}
\author[
   addressref={math},
   email={nnavarin@math.unipd.it}
]{\inits{N}\fnm{Nicol\`o} \snm{Navarin}}
\author[
   addressref={math},
   email={aiolli@math.unipd.it}
]{\inits{F}\fnm{Fabio} \snm{Aiolli}}
\author[
   addressref={math},
   email={sperduti@math.unipd.it}
]{\inits{A}\fnm{Alessandro} \snm{Sperduti}}
\author[
   addressref={cribi,bio},
   corref={cribi},                       % id of corresponding address, if any
   email={giorgio.valle@unipd.it}
]{\inits{G}\fnm{Giorgio} \snm{Valle}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% Enter the authors' addresses here        %%
%%                                          %%
%% Repeat \address commands as much as      %%
%% required.                                %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\address[id=cribi]{%
  \orgname{CRIBI Biotechnology Center, University of Padova},
  \street{viale G. Colombo, 3},
 % \postcode{24105}
  \city{Padova},
  \cny{Italy}
}
\address[id=sdb]{%                           % unique id
  \orgname{Department of Women's and Children's Health, University of Padova}, % university, etc
  \street{via Giustiniani, 3},                     %
  %\postcode{}                                % post or zip code
  \city{Padova},                              % city
  \cny{Italy}                                    % country
}
\address[id=math]{%                           % unique id
  \orgname{Department of Mathematics, University of Padova}, % university, etc
  \street{via Trieste, 63},                     %
  %\postcode{}                                % post or zip code
  \city{Padova},                              % city
  \cny{Italy}                                    % country
}
\address[id=iit]{%                           % unique id
  \orgname{Istituto Italiano di Tecnologia}, % university, etc
  \street{Via Morego, 30},                     %
  %\postcode{}                                % post or zip code
  \city{Genoa},                              % city
  \cny{Italy}                                    % country
}
\address[id=bio]{%                           % unique id
  \orgname{Department of Biology, University of Padova}, % university, etc
  \street{viale G. Colombo, 3},                     %
  %\postcode{}                                % post or zip code
  \city{Padova},                              % city
  \cny{Italy}                                    % country
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% Enter short notes here                   %%
%%                                          %%
%% Short notes will be after addresses      %%
%% on first page.                           %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{artnotes}
%\note{Sample of title note}     % note to the article
\note[id=n1]{Equal contributor} % note, connected to author
\end{artnotes}

%\end{fmbox}% comment this for two column layout

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% The Abstract begins here                 %%
%%                                          %%
%% Please refer to the Instructions for     %%
%% authors on http://www.biomedcentral.com  %%
%% and include the section headings         %%
%% accordingly for your article type.       %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{abstractbox}

\begin{abstract} % abstract
\parttitle{Background} %if any
The uncovering of genes linked to human diseases is a pressing challenge in molecular biology and precision medicine. This task is often hindered by the large number of candidate genes and by the heterogeneity of the available information. Computational methods for the prioritization of candidate genes can help to cope with these problems. In particular, kernel-based methods are a powerful resource for the integration of heterogeneous biological knowledge, however, their practical implementation is often precluded by their limited scalability.

\parttitle{Results} %if any
We propose Scuba, a scalable kernel-based method for gene prioritization. It implements a novel multiple kernel learning approach, based on a semi-supervised perspective and on the optimization of the margin distribution. Scuba is optimized to cope with strongly unbalanced settings where known disease genes are few and large scale predictions are required. Importantly, it is able to efficiently deal both with a large amount of candidate genes and with an arbitrary number of data sources. As a direct consequence of scalability, Scuba integrates also a new efficient strategy to select optimal kernel parameters for each data source. We performed cross-validation experiments and simulated a realistic usage setting, showing that Scuba outperforms a wide range of state-of-the-art methods.

\parttitle{Conclusions} 
Scuba achieves state-of-the-art performance and has enhanced scalability compared to existing kernel-based approaches for genomic data. This method can be useful to prioritize candidate genes, particularly when their number is large or when input data is highly heterogeneous. The code is freely available at https://github.com/gzampieri/Scuba.
\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% The keywords begin here                  %%
%%                                          %%
%% Put each keyword in separate \kwd{}.     %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{keyword}
\kwd{gene prioritization}
\kwd{genetic disease}
\kwd{kernel methods}
\kwd{semi-supervised learning}
\end{keyword}

% MSC classifications codes, if any
%\begin{keyword}[class=AMS]
%\kwd[Primary ]{}
%\kwd{}
%\kwd[; secondary ]{}
%\end{keyword}

\end{abstractbox}
%
\end{fmbox}% uncomment this for twcolumn layout

\end{frontmatter}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% The Main Body begins here                %%
%%                                          %%
%% Please refer to the instructions for     %%
%% authors on:                              %%
%% http://www.biomedcentral.com/info/authors%%
%% and include the section headings         %%
%% accordingly for your article type.       %%
%%                                          %%
%% See the Results and Discussion section   %%
%% for details on how to create sub-sections%%
%%                                          %%
%% use \cite{...} to cite references        %%
%%  \cite{koon} and                         %%
%%  \cite{oreg,khar,zvai,xjon,schn,pond}    %%
%%  \nocite{smith,marg,hunn,advi,koha,mouse}%%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%% start of article main body
% <put your article body there>

%%%%%%%%%%%%%%%%
%% Background %%
%%
\section*{Background}
The identification of the genes underlying human diseases is a major goal in current molecular genetics research. Dramatic progresses have been made since the 1980s, when only a few DNA loci were known to be related to disease phenotypes. Nowadays opportunities for the diagnosis and the design of new therapies are progressively growing, thanks to several technological advances and the application of statistical or mathematical techniques. For instance, positional cloning has allowed to map a vast portion of known Mendelian diseases to their causative genes \cite{strachan,botstein}. However, despite the huge advances, much remains to be discovered. On December 21\textsuperscript{st} 2016, the Online Mendelian Inheritance in Man database (OMIM) registered 4,908 Mendelian phenotypes of known molecular basis and 1,483 Mendelian phenotypes of unknown molecular origin \cite{omim}. Moreover, 1,677 more phenotypes were suspected to be Mendelian. But it is among oligogenic and poligenic (and multifactorial) pathologies that the most remains to be elucidated: for the majority of them, only a few genetic loci are known \cite{strachan,botstein}.

Independently of the type of disease, the search of causative genes usually concerns a large number of suspects. It is therefore necessary to recognise the most promising candidates to submit to additional investigations, as experimental procedures are often expensive and time consuming. Gene prioritization is the task of ordering genes from the most promising to the least. In traditional genotype-phenotype mapping approaches - as well as in genome-wide association studies - the first step is the identification of the genomic region(s) wherein the genes of interest lie. Once the candidate region is identified, the genes there residing are prioritized and finally analysed for the presence of possible causative mutations \cite{strachan}. More recently, in new generation sequencing studies this process is inverted as the first step is the identification of mutations, followed by prioritization and final validation \cite{salgado}. Prioritization criteria are usually based on functional relationships, co-expression and other clues linking genes together. In general, all of them follow the ``guilt-by-association" principle, i.e. disease genes are sought by looking for similarities to genes already associated to the pathology of interest \cite{strachan}.

In the last few years, computational techniques have been developed to aid researchers in this task, applying both statistics and machine learning \cite{moreau}. Thanks to the advent of high-throughput technologies and new generation sequencing, a huge amount of data is in fact available for this kind of investigations. In particular, computational methods are essential for multi-\emph{omics} data integration, that has been recognised as a valuable strategy for understanding genotype-phenotype relationships \cite{ritchie}. In fact, clues are often embedded in different data sources and only their combination leads to the emergence of informative patterns. Furthermore, incompleteness and noise of the single sources can be overcome by inference across multiple levels of knowledge.

Several popular algorithms for pattern analysis are based on \emph{kernels}, which are mathematical transformations that permit to estimate the similarity among items (in our case genes) taking into account complex data relations \cite{cristianini}. Importantly, kernels provide a universal encoding for any kind of knowledge representation, e.g. vectors, trees or graphs. When data integration is required, a multiple kernel learning (MKL) strategy allows a data-driven weighting/selection of meaningful information \cite{gonen}. The goal of MKL is indeed to learn optimal kernel combinations starting from a set of predefined kernels obtained by various data sources. Through MKL the issue of combining different data types is then solved by converting each dataset in a kernel matrix.

Numerous MKL approaches have been proposed for the integration of genomic data \cite{wang,borgwardt} and some of them have been applied to gene prioritization \cite{debie,mkl1class,prodige,zakeri}. De Bie \emph{et al} formulated the problem as a one-class support vector machine (SVM) optimization task \cite{debie}, while Mordelet and Vert tackled it through a biased SVM in a \emph{positive-unlabelled} framework \cite{prodige,chapelle}. Recently, Zakeri \emph{et al} proposed an approach for learning non-linear log-euclidean kernel combinations, showing that it can more effectively detect complementary biological information compared to linear combinations-based approaches \cite{zakeri}. However, as highlighted in a recent work by Wang \emph{et al} \cite{wang}, current methods share two limitations: high computational costs - given by a (at least) quadratic complexity in the number of training examples - and the difficulty to predefine optimal kernel functions to be fed to the MKL machine.

In this work we tackle these issues by proposing a novel scalable gene prioritization method based on a particular MKL approach \cite{easymkl}. By this approach, the optimal kernel is efficiently computed by maximizing the distance between positive and negative examples and optimizing the margin distribution \cite{komd}. This permits to obtain a high scalability relatively to the number of kernels, with a linear time complexity and a practically constant memory requirement. However, this approach assumes comparable label noise in the two example distributions, which does not reflect the case in consideration. Moreover, it does not scale with the number of training examples. Here we introduce a new algorithm, specifically adapted to a \emph{positive-unlabelled} unbalanced framework and we apply it to gene prioritization for the first time. The new learning algorithm has an additional gain in scalability that comes particularly useful when large numbers of genes have to be prioritized. This scalability allows us to transform each data source by multiple kernels and alleviates the issue of defining appropriate base kernels for each source. We called the proposed method Scuba (SCalable UnBAlanced gene prioritization).

From an experimental point of view, here we focus on the integration of multiple gene networks whose edges symbolize functional relationships from heterogeneous sources and we employ two different test settings. In the first setting, we reproduce the procedure presented in a previous work by Chen \emph{et al} \cite{f3pc}, built upon cross-validation experiments \cite{devijver} on collections of known disease genes. This kind of evaluation is useful to compare different methods, but results may suffer from overestimation due to the reliance of many data repositories on medical literature or external data sources like OMIM \cite{omim}. Such dependence introduces a bias that may favour the retrieval of known disease genes. Thus, as a second validation we employ a more realistic setting, following a previous evaluation of gene prioritization tools by B\"ornigen \emph{et al} \cite{bornigen}. Here performance measures focus on the ability of predicting disease genes discovered subsequently to the last update of datasets.

Overall, we compare Scuba with other 14 gene prioritization systems, including other 2 kernel-based methods and 8 web tools. We find that Scuba has competitive accuracy and in particular yields the best results in genome-wide prioritizations, showing its value for large-extent applications.

\section*{Methods}
In this section, we first introduce and formalize some concepts that will be used throughout this paper. Then, we present the proposed approach in detail.

\textbf{Disease gene prioritization}: Let us consider a set of genes $\mathcal{G} = \lbrace g_1, g_2, \ldots, g_N \rbrace$ that represents either the global set of genes in the genome or a subset of it. Given another set $\mathcal{P} = \lbrace g_1, g_2, \ldots, g_m \rbrace, \, \mathcal{P} \subset \mathcal{G}$ containing genes known to be associated to a genetic disease, gene prioritization is the task that aims to rank genes in the set of candidates $\mathcal{U} = \mathcal{G} \setminus \mathcal{P}$ $\textcolor{red}{= \lbrace g_{m+1}, g_{m+2}, \ldots, g_N \rbrace}$ according to their likelihood of being related to that disease. Genes in $\mathcal{P}$ are labelled as \emph{positive} and represent a secure source of information. In contrast, candidate genes in $\mathcal{U}$ are technically \emph{unlabelled}, as we expect that some of them may be associated to the disease but we do not know which ones. Under this notation, this problem can be posed as a \emph{positive-unlabelled} (PU) learning task \cite{prodige,chapelle}.

\textbf{Kernel}: \textit{Kernels} can be informally seen as similarity measures between pairs of data examples. Mathematically, such similarities are defined by inner products between vectors of corresponding examples in a Hilbert space $\mathcal{H}$, without the need of an explicit transformation to that space. A kernel function $k$ on $\mathcal{X} \times \mathcal{X} $ is then formally defined as:
\begin{align*}
	k: \mathcal{X} &\times \mathcal{X} \longrightarrow \mathcal{R}\\
	k(x_{1}, x_{2}) &= \mathcal{<}\phi(x_{1}),\phi(x_{2})\mathcal{>} \, ,
\end{align*}
where $x_1, x_2 \in \mathcal{X}$, $\phi$ is a mapping $\phi:\mathcal{X} \longrightarrow \mathcal{H}$ and $k$ needs to be (1) symmetric, i.e. $k(x_1,x_2)=k(x_2,x_1)$ (2) semi-definite, i.e. the kernel matrix defined by $k_{ij} = k(x_i, x_j)$ has all eigenvalues $\geq 0$. Kernels can be used to define similarities starting from various data types, like graph nodes.

\textbf{Graph node kernel}: A graph $G = (V,E)$ is a structure consisting of a node set $V=\{v_1,\dots,v_N\}$ and an edge set $E=\{(v_i,v_j) | v_i,v_j \in V)\}$. A graph node kernel aims at defining a similarity between any couples of nodes in a graph. A considerable number of graph node kernels have been introduced. The most popular is the diffusion kernel \cite{dk} which is based on the heat diffusion phenomenon. The key idea is to allow a given amount of {\em heat} on each node and let it {\em diffuse} through the edges. The similarity between two nodes $v_{i}, v_{j}$ is then measured as the amount of heat starting from $v_{i}$ and reaching $v_{j}$ over an infinite time interval. In the diffusion kernel the heat flow is proportional to the number of paths connecting two nodes, introducing a bias that penalizes peripheral nodes with respect to central ones. This problem is tackled by a modified version called Markov exponential diffusion kernel (MEDK) \cite{mrf} where a Markov matrix replaces the adjacency matrix.  Another kernel called Markov diffusion kernel (MDK) \cite{mdk}, exploits the notion of {\em diffusion distance}, a measure of similarity between patterns of heat diffusion. The regularized Laplacian kernel (RLK) \cite{rlk} implements instead a normalized version of the random walk with restart model and defines the node similarity as the number of paths connecting two nodes with different lengths.

\subsection*{Scalable Multiple Kernel Learning: EasyMKL}
We approach the problem of disease gene prioritization by employing a graph-based integration in which we use graph node kernels to extract gene information and encode it in the form of kernel matrices. However, a big challenge is how to effectively combine kernels when building predictive systems. This challenge can be solved by MKL. In the following, we first formalize the MKL problem and we then briefly introduce a scalable MKL algorithm named EasyMKL \cite{easymkl}.

Given a set of pre-defined kernels, multiple kernel learning is a task that aims at finding an optimal kernel combination: 
\begin{equation}
\textbf{K} = \psi(\textbf{K}_1, \textbf{K}_2,\ldots, \textbf{K}_R) .
\end{equation}
Recently, many MKL methods have been proposed \cite{gonen,wang}. However, most of them require a long computation time and a high memory consumption, especially when the number of pre-defined kernels is high. To tackle these limitations, a scalable multiple kernel learning named EasyMKL has been previously proposed \cite{easymkl}. This method focuses on learning a linear combination of the input kernels with positive linear coefficients, namely
\begin{equation}
\textbf{K} =  \sum_{r=1}^{R} {\eta_r \textbf{K}_r}, \ \eta_r \geq 0 \, ,
\end{equation}
where $\eta=(\eta_{1}, \dots , \eta_{R})$ is the coefficient vector. \textcolor{red}{In a fully supervised binary task, EasyMKL computes the optimal kernel by maximizing the distance between positive and negative examples. The base learner is a kernel-based approach for the optimization of the margin distribution in binary classification or ranking \cite{komd}.}

\textcolor{red}{In order to present its formulation, let us first define the probability distribution $\gamma \in \mathbb{R}^{N}_{+}$ representing weights assigned to training examples and living in the domain $\Gamma = \{ \gamma \in \mathbb{R}^{N}_{+} | \sum_{i \in \mathcal{P}} \gamma_{i}=1, \sum_{i \in \mathcal{N}} \gamma_{i}=1\}$, where $\mathcal{N}$ is the set of negative examples. From this definition, it follows that any element $\gamma \in \Gamma$ represents a pair of points in the input space: the first one is constrained to the convex hull of positive training examples and the second one to the convex hull of negative training examples. As stated above, EasyMKL maximizes the distance between positive and negative examples, optimizing the margin distribution at the same time. Under this notation, the task can be posed as a min-max problem over variables $\gamma$ and $\eta$ as follows:}
\begin{equation}
\max\limits_{\eta:\| \eta \|_2 \leq 1}\min\limits_{\gamma \in \Gamma}\, (1 - \lambda) \gamma^{\top}\\ \textbf{Y} (\sum_{r}{\eta_{r} \textbf{K}_{r}})\textbf{Y} \gamma + \lambda \, \gamma^{\top} \gamma \, .
\end{equation}
Here $\textbf{Y}$ is a diagonal matrix containing the vector of example labels\textcolor{red}{, +1 for positive and -1 for negative. Optimization of the first term alone leads to an optimal probability distribution $\gamma^{*}$ representing the two nearest points in the convex hulls of positive and negative examples, equally to a hard SVM task using a kernel $\textbf{K}$ \cite{komd}.} The second term represents a quadratic regularization over $\gamma$ whose objective solution is the squared distance between positive and negative centroids in the feature space. The regularization parameter $\lambda \in [0,1]$ permits to tune the objective to optimize, by balancing between the two critical values $\lambda=0$ and $\lambda=1$. \textcolor{red}{When $\lambda=0$ we obtain a pure hard SVM objective, while when $\lambda=1$ we get a centroid-based solution.}

It can be shown that this problem has analytical solution in the $\eta$ variable, so that the previous expression can be reshaped into: 
\begin{equation}\label{easymkl_quad_opt}
\min \limits_{\gamma \in \Gamma} \, (1 - \lambda) \gamma^{\top} \textbf{Y} \textbf{K}^{s} \textbf{Y} \gamma + \lambda \, \gamma^{\top} \gamma \, ,
\end{equation}
where $\textbf{K}^{s}=\sum_{r}^{R}\textbf{K}_r$ is the sum of the pre-defined kernels. This minimization can be efficiently solved and only requires the sum of the kernels. The computation of the kernel summation can be easily implemented incrementally and only two matrices need to be stored in memory at a time. As shown in \cite{easymkl}, EasyMKL can deal with an arbitrary number of kernels using a fixed amount of memory and a linearly increasing computation time.

Once the problem in Eq. \ref{easymkl_quad_opt} is solved, we \textcolor{red}{have an optimal distribution $\gamma^{*}$ and} we are able to obtain the optimal kernel weights $\eta_r^*$ by using the formula:
\begin{equation}\label{eta}
	\eta_r^* = \frac{\gamma^* \textbf{Y} \textbf{K}_r \textbf{Y} \gamma^*}{\sum_{r=1}^R \gamma^* \textbf{Y} \textbf{K}_r \textbf{Y} \gamma^*}.
\end{equation}
The optimal kernel is thus evaluated as $ \textbf{K}^{*} = \sum_{r}^R \eta_r^* \textbf{K}_r$. Finally, by replacing $\textbf{K}^{s}$ with $\textbf{K}^{*}$ in Eq. \ref{easymkl_quad_opt}, we can get the final probability distribution $\gamma^{*}$.

\subsection*{Unbalanced Multiple Kernel Learning: Scuba}
In the previous section we introduced EasyMKL, a scalable, efficient kernel integration approach. However, the gene prioritization task has two additional issues that complicate the work. First, our learning setting is not fully supervised: an assumption is that there are some positive examples hidden among the negatives and we want to retrieve them. Thus, we have the certainty about positive examples but not about negative ones. Second, the number of known disease genes is typically much smaller than the number of candidates, making the problem strongly unbalanced. For these reasons, inspired by a previous work \cite{pyros} we propose a new MKL algorithm based on EasyMKL that not only inherits its scalability, but also efficiently deals with an unbalanced setting.

In order to clearly present our method, we first need to highlight the different contributions given by positive and unlabelled examples. Therefore, we define $\textbf{K}^{+}$, $\textbf{K}^{-}$ and $\textbf{K}^{+-}$ the sub-matrices of $\textbf{K}^s$ pertaining to positive-positive, unlabelled-unlabelled and positive-unlabelled example pairs, respectively. Schematically, we have:
\begin{equation*}
\textbf{K}^{s} = \left( \begin{array}{cc}
\textbf{K}^{+} & \textbf{K}^{+-}\\
\textbf{K}^{-+} & \textbf{K}^{-}\\
\end{array} \right) \, .
\end{equation*}
being $\textbf{K}^{-+}$ the transpose of $\textbf{K}^{+-}$. In other words, $\textbf{K}^{+}$ contains similarities among positive examples \textcolor{red}{$g_i \in \mathcal{P}, i = 1, \dots m$}, $\textbf{K}^{-}$ contains similarities among unlabelled examples \textcolor{red}{$g_j \in \mathcal{U}, j = m+1, \dots N$} and $\textbf{K}^{+-}$ includes similarities between positive-unlabelled example pairs. In the same way, we define $\gamma_{+}$ and $\gamma_{-}$ as the probability vectors associated to positive and unlabelled examples, respectively.\\
Under this change of variables, we reformulate the problem as:
\begin{multline*}
	\min \limits_{\gamma \in \Gamma} \, \gamma_{+}^{\top} \textbf{K}^{+} \gamma_{+} - 2 \,\gamma_{+}^{\top} \textbf{K}^{+-} \gamma_{-} +  \gamma_{-}^{\top} \textbf{K}^{-} \gamma_{-} \\+ \lambda_{+} \gamma_{+}^{\top} \gamma_{+} + \lambda_{-} \gamma_{-}^{\top} \gamma_{-} \, .
\end{multline*}
In this new formulation, the original EasyMKL problem is obtained by setting $\lambda_+ = \lambda_- = \frac{\lambda}{1-\lambda}$. However, due to the unbalanced PU nature of the problem, we are interested in using two different regularizations among positive and unlabelled examples. In our case, we decide to fix \emph{a priori} the regularization parameter  $\lambda_{-} = +\infty$, corresponding to fixing $\lambda=1$ over unlabelled examples only. Then, the solution of part of the objective function is defined by the uniform distribution $\gamma_{-} = (\frac{1}{n},\frac{1}{n},\dots\frac{1}{n}) \equiv u$, where \textcolor{red}{$n=N-m$} is the number of unlabelled examples.\\
We inject this analytic solution of part of the problem in our objective function as
\begin{multline*}
	\min \limits_{\gamma \in \Gamma^{+}} \, \gamma_{+}^{\top} \textbf{K}^{+} \gamma_{+} - 2 \,\gamma_{+}^{\top} \textbf{K}^{+-} u +  u^{\top} \textbf{K}^{-} u \\+ \lambda_{+} \gamma_{+}^{\top} \gamma_{+} + \lambda_{-} u^{\top}u \, ,
\end{multline*}
where $\Gamma^{+} = \lbrace \gamma \in \mathcal{R}_+^m | \textcolor{red}{\sum_{i = 1}^{m} \gamma_i = 1, \gamma_j = 1/n \ \forall \, j = m+1, \dots N} \rbrace$ is the probability distribution domain where the distributions over the unlabelled examples correspond to the uniform distribution. It is trivial that $u^{\top} \textbf{K}^{-} u$ and $\lambda_{-} u^{\top}u$ are independent from the $\gamma_{+}$ variable. Then, they can be removed from the objective function obtaining
\begin{equation}\label{scuba_opt}
	\min \limits_{\gamma \in \Gamma^{+}} \, \gamma_{+}^{\top} \textbf{K}^{+} \gamma_{+} - 2 \, \gamma_{+}^{\top} \textbf{K}^{+-} u + \lambda_{+} \gamma_{+}^{\top} \gamma_{+} \, .
\end{equation}
In this expression, we only need to consider the entries of the kernel $\textbf{K}^{s}$ concerning the positive set, avoiding all the entries with indices in the unlabelled set. The complexity becomes quadratic in the number of positive examples $m$, which is always much smaller than the number of examples to prioritize. Moreover, this algorithm still depends linearly on the number of kernels $R$ and the overall time complexity is then $\mathcal{O}(m^2 \cdot R)$. In this way, we greatly simplify the optimization problem, while being able to take into account the diverse amount of noise present in positive and unlabelled example sets.\\
Like in the previous section, after solving the problem of Eq. \ref{scuba_opt} we use Eq. \ref{eta} to compute the optimal kernel weights. Next, we solve again the Scuba optimization problem to get the final optimal probability distribution $\gamma^{*}$. The likelihood of association to the disease for every gene is given by the vector of scores $s$ defined as
\begin{equation}\label{score_function}
	s = \textbf{K}^{*} \textbf{Y} \gamma^{*} \, ,
\end{equation}
\textcolor{blue}{where $\textbf{K}^{*}$ is the final kernel matrix, computed by $\textbf{K}^{*} = \sum_{r}^R \eta_r^* \textbf{K}_r$. We apply the formula \ref{score_function} to compute scores for the test genes as:
\begin{equation}\label{score_function}
	s = \textbf{K}_{}^{t*} \textbf{Y} \gamma^{*} \, ,
\end{equation}
where $\textbf{K}_{}^{t*}$ is the final test kernel matrix, obtained by taking the weighted sum over all test kernel matrices:
\begin{center}
$\textbf{K}_{}^{t*} = \sum_{r}^{} \eta_{r}^{*} \textbf{K}_{r}^{t}$,
\end{center}
$\textbf{K}_{}^{t*}, \textbf{K}_{r}^{t}$s are matrices of size $(p,q)$ with $p$, $q$ are the number of test genes and training genes, respectively. More particularly, a score for $s_i$ of the test gene $g_i$ is computed as: 
\begin{center}
$s_i = \sum_{j}^{} y_j \gamma_{j}^* \textbf{K}_{ij}^{t*}$.
\end{center}
In words, $s_i$ is the weighted sum over the weighted similarities between the test gene $g_i$ and all genes in the training set $g_j$s. Once we get the scores for test genes, candidate genes are then prioritized based on the score values.
}
\subsection*{Base kernels selection}\label{basekernels}
We leverage the scalability achieved by the new algorithm to ease the optimization of base kernels. As a general practical case, we start from a set of data sources $\mathcal{S} = \lbrace S_1, S_2,\ldots, S_L \rbrace$ representing various levels of biological information. \textcolor{blue}{We first construct a set of corresponding graphs derived from the set of data sources $S$ to obtain a set of graphs $G = \lbrace G_1, G_2, \ldots, G_L\rbrace$. We then apply different kernels with different parameter values on each $G_i \in \mathcal{G}$ by using kernel formulas. As a consequence, for each source $G_i$, we get a set of kernel matrices $\mathcal{K}_i = \lbrace K_{i1}, K_{i2},\ldots, K_{iH} \rbrace$}. By collecting all kernels from all $\mathcal{K}_i$, we achieve a final kernel matrix set $\mathcal{K}$ comprising $L\cdot H$ matrices. Next, all matrices in $\mathcal{K}$ and gene sets $\mathcal{P}$ and $\mathcal{U}$ are fed into Scuba to obtain the optimal kernel $K$. In this way, we directly use MKL to perform an automatic selection of optimal kernel parameters. The final kernel and the disease gene set $\mathcal{P}$ are then employed to train a model, which is used to generate a score list for candidate genes in $\mathcal{U}$ through Eq. \ref{score_function}. The score assigned to a candidate expresses the likelihood of it being associated to the disease.

\subsection*{Experimental workflow}
We employ Scuba to prioritize candidate genes starting from multiple gene networks, obtained by various data sources. We transform every network by means of multiple graph node kernels as explained in the previous section. In the cross-validation experimental setting we use MEDK to estimate the similarity among genes, just like in \cite{f3pc}. In the unbiased setting we use MDK and RLK, selected by validating on training sets.

In both settings, we fix the number of kernel matrices per data source $H=3$ and learn the regularization parameter $\lambda_{+}$ by employing k-fold cross validation on the training set, using the the grid of values $\lbrace 0, \ 0.1,\ 0.2,\ \ldots, 1 \rbrace$. \textcolor{blue}{Kernel parameter values are set as follows: $\lbrace 0.01,\ 0.04,\ 0.07 \rbrace$ for MEDK, suggested in \cite{dir} and used in \cite{mrf}, $\lbrace 2,\ 4,\ 6 \rbrace$ for MDK and $\lbrace 1,\ 10,\ 100 \rbrace$ for RLK, suggested in \cite{mdk}.}

\subsection*{Data sources}
We employ several biological data sources to test Scuba, presented in the following.
\begin{itemize}
	\item \textbf{Human Protein Reference Database} (\textbf{HPRD}) \cite{hprd}. The HPRD resource provides protein interaction data which we implement as an unweighted graph, where genes are linked if their corresponding proteins interact.
	\item \textbf{BioGPS} \cite{biogps}. It contains expression profiles for 79 human tissues, which are measured by using the Affymetrix U133A array. Gene co-expression, defined by pairwise Pearson correlation coefficients (PCC), is used to build an unweighted graph. A pair of genes are linked by an edge if the PCC value is larger than 0.5.
	\item \textbf{Pathways}. Pathway datasets are obtained from the database of KEGG \cite{kegg}, Reactome \cite{reactome}, PharmGKB \cite{pharmgkb} and PID \cite{pid}, which contain 280, 1469, 99 and 2679 pathways, respectively. A pathway co-participation network is constructed by connecting genes that co-participate in any pathway.
	\item \textbf{String} \cite{string}. The String database gathers protein information covering seven levels of evidence: genomic proximity in procaryotes, fused genes, co-occurrence in organisms, co-expression, experimentally validated physical interactions, external databases and text mining. Overall, these aspects focus on functional relationships that can be seen as edges of a weighted graph, where the weight is given by the reliability of that relationship. To perform the unbiased evaluation we employed the version 8.2 of String, from which we extracted functional links among 17078 human genes.
\end{itemize}

The first three datasets were obtained directly from Chen \emph{et al} \cite{f3pc}, already preprocessed in such a way that all of them represent exactly the same 7311 genes. We employed this data without any further processing.

Known gene-disease associations employed in the cross-validation experimental setting were taken from a work of Goh \emph{et al} which defines classes of related diseases \cite{goh}. Training and candidate gene sets used in the second set of experiments (section \emph{Unbiased evaluation}) were obtained from the supplementary material of the unbiased evaluation of gene prioritization tools performed by B\"ornigen \emph{et al} \cite{bornigen}. Finally, gene-disease associations from the Human Phenotype Ontology were used, belonging to builds 29 and 117 \cite{hpo}.

\subsection*{Other kernel-based gene prioritization methods}
We compare Scuba with other two kernel methods for gene prioritization. The first one implements a one class approach to MKL, slightly modifying the formulation of the method of De Bie \emph{et al} \cite{debie}. In the corresponding work \cite{mkl1class}, authors show that this newer approach reaches higher performances in ranking. In the following, we refer to it as MKL1class. The second method we consider is ProDiGe, a PU approach that combines MKL and multitask learning \cite{prodige}. We focus on its first version without multitask learning, as our purpose is to study performances in terms of the MKL framework. We ran ProDiGe using the default parameters indicated in the corresponding paper: number of bagging iterations $B=30$ and regularization parameter $C=1$. In the same way, we set the regularization parameter $\nu=0.5$ for MKL1class.

\section*{Results}
In this section, we describe the tests made to evaluate our proposed method, which follow two different experimental procedures. In the first setting, we aim at estimating Scuba performance in a standard validation framework. In the second setting we evaluate it by an unbiased approach, making a comparison with prioritization tools available on the web and with two state-of-the-art kernel-based methods.

\subsection*{Cross-validation}
As a first evaluation of Scuba, we followed the experimental protocol used by Chen \emph{et al} to test predictive performance of other prioritization methods \cite{f3pc}. In this setting, we employed three data sets: BioGPS, HPRD and Pathways, which we borrowed from the authors of the work. To perform the experiments, we employed known gene-disease associations from OMIM, grouped into 20 classes on the basis of disease relatedness by Goh \emph{et al} \cite{goh}. Among those classes we selected the 12 with at least 30 confirmed genes. We then built a training set consisting of a positive set \textcolor{red}{$\mathcal{P}$ and an unlabelled set $\mathcal{U}$} for each of them. \textcolor{red}{$\mathcal{P}$} contains all its disease gene members. \textcolor{red}{$\mathcal{U}$} is constructed by randomly picking genes from known disease genes such that $\vert \textcolor{red}{\mathcal{U}} \vert = \frac{1}{2} \vert \textcolor{red}{\mathcal{P}} \vert$. The unknown genes relate to at least one disease class, but do not relate to the current class. We chose the genes in \textcolor{red}{$\mathcal{U}$} from the other disease genes because we assume that they were less likely to be associated to the considered class. In fact, disease genes are generally more studied and a potential association has more chances to have already been identified.

After that, leave-one-out cross validation was used to evaluate the performance of the algorithm. Iteratively, every gene in the training set was selected to be the test gene and the remaining genes in \textcolor{red}{$\mathcal{P}$} and \textcolor{red}{$\mathcal{U}$} were used to train the model. Once the model was trained, a score list for the test gene and the candidate genes \textcolor{red}{in $\mathcal{U}$} was computed. Then, we computed a decision score for each test gene representing the percentage of candidate genes ranked lower than it. We collected all decision scores for every gene in all disease classes to form a global decision score list. The performance of Scuba was measured by calculating the area under the curve (AUC) in the receiver-operating-characteristic plot obtained from the decision score list. The AUC expresses the probability that a randomly chosen disease gene is ranked above a randomly picked non-disease gene for any disease class.

Table \ref{table: MRF-comparison} illustrates the performance of different techniques in this experimental setting reported by Chen \emph{et al} \cite{f3pc}, and the performance of our proposed method. In the second column we show the significance of the difference between reported AUCs and Scuba AUC, assessed by means of separate pairwise comparisons \cite{hanley}. Scuba performs significantly better than the other methods, getting an AUC around 3.6\% greater than the second best performing technique, F3PC.

\subsection*{Unbiased evaluation}\label{unbiased}
Although the previous evaluation is useful to compare Scuba with other methods, predictive performance in cross-validation experiments may be inflated compared to real applications. Indeed, the retrieval of known disease genes can be facilitated by various means. One mean is the crosstalk between data repositories: for example, KEGG \cite{kegg} draws its information also from medical literature. Moreover, often the discovery of the link between a gene and a disease coincides with the discovery of a functional annotation or of a molecular interaction. In practice, instead, researchers are interested in novel associations, which in most cases are harder to find due to a lack of information around them.

In order to achieve a thorough evaluation of Scuba, we tested it in a more realistic setting, following the work of B\"{o}rnigen \emph{et al} \cite{bornigen}. In this study, eight gene prioritization web tools were benchmarked as follows. Newly discovered gene-disease associations were collected over a timespan of six months, gathering \textcolor{red}{42 test genes associated to a range of disorders}. As soon as a new association was discovered, \textcolor{red}{each web tool was queried with a disorder-specific set of positive genes $\mathcal{P}$ to prioritize a set of candidates $\mathcal{U}$ containing the corresponding test gene (or to prioritize the whole genome where possible). Rank positions of the 42 test genes were ultimately} used to assess the ability of the tools to successfully prioritize disease genes. The idea behind this procedure is to anticipate the integration of the associations in the data sources and so avoid biased predictions.

In order to test Scuba in this setting, we backdated our data to a time prior to May 15, 2010 by employing String v8.2 data \cite{string}. \textcolor{red}{After that, we recovered positive sets and test genes from the original publication and we followed its experimental protocol as follows \cite{bornigen}. We performed prioritizations for each test gene in two distinct cases: genome-wide and candidate set-based prioritizations. In a genome-wide prioritization all genes in the String dataset - except those in $\mathcal{P}$ - belong to $\mathcal{U}$ and were prioritized. In candidate set-based tests, the sets of candidates $\mathcal{U}$ were constructed by considering all genes with Ensembl \cite{ensembl} gene identifier within the chromosomal regions around the test genes, in order to get on average 100 candidates for each trial.} In both cases, we normalized the ranking positions over the total number of considered genes in order to get median, mean and standard deviation of the normalized ranks for test genes. We also computed the true positive rate (TPR) relatively to some representative thresholds (5\%, 10\% and 30\% of the ranking) and the AUC obtained by averaging over the 42 prioritizations.

Along with Scuba, we evaluated in this setting also MKL1class \cite{mkl1class} and ProDiGe \cite{prodige}, two state-of-the-art kernel based gene prioritization methods.  In Table \ref{unbiased1} it is possible to see performances for all three methods. The significance of rank median differences between Scuba and competing methods was assessed by Wilcoxon signed rank tests, one for each comparison. At a significance threshold of 0.05, Scuba achieves significantly higher performances in genome-wide tasks compared to both baselines. In the candidate set-based setting, it performs significantly better than ProDiGe and better, although not significantly, than MKL1class. These differences can be visually appreciated in Figure \ref{scatter}, where we compare the rank distributions \textcolor{red}{of test genes} obtained by the three methods. Scuba and MKL1class present moderate rank differences, particularly in the central region of the ranks. On the other hand, differences between Scuba and ProDiGe are smaller (Pearson $r=0.98$ in both cases) and almost all in favour of Scuba.

In Table \ref{unbiased2} we show results for Scuba compared to the results obtained in the work of B\"{o}rnigen \emph{et al}, pertaining to eight prioritization systems \cite{bornigen}. In genome-wide predictions, Scuba dominates over the other tools. On predictions over smaller candidate sets, it is still competitive although best results are achieved by GeneDistiller \cite{genedistiller}, Endeavour \cite{endeavour} and ToppGene \cite{toppgene}. It is important to underline that in this case considered tools rely on different data sources, so we are comparing different prioritization systems rather different algorithms. Furthermore, tools are in some cases unable to provide an answer to a given task, depending on the underlying data sources (for more details see the original work \cite{bornigen}). We report the fraction of prioritizations on which tools are actually evaluated as response rate. This table has the purpose of showing the potentiality of Scuba relatively to what is easily accessible by non-bioinformaticians. However, since we used the String data for instance Scuba is directly comparable with Pinta \cite{bornigen, pinta}.

Next, we expanded this validation by employing gene-disease annotations derived from the Human Phenotype Ontology (HPO) \cite{hpo}. This resource gathers information from several databases and makes available its monthly updates, permitting to trace the annotations history. We downloaded the HPO build 29 - dating March 2013 - and build 117 of February 2017. We compared the two annotations corresponding to these versions of HPO and extracted the gene-disease associations that were added in this time gap. We concentrated on the multifactorial diseases covered in the previous analysis, that could possibly have some previously undiscovered associations. We thus analyzed how the obtained genes are ranked in genome-wide prioritizations of the previous analysis, applying the same performance measures as before. The outcome is an analogous evaluation, but this time target genes are those extracted from HPO.

In Table \ref{hpo} results for Scuba, MKL1class and ProDiGe are shown. We can observe a slightly different trend compared to previous results, with Scuba and ProDiGe having very close performance and MKL1class being significantly worse than Scuba. As a confirmation, in Figure \ref{scatter} we can see that there is no clear difference between \textcolor{red}{test gene rank distributions for Scuba and ProDiGe}. Instead, MKL1class ranks several test genes neatly lower compared to Scuba, with the associated Pearson correlation coefficient dropping to $r=0.85$.

\section*{Discussion}
Gene prioritization is progressively becoming essential in molecular biology studies. In fact, we are assisting to a continuous proliferation of a variety of \emph{omic} data brought by technological advances. In the near future it is then likely that more heterogeneous knowledge will have to be combined. Moreover, the classes of biological agents to be prioritized are going to enlarge. For instance, we are only beginning to understand the complex regulation machinery involving non-coding RNA and epigenetic agents. It is estimated that around 90.000 human long non coding genes exist, whose functional implications are progressively emerging \cite{noncode}. Facing this challenge, the development of novel methods is still strongly needed in order to enhance predictive power and efficiency.

Compared to the considered benchmark kernel methods - MKL1class and ProDiGe - Scuba has some important advantages. ProDiGe is one of the first proposed kernel-based PU learning method for gene prioritization \cite{prodige}. It implements a PU learning strategy based on a biased SVM, which over-weights positive examples during training. In order to reach scalability to large datasets, it leverages a bagging procedure. Like ProDiGe, Scuba implements a learning strategy based on a binary classification set up, but from a different perspective. In a PU problem, the information on positive labels is assumed secure, while the information on negative labels is not. In terms of margin optimization, this translates in unbalanced entropy on the probability distributions associated to the two sets of training examples. It is then required to regularize more on the unlabelled class - having higher entropy - and in the limit of maximum uncertainty we get the uniform distribution.

MKL1class implements another effective approach for data integration, namely single class learning. This means that the model is obtained solely based on the distribution of known disease genes, disregarding unlabelled ones. Scuba has enhanced scalability compared to MKL1class, as it involves the optimization of the 1-norm of the margin vector from the different kernels. In contrast, MKL1class optimizes its 2-norm, which is more computationally demanding. Importantly, another distinctive feature of Scuba is a time complexity dependent on the number of positive examples and not on the number of total examples. As a consequence, Scuba can exploit the information on the whole data distribution and at the same time scale to large datasets without the need of sub-sampling the examples. This may be of great advantage as typically disease genes are orders of magnitude less numerous than the candidates.

Results from two different evaluation settings show that our proposed method Scuba outperforms many existing methods, particularly in genome-wide analyses. Compared to the two considered existing kernel-based methods, Scuba performances (considering AUC) are always higher, and often significantly higher. Moreover, Scuba has two main levels of scalability that make it particularly suitable for gene prioritization:
\begin{itemize}
	\item \textbf{Scalability on number of kernels}: Scuba is able to deal with a large number of kernels defined on different data sources. As a consequence, it can be useful to get a more unified view of the problem and to build more powerful predicting models.

	\item \textbf{Scalability on number of training examples}: In typical gene prioritization problems, the number of known disease genes is much smaller than the number of candidates. Scuba is designed to efficiently deal with unbalanced settings and at the same time take advantage of the whole candidates distribution.
\end{itemize}

Altogether, our results show that Scuba is a valuable tool to achieve efficient prioritizations, especially in large-scale investigations. A detailed overview on the validation results for single diseases is available in Supplementary Tables 1, 3, 4.

Finally, as it is visible in Supplementary Table 2, performance with multiple kernels might be close to those with single kernels. Nevertheless, feeding multiple kernels into Scuba alleviates the issue of choosing appropriate kernels for each data source, as implemented in our work. Importantly, this strategy can also provide multiple views on the same data and possibly increase performance. Nevertheless caution must be paid since the more kernels are combined and the more parameters have to be learned, thus increasing the risk of over-fitting. We advice then to moderate the number of kernel matrices generated from each data source.

\section*{Conclusion}
In this work, we propose a novel computational kernel-based method to guide the identification of novel disease genes. Our method takes advantage of complementary biological knowledge by combining heterogeneous data sources. Every source can be transformed by appropriate kernel functions in order to take full advantage of its information. Our original algorithm is scalable relatively to the size of input data, number of kernel transformations employed and number of training examples. Experimental results support the thesis that Scuba is an effective approach and can be applied in various disease domains.

Scuba only requires a collection of input genes and optionally a set of candidate genes. The simple requirements make it applicable to a wide range of laboratory investigations. Furthermore, Scuba can be potentially employed also in other prioritization problems, as long as a PU approach and the integration of heterogeneous biological knowledge are needed.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% Backmatter begins here                   %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{backmatter}

\section*{List of abbreviations}
\textbf{OMIM:} Online Mendelian Inheritance in Man database\\
\textbf{MKL:} Multiple Kernel Learning\\
\textbf{SVM:} Support Vector Machine\\
\textbf{Scuba:} SCalable UnBAlanced gene prioritization\\
\textbf{PU learning:} Positive-Unlabelled learning\\
\textbf{MEDK:} Markov Exponential Diffusion Kernel\\
\textbf{MDK:} Markov Diffusion Kernel\\
\textbf{RLK:} Regularized Laplacian Kernel\\
\textbf{EasyMKL:} Easy Multiple Kernel Learning\\
\textbf{HPRD:} Human Protein Reference Database\\
\textbf{PCC:} Pearson Correlation Coefficient\\
\textbf{AUC:} Area Under the receiver-operating-characteristic Curve\\
\textbf{TPR:} True Positive Rate\\
\textbf{DIR:} Data Integration Rank\\
\textbf{MRF:} Markov Random Field\\
\textbf{F3PC:} a logistic regression-based algorithm\\
\textbf{HPO:} Human Phenotype Ontology

\section*{Declarations}

\section*{Ethics approval and consent to participate}
  Not applicable.

\section*{Consent for publication}
  Not applicable.

\section*{Availability of data and material}
  The python code of the proposed algorithm is available in the GitHub repository at https://github.com/gzampieri/Scuba. The networks used in cross-validation experiments were borrowed by the supporting data of \cite{f3pc}. The String dataset is available at http://string-db.org/. Disease gene information was taken from the supplementary material of \cite{bornigen} and \cite{goh} and from the Human Phenotype Ontology archive at https://github.com/Human-Phenotype-Ontology/HPO-archive.

\section*{Competing interests}
  The authors declare that they have no competing interests.

\section*{Funding}
  This work was supported by the University of Padova, Strategic Project BIOINFOGEN.

\section*{Author's contributions}
  All authors contributed to the design of the study. GZ and DTV equally contributed to the implementation of the method, to performing the experiments and to writing the manuscript. GV and AS supervised the work. All authors read and approved the final manuscript.

\section*{Acknowledgements}
  Not applicable.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                  The Bibliography                       %%
%%                                                         %%
%%  Bmc_mathpys.bst  will be used to                       %%
%%  create a .BBL file for submission.                     %%
%%  After submission of the .TEX file,                     %%
%%  you will be prompted to submit your .BBL file.         %%
%%                                                         %%
%%                                                         %%
%%  Note that the displayed Bibliography will not          %%
%%  necessarily be rendered by Latex exactly as specified  %%
%%  in the online Instructions for Authors.                %%
%%                                                         %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% if your bibliography is in bibtex format, use those commands:
\bibliographystyle{bmc-mathphys} % Style BST file (bmc-mathphys, vancouver, spbasic).
\bibliography{Scuba_bmc_revision_2}      % Bibliography file (usually '*.bib' )
% for author-year bibliography (bmc-mathphys or spbasic)
% a) write to bib file (bmc-mathphys only)
% @settings{label, options="nameyear"}
% b) uncomment next line
%\nocite{label}

% or include bibliography directly:
% \begin{thebibliography}
% \bibitem{b1}
% \end{thebibliography}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                               %%
%% Figures                       %%
%%                               %%
%% NB: this is for captions and  %%
%% Titles. All graphics must be  %%
%% submitted separately and NOT  %%
%% included in the Tex document  %%
%%                               %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%
%% Do not use \listoffigures as most will included as separate files

\section*{Figures}

\begin{figure}[h!]
\caption{\csentence{Comparison of normalized ranks predicted by Scuba and competing kernel methods.}
	Normalized \textcolor{red}{test gene rank distributions} predicted by Scuba, MKL1class and ProDiGe for test genes in \textbf{(a)} genome-wide prioritizations in the unbiased evaluation of Table \ref{unbiased1} - \textbf{(b)} candidate set-based prioritizations in the unbiased evaluation of Table \ref{unbiased1} - \textbf{(c)} genome-wide prioritizations in the expanded unbiased evaluation of Table \ref{hpo}. In all cases, each point represents a test gene and lower values on the axes indicate better predictions. Genes lying on a diagonal have the same rank according to both methods considered on a plot. The further a gene lies above (below) a diagonal and the better it was ranked by Scuba (MKL1class/ProDiGe) compared to MKL1class/ProDiGe (Scuba). In each plot we show the Pearson correlation coefficient $r$ between the \textcolor{red}{test gene rank distributions} and its associated p-value.\label{scatter}}
\end{figure}

%\begin{figure}[h!]
%  \caption{\csentence{Sample figure title.}
%      Figure legend text.}
%      \end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                               %%
%% Tables                        %%
%%                               %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% Use of \listoftables is discouraged.
%%
\section*{Tables}
\begin{table*}[h!]
	\caption{The performance of different techniques in the experimental setting of Chen \emph{et al} \cite{f3pc} expressed in terms of AUC. Except for our proposed method Scuba, these results were taken from that work. The p-values indicate significance of the pairwise AUC differences with respect to Scuba AUC \cite{hanley}. Asterisks indicate significance of the test (p-value $<$ 0.05).\label{table: MRF-comparison}}
	\begin{tabular}{c c c}
		\hline\noalign{\vskip 1mm}
		Method & AUC & p-value \\
		\noalign{\vskip 1mm}\hline\noalign{\vskip 1mm}
		Scuba & \textbf{0.876} & - \\
		F3PC \cite{f3pc} & 0.830 & 1.39$\,\cdot$10$^{-4}$ * \\
		MRF \cite{mrf} & 0.731 & $<$10$^{-6}$ * \\
		DIR \cite{dir} & 0.716 & $<$10$^{-6}$ * \\
		GeneWanderer \cite{genewanderer} & 0.711 & $<$10$^{-6}$ * \\
		\noalign{\vskip 1mm}\hline
	\end{tabular}
\end{table*}

\begin{table*}[h!]
	\caption{Performances of Scuba, MKL1class and ProDiGe in the unbiased setting of B\"{o}rnigen \emph{et al} \cite{bornigen}. Values refer to predictions on all the 42 gene-disease associations. Rank difference p-values were obtained using Wilcoxon signed rank tests comparing separately Scuba/MKL1class and Scuba/ProDiGe ranks differences. Asterisks indicate significance of the tests at a threshold of 0.05.\label{unbiased1}}
	\begin{tabular}{c c c c c c c c}	
		\hline\noalign{\vskip 1mm}
		Tool/Method & Rank & Rank & TPR in top & TPR in top & TPR in top & AUC & Rank difference \\
		& median & average & 5\% (\%) & 10\% (\%) & 30\% (\%) & & p-value \\
		\noalign{\vskip 1mm}\hline\noalign{\vskip 1mm}
		\multicolumn{5}{l}{Genome-wide prioritization methods}\\[1mm]
		Scuba & \textbf{10.55} & \textbf{20.48}$\pm$23.53 & \textbf{33.3} & \textbf{47.6} & \textbf{78.6} & \textbf{0.80} & - \\
		MKL1class \cite{mkl1class} & 13.30 & 23.42$\pm$23.23 & 21.4 & \textbf{47.6} & 69.0 & 0.77 & 2.5$\,\cdot$10$^{-2}$ * \\
		ProDiGe \cite{prodige} & 11.73 & 24.45$\pm$27.33 & 31.0 & 45.2 & 71.4 & 0.76 & 3.0$\,\cdot$10$^{-7}$ * \\
		\noalign{\vskip 1mm}\hline\noalign{\vskip 1mm}
		\multicolumn{5}{l}{Candidate set-based prioritization methods}\\[1mm]
		Scuba & \textbf{12.95} & \textbf{23.32}$\pm$25.46 & \textbf{28.6} & \textbf{45.2} & \textbf{73.8} & \textbf{0.78} & -  \\
		MKL1class \cite{mkl1class} & 15.07 & 25.63$\pm$24.73 & 23.8 & 40.5 & 61.9 & 0.76 & 9.7$\,\cdot$10$^{-2}$ \\
		ProDiGe \cite{prodige} & 14.41 & 26.39$\pm$29.09 & 26.2 & 40.5 & 71.4 & 0.75 & 2.7$\,\cdot$10$^{-3}$ * \\
		\noalign{\vskip 1mm}\hline
	\end{tabular}
\end{table*}

\begin{table*}[h!]
	\caption{Performances of Scuba and of some gene prioritization web tools in the unbiased setting of B\"{o}rnigen \emph{et al} \cite{bornigen}. Response rate is the percentage of gene-disease associations considered by each tool. Values for Suspects were computed on the first 27 associations only (highlighted by $^{a}$).\label{unbiased2}}
	\begin{tabular}{c c c c c c c c}	
		\hline\noalign{\vskip 1mm}
		Tool/Method & Response & Rank & Rank & TPR in top & TPR in top & TPR in top & AUC\\
		& rate (\%) & median & average & 5\% (\%) & 10\% (\%) & 30\% (\%) & \\
		\noalign{\vskip 1mm}\hline\noalign{\vskip 1mm}
		\multicolumn{5}{l}{Genome-wide prioritization methods}\\[1mm]
		 Scuba & 100 & \textbf{10.55} & \textbf{20.48}$\pm$23.53 & \textbf{33.3} & \textbf{47.6} & \textbf{78.6} & \textbf{0.80}\\
		Candid \cite{candid} & 100 & 18.10 & 27.35$\pm$24.62 & 21.4 & 33.3 & 64.3 & 0.73\\
		Endeavour \cite{endeavour} & 100 & 15.49 & 21.47$\pm$22.37 & 28.6 & 38.1 & 71.4 & 0.79\\
		Pinta \cite{pinta} & 100 & 19.03 & 23.52$\pm$23.58 & 26.2 & 31.0 & 71.4 & 0.77\\
		\noalign{\vskip 1mm}\hline\noalign{\vskip 1mm}
		\multicolumn{5}{l}{Candidate set-based prioritization methods}\\[1mm]
		Scuba & 100 & 12.95 & 23.32$\pm$25.46 & 28.6 & 45.2 & 73.8 & 0.78\\
		Suspects \cite{suspects} & 88.9$^{a}$ & 12.77$^{a}$ & 24.64$\pm$26.42$^{a}$ & 33.3$^{a}$ & 33.3$^{a}$ & 63.0$^{a}$ & 0.76$^{a}$\\
		ToppGene \cite{toppgene} & 97.6 & 16.80 & 34.53$\pm$35.31 & \textbf{35.7} & 42.9 & 52.4 & 0.66\\
		GeneWanderer-RW	\cite{genewanderer} & 88.1 & 22.10 & 29.55$\pm$26.28 & 16.7 & 26.2 & 61.9 & 0.71\\
		Posmed-KS \cite{posmed} & 47.6 & 31.44 & 42.07$\pm$30.98 & 4.7 & 7.1 & 23.8 & 0.58\\
		GeneDistiller \cite{genedistiller} & 97.6 & \textbf{11.11} & \textbf{15.37}$\pm$13.77 & 26.2 & \textbf{47.6} & 78.6 & \textbf{0.85}\\
		Endeavour \cite{endeavour} & 100 & 11.16 & 18.41$\pm$21.39 & 26.2 & 42.9 & \textbf{90.5} & 0.82\\
		Pinta \cite{pinta} & 100 & 18.87 & 25.23$\pm$24.72 & 28.6 & 31.0 & 71.4 & 0.75\\
		\noalign{\vskip 1mm}\hline
	\end{tabular}
\end{table*}

\begin{table*}[h!]
	\caption{Performances of Scuba, MKL1class and ProDiGe in the expanded unbiased setting involving seven multifactorial diseases. Values refer to predictions on 48 gene-disease associations. Rank difference p-values were obtained using Wilcoxon signed rank tests comparing separately Scuba/MKL1class and Scuba/ProDiGe ranks differences. Asterisks indicate significance of the tests at a threshold of 0.05.\label{hpo}}
	\begin{tabular}{c c c c c c c c c}	
		\hline\noalign{\vskip 1mm}
		Method & Rank & Rank & TPR in top & TPR in top & TPR in top & TPR in top & AUC & Rank difference \\
		& median & average & 1\% (\%) & 5\% (\%) & 10\% (\%) & 30\% (\%) & & p-value \\
		\noalign{\vskip 1mm}\hline\noalign{\vskip 1mm}
		\multicolumn{4}{l}{Genome-wide prioritizations}\\[1mm]
		Scuba & 8.13 & \textbf{17.45}$\pm$22.33 & \textbf{10.4} & 41.7 & \textbf{58.3} & \textbf{79.2} & \textbf{0.83} & - \\
		MKL1class \cite{mkl1class} & 14.28 & 25.79$\pm$26.96 & 2.1 & 27.1 & 45.8 & 66.7& 0.74 & 1.2$\,\cdot$10$^{-5}$ * \\
		ProDiGe \cite{prodige} & \textbf{7.89} & 18.40$\pm$23.77 & \textbf{10.4} & \textbf{43.8} & 54.2 & \textbf{79.2} & 0.82 & 9.5$\,\cdot$10$^{-2}$ \\
		\noalign{\vskip 1mm}\hline
	\end{tabular}
\end{table*}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                               %%
%% Additional Files              %%
%%                               %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section*{Additional Files}
  \subsection*{Additional file 1 --- Supplementary Tables}
    PDF file containing experimental results for individual diseases and different kernel combinations.

%  \subsection*{Additional file 2 --- Sample additional file title}
%    Additional file descriptions text.

\end{backmatter}
\end{document}

### 2017-10-11 neurocomputing.tex
Add files via upload

\documentclass[review]{elsarticle}

\usepackage{lineno,hyperref}
\newtheorem{definition}{Definition}
\usepackage{amssymb,amsmath,array}
\usepackage{xcolor}% provides \colorlet
\usepackage{fixme}
\fxsetup{
    status=draft,
    author=,
    layout=inline,
    theme=color
}

\definecolor{fxnote}{rgb}{0.8000,0.0000,0.0000}
% define the background colour:
\colorlet{fxnotebg}{yellow}

% refedine the layout macro:
\makeatletter
\renewcommand*\FXLayoutInline[3]{%
  \@fxdocolon {#3}{%
    \@fxuseface {inline}%
    \colorbox{fx#1bg}{\color {fx#1}\ignorespaces #3\@fxcolon #2}}}
\makeatother

\modulolinenumbers[5]

\journal{Journal of \LaTeX\ Templates}

\newcommand{\IR}{\rm I\!R}

%%%%%%%%%%%%%%%%%%%%%%%
%% Elsevier bibliography styles
%%%%%%%%%%%%%%%%%%%%%%%
%% To change the style, put a % in front of the second line of the current style and
%% remove the % from the second line of the style you would like to use.
%%%%%%%%%%%%%%%%%%%%%%%

%% Numbered
%\bibliographystyle{model1-num-names}

%% Numbered without titles
%\bibliographystyle{model1a-num-names}

%% Harvard
%\bibliographystyle{model2-names.bst}\biboptions{authoryear}

%% Vancouver numbered
%\usepackage{numcompress}\bibliographystyle{model3-num-names}

%% Vancouver name/year
%\usepackage{numcompress}\bibliographystyle{model4-names}\biboptions{authoryear}

%% APA style
%\bibliographystyle{model5-names}\biboptions{authoryear}

%% AMA style
%\usepackage{numcompress}\bibliographystyle{model6-num-names}

%% `Elsevier LaTeX' style
\bibliographystyle{elsarticle-num}
%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

\begin{frontmatter}

\title{The Conjunctive Disjunctive Graph Node Kernel for Disease Gene Prioritization}
%\tnotetext[mytitlenote]{Fully documented templates are available in the elsarticle package on \href{http://www.ctan.org/tex-archive/macros/latex/contrib/elsarticle}{CTAN}.}

%% Group authors per affiliation:
\author{Dinh Tran Van}
\author{Alessandro Sperduti}
\address{Department of Mathematics, Padova University, Trieste, 63, 35121 Padova, Italy}
\author{Fabrizio Costa}
\address{Department of Computer Science, University of Exeter Exeter EX4 4QF, UK}
%\fntext[myfootnote]{Since 1880.}

%% or include affiliations in footnotes:
%\author[mymainaddress,mysecondaryaddress]{Elsevier Inc}
%\ead[url]{www.elsevier.com}

%\author[mysecondaryaddress]{Global Customer Service\corref{mycorrespondingauthor}}
%\cortext[mycorrespondingauthor]{Corresponding author}
%\ead{support@elsevier.com}

%\address[mymainaddress]{1600 John F Kennedy Boulevard, Philadelphia}
%\address[mysecondaryaddress]{360 Park Avenue South, New York}

\begin{abstract}
Biological relations encoded data are naturally represented in the form of graphs. Therefore, there is a high number of graph-based biological learning systems have been proposed and shown promissing results. One of the key points which determines the performance of graph-based learning systems is the denifition of node similarity measure. Node similarities are normally measured by graph node kernels. However, most available graph node kernels do not show high discriminative capacity since they share two common limitations. First they are based on the diffusion phenomenon which does not effectively exploit the nodes' configuration. Second they are not able to take into account the available information associated to nodes of graph.

In this paper, we propose an efficient graph node kernel that not only effectively exploits nodes' context, but also can integrate information which are available on the graph nodes. Empirical evaluation show that our proposed graph node kernel show state of the art concerning graph node kernels.

%Gene-disease associations are inferred on the basis of similarities between the proteins encoded by genes. Biological relationships used to define similarities range from interacting proteins, proteins that participate in pathways and protein expression profiles. Though graph kernel methods have become a prominent approach for association prediction, most solutions are based on a notion of information diffusion that does not capture the specificity of different network parts. Here we propose a graph kernel method that explicitly models the configuration of each gene's context. An empirical evaluation on several biological databases shows that our proposal achieves state-of-the-art results.
\end{abstract}

\begin{keyword}
Graph node kernels, graph decomposition, disease gene prioritization
\end{keyword}
\end{frontmatter}

\linenumbers
\section{Introduction}
Node similarity measure definition is the key that determines the performance of graph-based biological learning systems. The state of the art graph node kernels used to measure node similarity, are based on the notion of information diffusion, including LEDK \cite{kondor2002diffusion}, MEDK \cite{chen2014disease}, MDK \cite{fouss2006experimental}, RLK \cite{chebotarev2006matrix}, etc. These graph node kernels often show relatively low discriminative capacity, especially in cases of working with sparse graphs or graphs whose high numbers of missing links, since they share following limitations. First information is processed in an additive and independent fashion which prevents them from accurately modeling the configuration of each node's context. Second, they do not take into account information associated to nodes of graphs when they are available. These additional information normally provide a complement to graph topology. Therefore, they are potential to use to improve the expressiveness of graph node kernels.

We propose an effective convolutional graph node kernel, named \textit{Conjunctive disjunctive node kernel} (CDNK) which is able to \textit{i}) effectively exploit the nodes' context, \textit{ii}) process with information sticked on nodes of graphs.

Predictive systems for gene-disease associations are often based on a notion of similarity between genes. A common strategy is to encode relations between genes as a network and to use graph based techniques to make useful inferences. In the last decades, a number of graph kernel methods have been proposed that directly exploit transitive properties in biological networks. The prototypical method is the Diffusion kernel (DK) \cite{kondor2002diffusion} inspired by the heat diffusion phenomenon. The key idea is to allow a given amount of {\em heat} placed on nodes to {\em diffuse} through the edges. The similarity between two nodes $v_{i}, v_{j}$ is then defined as the amount of heat starting from $v_{i}$ and reaching $v_{j}$ within a given time interval. In DK the heat flow is proportional to the number of paths connecting two nodes, which introduces an undesired bias that penalize peripheral nodes w.r.t. central ones. This problem is tackled by a modified version of DK called Markov exponential diffusion kernel (MED) \cite{chen2014disease} where a Markov matrix replaces the Laplacian matrix.  Another kernel called Markov diffusion kernel (MD) \cite{fouss2006experimental}, exploits instead the notion of {\em diffusion distances}, a measure of similarity between patterns of heat diffusion. The Regularized Laplacian kernel (RL) \cite{chebotarev2006matrix} represents instead a normalized version of the random walk with restart model and defines the node similarity as the number of paths connecting two nodes with different lengths. All these approaches can be applied to dense networks with high degree nodes. A drawback of these approaches is however their relatively low discriminative capacity. This is in part due to the fact that information is processed in an additive and independent fashion which prevents them from accurately modeling the configuration of each gene's context. To address this issue here we propose to employ a {\em decompositional} graph kernel (DGK) \cite{haussler1999convolution} technique in which the similarity function between graphs can be formed by decomposing each graph into subgraphs and by devising a valid local kernel between the subgraphs. To exploit its higher discriminative capacity we first decompose the network in a collection of connected sparse graphs and then we develop a suitable kernel, that we call Conjunctive Disjunctive Node Kernel (CDNK).

\cite{mordelet2011prodige}, \cite{chen2014disease}, \cite{valentini2014extensive}
\section{Background}
In this section, we first introduce definitions and notations that are used to define our proposed method. We then present graph node kernels and problems of state of the art graph node kernel.
\subsection{Definitions and Notations}
A graph is a structure, notated as $G=(\mathbb{V},\mathbb{E})$ where $\mathbb{V}$ is the set of nodes and $\mathbb{E}$ is the set of links (edges).

We define the shortest path between $u$ and $v$, notated as $\mathcal{D}(u,v)$, as the number of edges on the shortest path between them. The \textit{neighborhood} of a node $u$ with radius $r$, $N_r(u) = \lbrace v\ |\ \mathcal{D}(u,v) \leq r \rbrace$, is the set of nodes at distance no greater than $r$ from $u$. The corresponding \textit{neighborhood subgraph} $\mathcal{N}_{r}^{u}$ is the  subgraph induced by the neighborhood (i.e. considering all the edges with endpoints in $N_r(u)$). The \textit{degree} of a node $u$, $deg(u) = |\mathcal{N}_{1}^{u}|$, is the cardinality of its neighborhood for $r=1$. The maximum node degree in the graph $G$ is $deg(G)$.

\begin{definition}{}
\textit{An adjacency matrix $\textbf{A}$ is a symmetric matrix used to characterize the direct links between vertices $v_{i}$ and $v_{j}$ in the graph. Any entry $A_{ij}$ is equal to $w_{ij}$ when there exists a link connecting $v_{i}$ and $v_{j}$, and is 0 otherwise.}. 
\end{definition}

\begin{definition}{}
\textit{The Laplacian matrix $\textbf{L}$ is defined as $\textbf{L} = \textbf{D}-\textbf{A}$, where $\textbf{D}$ is the diagonal matrix with non-null entries equal to the summation over the corresponding row of the adjacency matrix, i.e. $\textbf{D}_{ii}=\sum_j \textbf{A}_{ij}$.}
\end{definition}

\begin{definition}{}
\textit{Transition matrix of a graph $G$, notated as $P$, is a matrix whose each element $P_{ij} = A_{ij}/\sum_{i}^{}A_{ij}$ showing the probability of stepping on node $j$ from node $i$.}
\end{definition}

\subsection{Kernels defined on graphs}
Kernel methods, whose kernels, have emerged as one of the most powerful framework in machine learning which has been successfully applied in various fields. Kernels can be considered as similarity functions that allow to measure the similarity between any couple of objects in feature space and with any type of data representation.

Formally, a kernel, $k$, on $\mathbb{X} \times \mathbb{X}$ with $\mathbb{X}$ is a set of objects, is a function defined as $k: \mathbb{X} \times \mathbb{X}\longmapsto \mathbb{R}$ such that \textit{i}) $k$ is symmetric, meaning that $k(x_1,x_2) = k(x_2,x_1)$, where $x_1, x_2 \in \mathbb{X}$, and \textit{ii}) $k$ is positive semi-definite, that is $\sum_{i=1}^{N}\sum_{j=1}^{N} c_i c_jk(x_i,x_j) \geq 0$ for any $N>0$, $c_i, c_j \in \mathbb{R}$, and  $x_i, x_j \in \mathbb{X}$.

When the input of kernels is a set of graphs, we have graph kernels and when it is a set of graph nodes, we refer kernels as graph node kernels.
The problem of developing kernels defined on structured data in general and graphs in particular is made possible thanks to release of decomposition kernels proposed in \cite{haussler1999convolution}. Examples are kernels proposed in \cite{shervashidze2011weisfeiler, gartner2003survey, borgwardt2005shortest, costa2010fast}. Following we describe NSPDK \cite{costa2010fast}, a convolutional graph kernel which shows state of performance and it is later on used to define our proposed graph node kernel.

\subsubsection*{The Neighborhood Subgraph Pairwise Distance Kernel}
\label{sec:ndpdk}
The NSPDK \cite{costa2010fast} is an instance of convolution kernel \cite{haussler1999convolution} where given a graph $G \in \mathcal{G}$ and two rooted graphs $A_u, B_v$, the relation $R_{r,d}(A_u, B_v, G)$ is true {\em iff} $A_u \cong \mathcal{N}_r^u$ is (up to isomorphism $\cong$) a neighborhood subgraph of radius $r$ of $G$ and so is $B_v \cong  \mathcal{N}_r^v$, with roots at distance $\mathcal{D}(u,v)= d$. Figure \ref{sec:ndpdk} illustrates a pairwise neighborhood subgraph. We denote $R^{-1}$ as the inverse relation that returns all pairs of neighborhoods of radius $r$ at distance $d$ in $G$, $R^{-1}_{r,d}(G) = \lbrace A_u, B_v | R_{r,d}(A_u,B_v,G)=true\rbrace$. The kernel $\kappa_{r,d}$ over $\mathcal{G} \times \mathcal{G}$, counts the number of such fragments in common in two input graphs: 
\begin{center}
$\kappa_{r,d}(G,G^{'}) = 
\!\!\!\!\!\!\!\!\!\!\!\! 
\sum\limits_{\substack{A_u, B_v \ \in \ R_{r,d}^{-1}(G) \\ 
{A'}_{u'}, {B'}_{v'} \ \in \ R_{r,d}^{-1}(G')
}} \!\!\!\!\!\!\!\!\!\!\!\!  { { \textbf{1}_{A_{u} \cong A'_{u'}}} \cdot {
\textbf{1}_{B_{v} \cong B'_{v'}}} }$, 
\end{center} 
\noindent where $\textbf{1}_{A \cong B}$ is the \textit{exact matching function} that returns 1 if $A$ is
isomorphic to $B$ and 0 otherwise.  Finally, the NSPDK is defined as $K(G,G') = \sum\limits_{r}{\sum\limits_{d}{\kappa_{r,d}(G,G')}}$, where for efficiency reasons, the values of $r$ and $d$ are upper bounded to a given maximal $r^*$ and $d^*$, respectively.
\begin{figure}
\centering
\includegraphics[width=.9\textwidth]{nspdk}
\caption{Pairwise neighborhood subgraphs for the ``red'' node with $r=1$ and $d=3$}
\label{fig:nspdk}
\end{figure}

The state of the art kernels are based on diffusion phenomenon. In other words, they take paths connecting two nodes into account when measuring their similarity. In the following, we shortly introduce some of the most used graph node kernels.

\subsubsection{Laplacian exponential diffusion kernel}
One of the most well-known graph node kernels is the Laplacian exponential diffusion kernel \textbf{LEDK}, as it is widely used for exploiting discrete structures in general and graphs in particular. On the basis of the heat diffusion dynamics, Kondor and Lafferty proposed \textbf{LEDK} in \cite{kondor2002diffusion}: imagine to initialize each vertex with a given amount of heat and let it flow through the edges until an arbitrary instant of time. The similarity between any vertex couple $v_{i}$, $v_{j}$ is the amount of heat starting from $v_{i}$ and reaching $v_{j}$ within the given time. Therefore, \textbf{LEDK} can capture the long range relationship between vertices of a graph to define the global similarities. Below is the formula to compute \textbf{LEDK} values:
\begin{equation} 
\label{LEDK-formula}
K = e^{-\beta \textbf{L}} = \textbf{I} - \beta \textbf{L} + \frac{\beta \textbf{L}^{2}}{2!} - ...
\end{equation}
where $\beta$ is the diffusion parameter and is used to control the rate of diffusion and $\textbf{I}$ is the identity matrix. Choosing a consistent value for $\beta$ is very important: on the one side, if $\beta$ is too small, the local information cannot be diffused effectively and, on the other side, if it is too large, the local information will be lost. \textbf{LEDK} is positive semi-definite as proved in \cite{kondor2002diffusion}.
\subsubsection{Exponential diffusion kernel}
In \textbf{LEDK}, the similarity values between high degree vertices are generally higher compared to those between low degree ones. Intuitively, the more paths connect two vertices, the more heat can flow between them. This could be problematic since peripheral nodes have unbalanced similarities with respect to central nodes. In order to make the strength of individual vertices comparable, a modified version of \textbf{LEDK} is introduced by Chen et al in \cite{chen2014disease}, called Markov exponential diffusion kernel \textbf{MEDK} and given by the following formula:
\begin{equation} \label{MEDK-formula}
\textbf{K} = e^{-\beta \textbf{M}}
\end{equation}
The difference with respect to the Laplacian diffusion kernel is the replacement of \textbf{L} by the matrix $\textbf{M}=(\textbf{D}-\textbf{A}-n\textbf{I})/n$ where $n$ is the total number of vertices in graph. The role of $\beta$ is the same as for \textbf{LEDK}.
\subsubsection{Markov diffusion kernel}
The original Markov diffusion kernel \textbf{MDK} was introduced by Fouss et al. \cite{fouss2006experimental} and exploits the idea of diffusion distance, which is a measure of how similar the pattern of heat diffusion is among a pair of initialized nodes. In other words, it expresses how much nodes "influence" each other in a similar fashion. If their diffusion ways are alike, the similarity will be high and, vice versa, it will be low if they diffuse differently. This kernel is computed starting from the transition matrix \textbf{P} and by defining $\textbf{Z}(t) = \frac{1}{t}\sum_{\tau=1}^{t}\textbf{P}^{\tau}$, as follows:
\begin{equation} 
\label{MDK-formula}
\textbf{K} = \textbf{Z}(t)\textbf{Z}^{\top}(t)
\end{equation}

\subsubsection{Regularized Laplacian kernel}
Another popular graph node kernel function used in graph mining is the regularized Laplacian kernel \textbf{RLK}. This kernel function is introduced by Chebotarev and Shamis in \cite{chebotarev2006matrix} and represents a normalized version of the random walk with restart model. It is defined as follows:
\begin{equation} 
\label{RLK-formula}
\textbf{K} = \sum_{n=0}^{\infty}\beta^{n}(-\textbf{L})^n = (\textbf{I} + \beta \textbf{L})^{-1}
\end{equation}
where the parameter $\beta$ is again the diffusion parameter. \textbf{RLK} counts the paths connecting two nodes on the graph induced by taking \textbf{-L} as the adjacency matrix, regardless of the path length. Thus, a non-zero value is assigned to any couple of nodes as long as they are connected by any indirect path. \textbf{RLK} remains a relatedness measure even when diffusion factor is large, by virtue of the negative weights assigned to self-loops.

\section{Method}
We start from the type of similarity notion computed by a neighborhood based decomposition kernel between graph instances \cite{costa2010fast}, NSPDK, and adapt it to form a convolutional graph node kernel which aims at expressing the similarity between nodes in a single graph. Our intended graph node kernel takes an undirected, labeled graph as its input.

Given an input graph $G=(V,E)$, our method consists of three steps. In the first step, \textit{node labeling}, a labeling function is proposed to assign label for every node of the graph. This step is necessary since we desire to design a convolutional kernel which deals with labeled graphs. The second step includes a graph decompostion procedure which allows to decompose the graph into a set of linked sparse connected components. In this procedure, we define two different kinds of link: \textit{conjunctive} and \textit{disjunctive} in which they are treated in distinct manners later on. In the final step, the similarity between any node couple $u$ and $v$ is computed by adopting NSPDK on two neighborhood subgraphs rooted as $u$ and $v$ which extracted from the resulted graph after decomposing. Following, we describe each step in detail.

\subsection{Node Labeling} 
\label{sec:lab}
We are interested in investigating biological networks in general and genetic networks in specific. Therefore, in this section, we propose two different approaches to label for genetic graphs. It is worth to notice that genetic networks typically represent genes as nodes labeled with a gene identifier. Here we take a different approach and use the node labels to encode abstract information about the genes. In this way downstream machine learning algorithms can generalize from similar examples and allow the identification of overlooked but related genes. We experiment with two types of information: \textit{i}) topological information and \textit{ii}) functional information.\\

\textit{Topological labels} This information is simply based on the connectivity degree of the gene. The idea is that genes that have the same number of connections are more similar than genes with a different connectivity. Here we propose a node labeling function $\mathcal{L}$ that assigns labels for nodes by considering their degrees, such that nodes with similar degrees are assigned with a same label. Therefore, our function assigns the degree for nodes $u$ having degree less than or equal to a user defined threshold $T$ ($T=5$ in our experimental evaluation). However degree values larger than $T$ are subsequently discretized into $k$ levels. We treat nodes differently between nodes with high degrees and low degrees since we envisage that the properties of low-degree nodes whose different node degrees are more dissimilar in comparison with the ones of high degrees. Formally, the labeling function is defined as:
\begin{center}
$\mathcal{L}(u) = \left\{
	\begin{array}{ll}
		deg(u),\  & \mbox{if } deg(u) \leq T \\
		T+i,\ & \mbox{if } deg(u) > T
	\end{array}
\right.$,
\end{center}
where $i = \lceil \frac{deg(u)-T}{bin}\rceil$, $bin = \frac{deg(G)-T}{p - T}$ and $p$ ($p > T$) is the maximum number of symbols used. The value of $p$ depends on the degree distribution and can be tuned as a hyperparameter of the approach.\\

It is often that nodes in the graph are associated to extra information. These information can be considered as the complement to the graph topological information. Therefore, utilizing both information source are supposed to improve the performance of graph node kernels. In the following, we propose two different approaches to process with this kind of information. In the first approach, it is converted and used as the functional discrete labels for nodes, while it is processed and utilized as the real valued vectors in the second one. In both cases, we use \textit{Gene Ontology} \cite{gene2004gene} as an example of extra information associated to the graph nodes.\\

\textit{Functional discrete labels}
We use a gene ontology to construct binary vectors representing a bag-of-words encoding for each gene (\textit{i.e. each element of a binary vector is equal to 1 if its corresponding GO-term is associated to the gene, and is equal to 0 otherwise.}). The resulting vectors are then clustered using the k-means algorithm  into a user defined number of classes $K$ (tuned as a hyperparameter of the approach), so that genes with similar description profiles receive the same class identifier as label.\\

\textit{Functional real valued vector labels}
In addition to encoding the functional information as a discrete label we add a richer description by computing the similarity vector w.r.t. to each cluster. In this way we can fully exploit the latent description of the genes in terms of the different functional groups captured by the clustering procedure. Formally, given a vector $v \in I\!R^{26501}$ we compute a similarity vector $S(v)= {s_1, s_2, \ldots s_K}$ with entries $s_i = \frac{1}{1+ \ell(v,c_i)}$ where $\ell(v,c_i)$ is the Euclidean distance of $v$ from the center of the $i^{th}$ cluster $c_i = \frac{1}{|C_i|}\sum_{x \in C_i} x$ computed as the geometric mean of the elements in the cluster $C_i$.

\subsection{Network Decomposition} 
In genetic networks it is not uncommon to find nodes with high degrees. Unfortunately these cases cannot be effectively processed by a neighborhood based decomposition kernel (see convolution kernels presented in \ref{convolution-kernels}) since these are based on the notion of exact matches and the probability of finding identical neighborhoods decreases exponentially as the degree increases. This means that in a finite network it quickly becomes impossible to find any match and hence learn or generalize at all. Therefore, we propose a procedure to ``sparsify'' the network that is observed by the neighborhood kernel. In practice, we mostly keep the same the cardinality of the edge set. However we mark the edges with special attributes so that kernel is able to treat them differently when computing. The result is a procedure that decomposes the network in a linked collection of sparse sub-networks where each node has a reduced connectivity when considering the edges of a specific type. However the other edges are still available to connect the various sub-networks. We distinguish two types of edges: {\em conjunctive} and {\em disjunctive} edges. Nodes linked by conjunctive edges are going to be used jointly to define the notion of context and will be visible to the neighborhood graph kernel. Nodes linked by disjunctive edges are instead used to define features based only on the pairwise co-occurrence of the genes at the endpoints and are processed by our novel kernel.

\textit{Iterative k-core decomposition} \cite{alvarez2005k}: The node set is partitioned in two groups on the basis of the degree of each node w.r.t. a threshold degree $D$, the first part contains all nodes with degree smaller than or equal to $D$ and the second part the remaining ones. The node partition is used to induce the ``conjunctive'' vs ``disjunctive'' notion for the edge partition: edges that have both endpoints in the same part are marked as conjunctive, otherwise they are marked as disjunctive. We apply the k-core decomposition iteratively, where at each iteration we consider only the graph induced by the conjunctive edges. We stop iterating the decomposition after a user defined number of steps. Note that this decomposition does not alter the cardinality of the edge set, it is simply a procedure to mark each edge with the attribute conjunctive or disjunctive.

\begin{figure}
\centering
\includegraphics[width=.9\textwidth]{k_core}
\caption{K-core decomposition with degree threshold $D = 4$}
\label{fig:kcore-decomposition}
\end{figure}

\textit{Clique decomposition} \cite{tarjan1985decomposition}: To model the notion that nodes in a clique are tightly related, we summarize the whole clique with a new ``representative'' node. All the cliques (completely connected subgraphs) with a number of nodes greater than or equal to a given threshold size $C$ are identified. The endpoints of all edges incident on the clique's nodes are moved to the representative node. Disjunctive edges are introduced to connect each node in the clique to the representative. Finally all edges with both endpoints in the clique are removed.

In our work a network is transformed by applying first the iterative k-core decomposition and then the clique decomposition.

\begin{figure}
\centering
\includegraphics[width=.9\textwidth]{clique}
\caption{Clique decomposition with threshold $C = 4$}
\label{fig:clique-decomposition}
\end{figure}

\subsection{The Conjunctive Disjunctive Node Kernel}
We start from the Neighborhood Subgraph Pairwise Distance Kernel (NSPDK) \cite{costa2010fast} and adapt it to express the similarity between nodes in a single network. The key idea in NSPDK is to decompose graphs in small fragments and count how many pairs of fragments are shared between two instances. We introduce two improvements: \textit{i}) we partition the features according to the individual node's neighborhood, and \textit{ii}) we introduce a  distinction between ``disjunctive'' and ``conjunctive'' edges.

We define a node kernel $K(G_u,G_{u'})$ between two copies of the same network $G$ where we distinguish the nodes $u$ and $u'$ respectively. The idea is to define the features of a node $u$ as the subset of NSPDK features that always have the node $u$ as one of the roots. In addition we distinguish between two types of edges, called {\em conjunctive} and {\em disjunctive} edges. When computing distances to induce neighborhood subgraphs, only conjunctive edges are considered. When choosing the
pair of neighborhoods to form a single feature, we additionally consider roots $u$ and $v$ that are not at distance $d$ but such that $u$ is connected to $w$ via a disjunctive edge and such that $w$ is at distance $d$ from $v$ (Figure \ref{fig:cdnk} is an illustration). In this way disjunctive edges can still allow an {\em information flow} even if their
endpoints are only considered in a pairwise fashion and not jointly.

Formally, we define two relations: the \textit{conjunctive relation} $R^{\wedge}_{r,d}(A_u, B_v, G_u)$ identical to the NSPDK relation $R_{r,d}(A_u, B_v, G)$,  and (ii) $\mathcal{D}(u,v)= d$; the \textit{disjunctive relation} $R_{r,d}^{\vee}(A_u, B_v, G_u)$ is true {\em iff} (i)  $A_u \cong \mathcal{N}_r^u$ and $B_v \cong \mathcal{N}_r^v$ are true, (ii) $\exists w$ s.t. $\mathcal{D}(w,v)= d$, and (iii) $(u,w)$ is a disjunctive edge. We define $\kappa_{r,d}$ on the  inverse relations ${R^{\wedge}_{r,d}}^{ -1}$ and ${R^{\vee}_{r,d}}^{ -1}$:
\begin{center}
 $\kappa_{r,d}(G_u,G_{u'}) = \!\!\!\!\!\!\!\!\!\!\!\!
 \sum\limits_{\substack {A_u,{B}_{v} \in {R_{r,d}^{\wedge}}^{ -1}(G_u) \\ A'_{u'},{B'}_{v'} \in {R_{r,d}^{\wedge}}^{ -1}(G_{u'}) }} \!\!\!\!\!\!\!\!\!\!\!\!
  { \textbf{1}_{A_u \cong A'_{u'}} \cdot { \textbf{1}_{B_{v} \cong B'_{v'}}}}
+ \!\!\!\!\!\!\!\!\!\!\!\!
 \sum\limits_{\substack {A_u,{B}_{v} \in {R_{r,d}^{\vee}}^{ -1}(G_u) \\
  A'_{u'},{B'}_{v'} \in \ {R_{r,d}^{\vee}}^{ -1}(G_{u'}) }} \!\!\!\!\!\!\!\!\!\!\!\!
  { \textbf{1}_{A_u \cong A'_{u'}} \cdot { \textbf{1}_{B_{v} \cong B'_{v'}}}}
  $.
\end{center}
The CDNK is finally defined as $K(G_u,G_{u'}) = \sum\limits_{r}{\sum\limits_{d}{\kappa_{r,d}(G_u,G_{u'})}}$, where once again for efficiency reasons, the values of $r$ and $d$ are upper bounded to a given maximal $r^*$ and $d^*$.

\begin{figure}
\centering
\includegraphics[width=.9\textwidth]{cdnk}
\caption{Pairwise neighborhood subgraphs for the ``red'' node with $r=1$ and $d=3$ using ``conjuction'' and ``disjuctive'' edges}
\label{fig:cdnk}
\end{figure}

\subsection{Real valued node information}
In order to integrate the information of real vectors we proceed as follows. We compute a sparse vector representation for the neighborhood graph rooted in node $v$ following \cite{costa2010fast}: for each neighborhood subgraph we calculate the quasi-isomorphism certificate hash code; we then combine the hashes for the pair of neighborhoods and use the resulting integer as a feature indicator. This yields a direct sparse vector representation (associated to node $u$ in graph $G$) $f: G_u \longmapsto I\!R^L$ where $L \approx 10K-1M$. Given the real valued vector information (associated to node $u$ in graph $G$) $g: G_u \longmapsto I\!R^K$ computed as the multi-class similarity to the $K$ clusters (c.f.r. Section \ref{sec:lab}), we update the computation of CDNK considering the discrete convolution of the discrete information with the real valued information:   
\begin{equation}
K(G_u,G_{u'}) = \left< f(G_u) \bigotimes g(G_u), f(G_{u'}) \bigotimes g(G_{u'}) \right>,
\end{equation}
where the discrete convolution is defined as: 
$(f \bigotimes g)[n]=\sum _{m=0}^{K-1}f[n-m]g[m].$
In words, we are starting a scaled copy of the real valued vector at the position indicated by each feature computed on the basis of the discrete information. Intuitively, when both the real valued and the discrete information match, the kernel computes a large similarity, but if there is a discrepancy in either one of the sources of information, the similarity will be penalized.

\section{Empirical Evaluation}
In this section, we desire to evaluate the performance of CDNK and compare it with the state of the art graph node kernels. In order to do that, we employ disease gene prioritization problem where the aim is to build a learning system which allows to prioritize candidate genes based on their probabilities of being associated to a given genetic disease.

Given a genetic graph and a list of training genes for a genetic disease (training set consists both positive and negative genes), we first apply  a graph node kernel to compute kernel matrix. This kernel matrix together with training gene set are used as input of a learning algorithm to construct a system. The trained system is used to prioritize for candidate genes.

To compare the performance of graph node kernels, we fix the learning algorithm and substitute graph node kernels. The performance of the system obtained in different cases reflect the performance of graph node kernels.

Datasets: We carry out in two separate networks derived from BioGPS and Pathways datasets. Following are the short description of these networks.
We carry out experiments on two separate datasets and we follow experiment's procedure carried out in \cite{chen2014disease} where 
12 diseases \cite{goh2007human} are used in each disease is associated to at least 30 positive genes. For each disease, we construct a positive set $\mathcal{P}$ with all positive (confirmed) disease genes, and a negative set $\mathcal{N}$ which contains random genes associated at least to one disease class which is not related to the class that is defining the positive set. In \cite{chen2014disease} the ratio between the dataset sizes is chosen as $\vert \mathcal{N} \vert = \frac{1}{2} \vert \mathcal{P} \vert$. The predictive performance of each method is evaluated via a leave-one-out cross validation: one gene is kept out in turn and the rest are used to train an SVM model. We compute a decision score $q_i$ for the test gene $g_i$ as the top percentage value of score $s_i$ among all candidate gene scores. We collect all decision scores for every gene in the training set to form a global decision score list on which we compute the AUC-ROC.

\begin{itemize}
\item \textbf{BioGPS:} A gene co-expression network is constructed from BioGPS dataset, which contains 79 tissues, measured with the Affymetrix U133A array. Edges are inserted when the pairwise Pearson correlation coefficient (PCC) between genes is larger than 0.5.

\item \textbf{Pathways:} Pathway information is retrieved from KEGG, Reactome, PharmGKB and the Pathway Interaction Database. If a couple of proteins co-participate in any pathway, the two corresponding genes are linked.  
\end{itemize}

\textbf{Model Selection:}
The hyper parameters of the various methods are tuned using a k-fold strategy. However due to the non i.i.d. nature of the problem, we employ a stronger setup to ensure no information leakage. The dataset on which we are validating the performance is never subsequently used in the predictive performance estimate. The values for diffusion parameter in DK and MED are sampled in $\lbrace 10^{-3}, 10^{-3}, 10^{-2}, 10^{-1} \rbrace$, time steps in MD in $\lbrace 1, 10, 100 \rbrace$ and RL parameter in $\lbrace 1, 4, 7 \rbrace$. For CDNK, the degree threshold values are sampled in $\lbrace 10,\ 15,\ 20 \rbrace$, clique size threshold in $\lbrace 4,\ 5 \rbrace$, maximum radius in $\lbrace 1,\ 2 \rbrace$, maximum distance in $\lbrace 2,\ 3,\ 4 \rbrace$, number of clusters $K$ in $\lbrace 5,\ 7 \rbrace$. Finally, the regularization trade off parameter $C$ for the SVM is sampled in $\lbrace 10^{-5},  \ 10^{-4}, \ 10^{-3},\ 10^{-2},\ 10^{-1}, 1,\ 10,\ 10^2 \rbrace$.

Table \ref{table_biogps} and \ref{table_pathways} show the AUC performance of the system corresponding to different cases of employing graph node kernels. Predictive models based on CDNK variations are ranked in top positions in all cases on BioGPS dataset and 9 out of 11 cases on Pathways when compared to diffusion based kernels. CDNK is ranking first when considering both the average AUC-ROC and the average rank with a difference, w.r.t. the best diffusion kernel, ranging from 5.4$\%$ to 10$\%$ and from 1.2$\%$ to 3.2$\%$ on BioGPS and Pathways, respectively. Regarding the variations of CDNK, the integration of real valued vector information improves the performance in most cases on using discrete labels only. The improvement in average ranges from 1.3$\%$ to 4.6$\%$. Note that also in the case where only topological information is used to induce the discrete label (CDNK3 in the Table \ref{table_biogps} and \ref{table_pathways}), the proposed approach improves on state-of-the-art.

\begin{table}
\caption{\textit {Predictive performance on 11 gene-disease associations in percentage using network induced by the BioGPS. Best results in bold. We report the AUC ROC and the rank for each kernel method.  CDNK1 = ontology for discrete labels, CDNK2 = ontology for both discrete and vector labels, CDNK3 = node degree for discrete labels and CDNK4 = node degree for discrete labels and ontology for vector label.}}
\vspace*{5pt}
\centering
\setlength{\tabcolsep}{1mm}
\begin{tabular}{|c|c|c|c|c|c|c|c|c|}
\hline
         & \multicolumn{8}{c|}{\textbf{BioGPS}}\\
 \hline
Disease & DK & MD & MED & RL & CDNK1 & CDNK2 & CDNK3 & CDNK4\\

 \hline
1 & 51.9/8 & 57.4/7 & 59.0/6 & 59.2/5 & 65.1/4 & 69.5/2 & 69.3/3 & \textbf{70.3/1} \\
2 & 81.7/5 & 78.5/6 & 75.2/7 & 75.0/8 & 88.3/2 & \textbf{88.8/1} & 85.1/4 & 86.8/3 \\
3 & 64.3/7 & 59.6/8 & 71.6/3 & 71.8/2 & 65.5/5 & \textbf{72.5/1} & 64.7/6 & 66.4/4 \\
4 & 65.3/7 & 58.2/8 & 67.8/6 & 67.8/5 & 71.9/4 & \textbf{78.7/1} & 73.9/3 & 77.0/2 \\
5 & 64.0/8 & 64.1/7 & 66.5/5 & 66.2/6 & 75.9/4 & 76.2/3 & 76.9/2 & \textbf{77.4/1} \\
6 & 74.6/5 & 70.2/8 & 71.0/7 & 71.2/6 & 79.3/2 & \textbf{83.7/1} & 76.7/4 & 79.0/3 \\
7 & 73.0/5 & 66.7/8 & 75.4/3 & 75.6/2 & 68.8/6 & 73.9/4 & 67.3/7 & \textbf{76.9/1} \\
8 & 74.4/8 & 76.8/3 & 76.2/5 & 76.4/4 & 74.7/7 & \textbf{77.7/1} & 76.0/6 & 76.8/2 \\
9 & 71.5/2 & 65.6/8 & 67.7/5 & 69.9/3 & 66.8/7 & \textbf{71.7/1} & 67.1/6 & 68.1/4 \\
10 & 54.0/6 & 50.3/8 & 56.1/5 & 51.1/7 & 77.6/4 & \textbf{82.7/1} & 80.5/2 & 80.0/3 \\
11 & 58.2/7 & 51.3/8 & 59.3/6 & 59.3/5 & 71.8/4 & \textbf{80.2/1} & 75.3/3 & 77.1/2 \\
\hline
$\overline{AUC}$ & 66.6 & 63.5 & 67.8 & 67.6 & 73.2 & \textbf{77.8} & 73.9 & 76.0 \\
$\overline{Rank}$ & 	6.18 & 7.18 & 5.27 & 4.82 & 4.45 & \textbf{1.55} & 4.18 & 2.36 \\
 \hline 
\end{tabular}
\label{table_biogps}
\end{table}
%==================================================
\begin{table}
\caption{\textit {Predictive performance on 11 gene-disease associations in percentage using network induced by the Pathways. Best results in bold. We report the AUC ROC and the rank for each kernel method.  CDNK1 = ontology for discrete labels, CDNK2 = ontology for both discrete and vector labels, CDNK3 = node degree for discrete labels and CDNK4 = node degree for discrete labels and ontology for vector label.}}
\vspace*{5pt}
\centering
\setlength{\tabcolsep}{1mm}
\begin{tabular}{|c|c|c|c|c|c|c|c|c|}
\hline
         & \multicolumn{8}{c|}{\textbf{Pathways}} \\
 \hline
Disease & DK & MD & MED & RL & CDNK1 & CDNK2 & CDNK3 & CDNK4 \\

 \hline
1 & 74.7/8 & 76.4/7 & 78.7/6 & 78.8/5 & 80.2/3 & \textbf{82.4/1} & 80.6/2 & 79.4/4 \\
2 & 55.1/8 & 64.9/7 & 76.6/6 & 76.6/5 & 81.1/2 & 80.3/3 & 79.8/4 & \textbf{82.2/1} \\
3 & 55.0/8 & 62.7/7 & 64.1/5 & 65.6/4 & 67.1/3 & 63.6/6 & 68.5/2 & \textbf{71.5/1} \\
4 & 54.3/8 & 65.2/7 & \textbf{73.7/1} & 73.7/2 & 66.1/5 & 68.1/3 & 67.5/4 & 65.6/6 \\
5 & 52.9/8 & 55.7/7 & 62.7/5 & 62.7/6 & 68.3/2 & \textbf{69.6/1} & 66.3/4 & 66.8/3 \\
6 & 83.4/8 & 92.7/7 & 96.5/2 & \textbf{96.5/1} & 93.0/6 & 94.1/5 & 94.4/4 & 95.5/3 \\
7 & 84.5/8 & 88.3/7 & 89.4/3 & 89.5/2 & 88.5/6 & 88.5/5 & 88.7/4 & \textbf{90.5/1} \\
8 & 53.7/8 & 65.6/7 & 72.0/6 & 72.3/5 & 72.5/3 & 72.5/2 & 72.3/4 & \textbf{76.5/1} \\
9 & 52.5/8 & 64.9/5 & 64.2/7 & 64.2/6 & \textbf{81.3/1} & 81.0/2 & 79.6/3 & 78.8/4 \\
10 & 68.8/6 & 65.4/8 & 74.4/5 & 74.4/4 & 66.9/7 & 76.8/2 & 76.1/3 & \textbf{79.5/1} \\
11 & 53.7/8 & 69.2/7 & 74.6/5 & 74.1/6 & 77.0/2 & \textbf{78.7/1} & 75.4/4 & 77.0/3 \\
\hline
$\overline{AUC}$ & 62.6 & 70.1 & 75.2 & 75.3 & 76.5 & 77.8 & 77.2 & \textbf{78.5} \\
$\overline{Rank}$ & 7.82 & 6.91 & 4.64 & 4.18 & 3.64 & 2.82 & 3.45 & \textbf{2.55} \\
 \hline 
\end{tabular}
\label{table_pathways}
\end{table}

\section{Conclusions}
We have shown how decomposing a network in a set of connected sparse graphs allows us to take advantage of the discriminative power of CDNK, a novel decomposition kernel, to achieve state-of-the-art results. Moreover, we have also introduced the way to integrate ``side'' information in form of real valued vectors when it is available on graph to get even better performance of CDNK. In future work we will investigate how to \textit{i}) decompose networks in a data driven way and \textit{ii}) extend the CDNK approach to gene-disease association problems exploiting multiple heterogeneous information sources in a joint way.
%\section{The Elsevier article class}
%
%\paragraph{Installation} If the document class \emph{elsarticle} is not available on your computer, you can download and install the system package \emph{texlive-publishers} (Linux) or install the \LaTeX\ package \emph{elsarticle} using the package manager of your \TeX\ installation, which is typically \TeX\ Live or Mik\TeX.
%
%\paragraph{Usage} Once the package is properly installed, you can use the document class \emph{elsarticle} to create a manuscript. Please make sure that your manuscript follows the guidelines in the Guide for Authors of the relevant journal. It is not necessary to typeset your manuscript in exactly the same way as an article, unless you are submitting to a camera-ready copy (CRC) journal.
%
%\paragraph{Functionality} The Elsevier article class is based on the standard article class and supports almost all of the functionality of that class. In addition, it features commands and options to format the
%\begin{itemize}
%\item document style
%\item baselineskip
%\item front matter
%\item keywords and MSC codes
%\item theorems, definitions and proofs
%\item lables of enumerations
%\item citation style and labeling.
%\end{itemize}
%
%\section{Front matter}
%
%The author names and affiliations could be formatted in two ways:
%\begin{enumerate}[(1)]
%\item Group the authors per affiliation.
%\item Use footnotes to indicate the affiliations.
%\end{enumerate}
%See the front matter of this document for examples. You are recommended to conform your choice to the journal you are submitting to.
%
%\section{Bibliography styles}
%
%There are various bibliography styles available. You can select the style of your choice in the preamble of this document. These styles are Elsevier styles based on standard styles like Harvard and Vancouver. Please use Bib\TeX\ to generate your bibliography and include DOIs whenever available.
%
%Here are two sample references: \cite{Feynman1963118,Dirac1953888}.
\section*{Funding}
This work was supported by the University of Padova, Strategic Project BIOINFOGEN.
\section*{References}

\bibliography{mybibfile}

\end{document}

### 2017-10-22 Response_neurocomputing.tex
Add files via upload

--- 2017-10-11 Response_3.tex
@@
 
-\underline{Answer}:
+\underline{Answer}: in LEDK, the similarity values between high degree nodes are generally higher compared to those between low degree ones.
+Intuitively, the more paths connect two vertices, the more heat can flow between them.
+This could be problematic since peripheral nodes have unbalanced similarities with respect to central nodes.
+MEDK adapt LEDK by adding a normalization in order to make the strength of individual vertices comparable.
 
 \item \textit{lines 20-25: what does "process in an additive and independent fashion" mean in this context.
-Similarly as before, please justify your claim about the poor discriminative power of these kernels;
-}
+Similarly as before, please justify your claim about the poor discriminative power of these kernels;}
 
@@
 
-\underline{Answer}:
+\underline{Answer}: the kernel extracts features as pairwise neighiborhood subgraphs.
+If the degree of nodes is high, it leads to big extracted neighborhood subgraphs, and it is difficult to find big graphs which are isomorphic.
+Therefore, we state that the probability to find matches exponentially decreases w.r.t the node degree.
 
@@
 How robust is this clustering to very incomplete information as found in GO databases?
-In a similar fashion, can you provide an evaluation of your
-decomposition on the networks you are working with?}
+In a similar fashion, can you provide an evaluation of your decomposition on the networks you are working with?}
 
 \underline{Answer}: 
-
+\begin{itemize}
+\item We have modified the title of the paper toward to disease gene prioritization.
+\item In the method, we consider the topological labeling as the discrete label
+\item The number of clusters used in the functional labeling is chosen from a set of predefined values through the model selection.
+\end{itemize}
 \subsection*{3} \textit{The paper is extremely badly organized, with some notions (like conjunctive and disjunctive edges) used before they are defined.
@@
 
-\underline{Answer}:
+\underline{Answer}: we have added in the new version of the paper.
 
@@
 
-\underline{Answer}:
+\underline{Answer}: the organization of the paper is modified based on the reviewer's comments.
 
@@
 \underline{Answer}: 
+We have modified based on the reviewer's comment.
 
@@
 \underline{Answer}: 
-
+We already modified based on the reviewer's comment.
 \subsection*{4} \textit{The experimental part is very light.
@@
 
-\underline{Answer}:
-
+\underline{Answer}: we follow the experimental setting, including dataset BioGPS and Pathways, performed in [2]. 
 \section{Reviewer 2}
@@
 
-\underline{Answer}:
+\underline{Answer}: we have modified based on the reviewer's comment.
 
@@
 
-\underline{Answer}: The large number of parameters is a drawback of our propoased graph node kernel. However, some of them are suggested to have smale range of values to be effective.
+\underline{Answer}: We have added a new section named Parameter Space with the content as follow.
+
+The kernel consists of five parameters: the threshold degree $D$, clique size threshold $C$, maximal radius $r^*$, maximal distance $d^*$ and number of clusters $P$ clusters in which their values are in $\mathbb{N}^*$.
+In order to choose the optimal tuple of parameters for kernel in a specific setting, a model selection procedure is normally adopted from a determined subset of parameter space.
+The parameter space of our kernel seems large due to the relatively high number of parameters.
+However, most parameter values are supposed to be in limited natural ranges.
+Therefore, it is actually not too big.
+In particular, the clique size threshold $C$ is recommended to be in $\lbrace 4, 5\rbrace$.
+The maximal radius $r^*$ is set with a value smaller than or equal to $3$ since it will lead to big neighborhood subgraphs.
+The maximal distance $d^*$ is assigned with values less than or equal to $4$ because the value of the maximal shortest distance between nodes in a connected component is often not too high.
+The values of the threshold degree $D$ should neither be too high and too low.
+If it is high, we have to face with high degree node problem, and if it is too low, the obtained graph contains of too sparse connected components.
+Therefore, we suggest the value for $D$ in $\left[6, 20 \right]$
 

### 2017-10-22 neurocomputing.tex
Add files via upload

@@
 \begin{abstract}
-Biological relations encoded data are naturally represented in the form of graphs.
-Therefore, there is a high number of graph-based biological learning systems have been proposed and shown promissing results.
-One of the key points which determines the performance of graph-based learning systems is the denifition of node similarity measure.
+Biological relations encoded in data sources are naturally represented in the form of graphs.
+As a consequence, a high number of graph-based biological learning systems have been proposed and shown promissing results.
+In graph-based learning systems, one of the key points which determines their performance is the denifition of graph node similarity measure.
 Node similarities are normally measured by graph node kernels.
@@
 First they are based on the diffusion phenomenon which does not effectively exploit the nodes' configuration.
-Second they are not able to take into account the available information associated to nodes of graph.
+Second they are not able to proceed the auxiliary information associated to graph nodes.
 
-In this paper, we propose an efficient graph node kernel that not only effectively exploits nodes' context, but also can integrate information which are available on the graph nodes.
-Empirical evaluation show that our proposed graph node kernel show state of the art concerning graph node kernels.
+In this paper, we propose an efficient graph node kernel that not only effectively exploits nodes' context, but also process with additional information which available on the graph nodes.
+Empirical evaluation on disease gene prioritization show that our proposed graph node kernel shows the state of the art concerning graph node kernels.
 
@@
 \section{Introduction}
-Node similarity measure definition is the key that determines the performance of graph-based biological learning systems.
-The state of the art graph node kernels used to measure node similarity, are based on the notion of information diffusion, including LEDK \cite{kondor2002diffusion}, MEDK \cite{chen2014disease}, MDK \cite{fouss2006experimental}, RLK \cite{chebotarev2006matrix}, etc.
-These graph node kernels often show relatively low discriminative capacity, especially in cases of working with sparse graphs or graphs whose high numbers of missing links, since they share following limitations.
+The release of advanced technologies is one of the main reasons for the
+revolution in various scientific research fields. In Biological and Medical
+domain, modern technologies are making it not only easier but also more
+economical than ever to undertake experiments and creating applications.
+As a consequence, a vast amount of biological data in terms of volume
+and type is generated through scientific experiments, published literature,
+high-throughput experiment technology, and computational analysis. This
+huge quantity of data are saved as biological datasets and made discoverable
+through web browsers, application programming interfaces, scalable search
+technology and extensive cross-referencing between databases. Biological
+databases normally contain information about gene function, structure, lo-
+calization, clinical effects of mutations and similarities of biological sequences and structures.
+
+In Biomedical, disease-gene association recovery is a major goal in molecular  that has received much attention from many researchers.
+Despite that fact that a big progress has been made in the last decades, a number of genes known to be related to a genetic disease is normally limited.
+In order to find out the complement set of the known disease gene set, one way is to search for whole genome or specific regions that often contain a large number of suspected genes (candidate genes).
+This is obvious not a good idea as it is expensive not only in term of time consuming but also from financial aspect.
+For this reason, a considerable number of gene prioritization methods have been proposed.
+Predictive systems for gene-disease associations are often based on a notion of similarity between genes.
+A common strategy is to encode relations between genes as a network and to use graph based techniques to make useful inferences \cite{mordelet2011prodige}, \cite{chen2014disease}, \cite{valentini2014extensive}.
+
+One of the key points that determines the performance of graph-based biological learning systems in general and graph-based disease gene prioritization systems in particular is the definition of node similarity measure.
+The node similarity is often measured by graph node kernels \cite{kondor2002diffusion}, \cite{chen2014disease}, \cite{fouss2006experimental}, \cite{chebotarev2006matrix}.
+However, the state of the art graph node kernels used to measure node similarity, are based on the notion of information diffusion.
+These graph node kernels often show relatively low discriminative capacity, especially in cases of working with sparse graphs, graphs have high numbers of missing links, since they share following limitations.
 First information is processed in an additive and independent fashion which prevents them from accurately modeling the configuration of each node's context.
@@
 
-Predictive systems for gene-disease associations are often based on a notion of similarity between genes.
-A common strategy is to encode relations between genes as a network and to use graph based techniques to make useful inferences.
-In the last decades, a number of graph kernel methods have been proposed that directly exploit transitive properties in biological networks.
-The prototypical method is the Diffusion kernel (DK) \cite{kondor2002diffusion} inspired by the heat diffusion phenomenon.
-The key idea is to allow a given amount of {\em heat} placed on nodes to {\em diffuse} through the edges.
-The similarity between two nodes $v_{i}, v_{j}$ is then defined as the amount of heat starting from $v_{i}$ and reaching $v_{j}$ within a given time interval.
-In DK the heat flow is proportional to the number of paths connecting two nodes, which introduces an undesired bias that penalize peripheral nodes w.r.t.
-central ones.
-This problem is tackled by a modified version of DK called Markov exponential diffusion kernel (MED) \cite{chen2014disease} where a Markov matrix replaces the Laplacian matrix.
-Another kernel called Markov diffusion kernel (MD) \cite{fouss2006experimental}, exploits instead the notion of {\em diffusion distances}, a measure of similarity between patterns of heat diffusion.
-The Regularized Laplacian kernel (RL) \cite{chebotarev2006matrix} represents instead a normalized version of the random walk with restart model and defines the node similarity as the number of paths connecting two nodes with different lengths.
-All these approaches can be applied to dense networks with high degree nodes.
-A drawback of these approaches is however their relatively low discriminative capacity.
-This is in part due to the fact that information is processed in an additive and independent fashion which prevents them from accurately modeling the configuration of each gene's context.
-To address this issue here we propose to employ a {\em decompositional} graph kernel (DGK) \cite{haussler1999convolution} technique in which the similarity function between graphs can be formed by decomposing each graph into subgraphs and by devising a valid local kernel between the subgraphs.
-To exploit its higher discriminative capacity we first decompose the network in a collection of connected sparse graphs and then we develop a suitable kernel, that we call Conjunctive Disjunctive Node Kernel (CDNK).
-
-\cite{mordelet2011prodige}, \cite{chen2014disease}, \cite{valentini2014extensive}
 \section{Background}
-In this section, we first introduce definitions and notations that are used to define our proposed method. We then present graph node kernels and problems of state of the art graph node kernel.
+In this section, we first introduce definitions and notations that are used to define our proposed method. We then describe the state-of-the-art concerning graph node kernels.
 \subsection{Definitions and Notations}
-A graph is a structure, notated as $G=(\mathbb{V},\mathbb{E})$ where $\mathbb{V}$ is the set of nodes and $\mathbb{E}$ is the set of links (edges).
+A graph is a structure $G=(\mathbb{V},\mathbb{E}, \mathcal{L}_1, \mathcal{L}_2)$ where $\mathbb{V}$, $\mathbb{E}, \mathcal{L}_1, \mathcal{L}_2$ are the vertex (node) set, link (edge) set, discrete labeling fuction and real vector labeling function, respectively.
+The functions $\mathcal{L}_1, \mathcal{L}_2$ are defined as:
+\begin{itemize}
+\item $\mathcal{L}_1: \mathbb{V} \longmapsto \mathbb{L}$, where $\mathbb{L}$ is a set of discrete labels.
+$\mathcal{L}_1$ assigns a single discrete label $\ell \in \mathbb{L}$ for each ndoe $v \in \mathbb{V}$, $\mathcal{L}_1(v) = \ell$.
+\item $\mathcal{L}_2: \mathbb{V} \longmapsto \mathbb{R}^n$.
+$\mathcal{L}_2$ assigns a single real vector label $(v_1,v_2,\ldots,v_n) \in \mathbb{R}^n$ for each ndoe $v \in \mathbb{V}$, $\mathcal{L}_2(v) = (v_1,v_2,\ldots,v_n)$.
+\end{itemize}
 
@@
 \begin{definition}{}
-\textit{Transition matrix of a graph $G$, notated as $P$, is a matrix whose each element $P_{ij} = A_{ij}/\sum_{i}^{}A_{ij}$ showing the probability of stepping on node $j$ from node $i$.}
+\textit{Transition matrix of a graph $G$, notated as $P$, is a matrix whose each element value, $P_{ij} = A_{ij}/\sum_{i}^{}A_{ij}$, is proportional to the probability of stepping on node $j$ from node $i$.}
 \end{definition}
 
-\subsection{Kernels defined on graphs}
-Kernel methods, whose kernels, have emerged as one of the most powerful framework in machine learning which has been successfully applied in various fields.
-Kernels can be considered as similarity functions that allow to measure the similarity between any couple of objects in feature space and with any type of data representation.
+\subsection{Kernels on graphs}
+Kernel methods have emerged as one of the most powerful framework in machine learning, which has been successfully applied in various domains, due to their modularity.
+In kernel methods, the definition of kernel functions is independent from the design of the learning algorithm.
+A kernel function can be  considered as the similarity measure between example couples in the feature space.
+Interestingly, It can be defined in feature space and with any type of data representation.
 
@@
 
-When the input of kernels is a set of graphs, we have graph kernels and when it is a set of graph nodes, we refer kernels as graph node kernels.
-The problem of developing kernels defined on structured data in general and graphs in particular is made possible thanks to release of decomposition kernels proposed in \cite{haussler1999convolution}.
-Examples are kernels proposed in \cite{shervashidze2011weisfeiler, gartner2003survey, borgwardt2005shortest, costa2010fast}.
+Canonical machine learning methods take vectorial data, data that are represented by vectors of features, as their input.
+However, there are many fields where data are not naturally represented by vectors, but structured forms in which graph is one of the most popular representation.
+Therefore, the task of developing methods which are able to learn from structured data in general or graphs in particular is very important.
+In this thesis, we focus on graphs, a special type of structured data representation.
+
+The problem of developing kernels defined on structured data in general and graphs in particular is made possible thanks to the release of decomposition kernels proposed in \cite{haussler1999convolution}.
+If the input of a kernel is a set of graphs, it is called a graph kernel and if it is a set of graph nodes, the kernel is referred as graph node kernel.
+Both graph kernels and graph node kernels are widely applied to build graph-based learning systems ranging from social networks, to recommendation systems, to biology.
+In the following, we shortly describe some of the kernels in each category.
+
+%Examples are kernels proposed in \cite{shervashidze2011weisfeiler, gartner2003survey, borgwardt2005shortest, costa2010fast}.
 Following we describe NSPDK \cite{costa2010fast}, a convolutional graph kernel which shows state of performance and it is later on used to define our proposed graph node kernel.
+\subsubsection{Graph kernels}
+\label{graph-kernels}
+The task of designing effient and expressive graph kernels play an important role in the development of graph-based predictive systems.
+Existing graph kernels are decompositional kernels and can be classified into two categories: sequence-based graph kernels and subgraph-based graph kernels.
+The sequence-based graph kernels decompose graphs into ``parts" in sequence-based forms, such as paths and walks.
+Typical examples of sequence-based graph kernels are product graph kernel, \cite{gartner2003survey}, shortest path kernels \cite{borgwardt2005shortest}.
+The subgraph-based graph kernels dissolve graphs into subgraphs.
+Examples include Weisfeiler-Lehman kernels \cite{shervashidze2009fast, shervashidze2011weisfeiler}, The Neighborhood Subgraph Pairwise Distance Kernel (NSPDK) \cite{costa2010fast}.
+Following we describe NSPDK since it is later adopted to develop our graph node kernel (presented in \ref{method}).
 
-\subsubsection*{The Neighborhood Subgraph Pairwise Distance Kernel}
-\label{sec:ndpdk}
 The NSPDK \cite{costa2010fast} is an instance of convolution kernel \cite{haussler1999convolution} where given a graph $G \in \mathcal{G}$ and two rooted graphs $A_u, B_v$, the relation $R_{r,d}(A_u, B_v, G)$ is true {\em iff} $A_u \cong \mathcal{N}_r^u$ is (up to isomorphism $\cong$) a neighborhood subgraph of radius $r$ of $G$ and so is $B_v \cong  \mathcal{N}_r^v$, with roots at distance $\mathcal{D}(u,v)= d$.
@@
 
+\subsubsection{Graph node kernels}
 The state of the art kernels are based on diffusion phenomenon.
@@
 
-\subsubsection{Laplacian exponential diffusion kernel}
-One of the most well-known graph node kernels is the Laplacian exponential diffusion kernel \textbf{LEDK}, as it is widely used for exploiting discrete structures in general and graphs in particular.
-On the basis of the heat diffusion dynamics, Kondor and Lafferty proposed \textbf{LEDK} in \cite{kondor2002diffusion}: imagine to initialize each vertex with a given amount of heat and let it flow through the edges until an arbitrary instant of time.
+\begin{itemize}
+
+\item Laplacian exponential diffusion kernel: One of the most well-known graph node kernels is the Laplacian exponential diffusion kernel LEDK, as it is widely used for exploiting discrete structures in general and graphs in particular.
+On the basis of the heat diffusion dynamics, Kondor and Lafferty proposed LEDK in \cite{kondor2002diffusion}: imagine to initialize each vertex with a given amount of heat and let it flow through the edges until an arbitrary instant of time.
 The similarity between any vertex couple $v_{i}$, $v_{j}$ is the amount of heat starting from $v_{i}$ and reaching $v_{j}$ within the given time.
-Therefore, \textbf{LEDK} can capture the long range relationship between vertices of a graph to define the global similarities.
-Below is the formula to compute \textbf{LEDK} values:
+Therefore, LEDK can capture the long range relationship between vertices of a graph to define the global similarities.
+Below is the formula to compute LEDK values:
 \begin{equation} 
 \label{LEDK-formula}
-K = e^{-\beta \textbf{L}} = \textbf{I} - \beta \textbf{L} + \frac{\beta \textbf{L}^{2}}{2!} - ...
+K = e^{-\beta L} = I - \beta L + \frac{\beta L^{2}}{2!} - ...
 \end{equation}
-where $\beta$ is the diffusion parameter and is used to control the rate of diffusion and $\textbf{I}$ is the identity matrix.
+where $\beta$ is the diffusion parameter and is used to control the rate of diffusion and $I$ is the identity matrix.
 Choosing a consistent value for $\beta$ is very important: on the one side, if $\beta$ is too small, the local information cannot be diffused effectively and, on the other side, if it is too large, the local information will be lost.
-\textbf{LEDK} is positive semi-definite as proved in \cite{kondor2002diffusion}.
-\subsubsection{Exponential diffusion kernel}
-In \textbf{LEDK}, the similarity values between high degree vertices are generally higher compared to those between low degree ones.
+LEDK is positive semi-definite as proved in \cite{kondor2002diffusion}.
+
+\item Exponential diffusion kernel: In LEDK, the similarity values between high degree vertices are generally higher compared to those between low degree ones.
 Intuitively, the more paths connect two vertices, the more heat can flow between them.
 This could be problematic since peripheral nodes have unbalanced similarities with respect to central nodes.
-In order to make the strength of individual vertices comparable, a modified version of \textbf{LEDK} is introduced by Chen et al in \cite{chen2014disease}, called Markov exponential diffusion kernel \textbf{MEDK} and given by the following formula:
+In order to make the strength of individual vertices comparable, a modified version of LEDK is introduced by Chen et al in \cite{chen2014disease}, called Markov exponential diffusion kernel MEDK and given by the following formula:
 \begin{equation} \label{MEDK-formula}
-\textbf{K} = e^{-\beta \textbf{M}}
+K = e^{-\beta M}
 \end{equation}
-The difference with respect to the Laplacian diffusion kernel is the replacement of \textbf{L} by the matrix $\textbf{M}=(\textbf{D}-\textbf{A}-n\textbf{I})/n$ where $n$ is the total number of vertices in graph.
-The role of $\beta$ is the same as for \textbf{LEDK}.
-\subsubsection{Markov diffusion kernel}
-The original Markov diffusion kernel \textbf{MDK} was introduced by Fouss et al.
+The difference with respect to the Laplacian diffusion kernel is the replacement of $L$ by the matrix $M=(D-A-nI)/n$ where $n$ is the total number of vertices in graph.
+The role of $\beta$ is the same as for LEDK.
+
+\item Markov diffusion kernel: The original Markov diffusion kernel MDK was introduced by Fouss et al.
 \cite{fouss2006experimental} and exploits the idea of diffusion distance, which is a measure of how similar the pattern of heat diffusion is among a pair of initialized nodes.
@@
 If their diffusion ways are alike, the similarity will be high and, vice versa, it will be low if they diffuse differently.
-This kernel is computed starting from the transition matrix \textbf{P} and by defining $\textbf{Z}(t) = \frac{1}{t}\sum_{\tau=1}^{t}\textbf{P}^{\tau}$, as follows:
+This kernel is computed starting from the transition matrix $P$ and by defining $Z(t) = \frac{1}{t}\sum_{\tau=1}^{t}P^{\tau}$, as follows:
 \begin{equation} 
 \label{MDK-formula}
-\textbf{K} = \textbf{Z}(t)\textbf{Z}^{\top}(t)
+K = Z(t)Z^{\top}(t)
 \end{equation}
 
-\subsubsection{Regularized Laplacian kernel}
-Another popular graph node kernel function used in graph mining is the regularized Laplacian kernel \textbf{RLK}.
+\item Regularized Laplacian kernel: Another popular graph node kernel function used in graph mining is the regularized Laplacian kernel RLK.
 This kernel function is introduced by Chebotarev and Shamis in \cite{chebotarev2006matrix} and represents a normalized version of the random walk with restart model.
@@
 \label{RLK-formula}
-\textbf{K} = \sum_{n=0}^{\infty}\beta^{n}(-\textbf{L})^n = (\textbf{I} + \beta \textbf{L})^{-1}
+K = \sum_{n=0}^{\infty}\beta^{n}(-L)^n = (I + \beta L)^{-1}
 \end{equation}
 where the parameter $\beta$ is again the diffusion parameter.
-\textbf{RLK} counts the paths connecting two nodes on the graph induced by taking \textbf{-L} as the adjacency matrix, regardless of the path length.
+RLK counts the paths connecting two nodes on the graph induced by taking $-L$ as the adjacency matrix, regardless of the path length.
 Thus, a non-zero value is assigned to any couple of nodes as long as they are connected by any indirect path.
-\textbf{RLK} remains a relatedness measure even when diffusion factor is large, by virtue of the negative weights assigned to self-loops.
+RLK remains a relatedness measure even when diffusion factor is large, by virtue of the negative weights assigned to self-loops.
 
+\end{itemize}
 \section{Method}
+\label{method}
 We start from the type of similarity notion computed by a neighborhood based decomposition kernel between graph instances \cite{costa2010fast}, NSPDK, and adapt it to form a convolutional graph node kernel which aims at expressing the similarity between nodes in a single graph.
@@
 
-Given an input graph $G=(V,E)$, our method consists of three steps.
-In the first step, \textit{node labeling}, a labeling function is proposed to assign label for every node of the graph.
-This step is necessary since we desire to design a convolutional kernel which deals with labeled graphs.
-The second step includes a graph decompostion procedure which allows to decompose the graph into a set of linked sparse connected components.
-In this procedure, we define two different kinds of link: \textit{conjunctive} and \textit{disjunctive} in which they are treated in distinct manners later on.
-In the final step, the similarity between any node couple $u$ and $v$ is computed by adopting NSPDK on two neighborhood subgraphs rooted as $u$ and $v$ which extracted from the resulted graph after decomposing.
-Following, we describe each step in detail.
-
-\subsection{Node Labeling} 
-\label{sec:lab}
-We are interested in investigating biological networks in general and genetic networks in specific.
-Therefore, in this section, we propose two different approaches to label for genetic graphs.
-It is worth to notice that genetic networks typically represent genes as nodes labeled with a gene identifier.
-Here we take a different approach and use the node labels to encode abstract information about the genes.
-In this way downstream machine learning algorithms can generalize from similar examples and allow the identification of overlooked but related genes.
-We experiment with two types of information: \textit{i}) topological information and \textit{ii}) functional information.\\
-
-\textit{Topological labels} This information is simply based on the connectivity degree of the gene.
-The idea is that genes that have the same number of connections are more similar than genes with a different connectivity.
-Here we propose a node labeling function $\mathcal{L}$ that assigns labels for nodes by considering their degrees, such that nodes with similar degrees are assigned with a same label.
-Therefore, our function assigns the degree for nodes $u$ having degree less than or equal to a user defined threshold $T$ ($T=5$ in our experimental evaluation).
-However degree values larger than $T$ are subsequently discretized into $k$ levels.
-We treat nodes differently between nodes with high degrees and low degrees since we envisage that the properties of low-degree nodes whose different node degrees are more dissimilar in comparison with the ones of high degrees.
-Formally, the labeling function is defined as:
-\begin{center}
-$\mathcal{L}(u) = \left\{
-	\begin{array}{ll}
-		deg(u),\  & \mbox{if } deg(u) \leq T \\
-		T+i,\ & \mbox{if } deg(u) > T
-	\end{array}
-\right.$,
-\end{center}
-where $i = \lceil \frac{deg(u)-T}{bin}\rceil$, $bin = \frac{deg(G)-T}{p - T}$ and $p$ ($p > T$) is the maximum number of symbols used.
-The value of $p$ depends on the degree distribution and can be tuned as a hyperparameter of the approach.\\
-
-It is often that nodes in the graph are associated to extra information.
-These information can be considered as the complement to the graph topological information.
-Therefore, utilizing both information source are supposed to improve the performance of graph node kernels.
-In the following, we propose two different approaches to process with this kind of information.
-In the first approach, it is converted and used as the functional discrete labels for nodes, while it is processed and utilized as the real valued vectors in the second one.
-In both cases, we use \textit{Gene Ontology} \cite{gene2004gene} as an example of extra information associated to the graph nodes.\\
-
-\textit{Functional discrete labels}
-We use a gene ontology to construct binary vectors representing a bag-of-words encoding for each gene (\textit{i.e.
-each element of a binary vector is equal to 1 if its corresponding GO-term is associated to the gene, and is equal to 0 otherwise.}).
-The resulting vectors are then clustered using the k-means algorithm  into a user defined number of classes $K$ (tuned as a hyperparameter of the approach), so that genes with similar description profiles receive the same class identifier as label.\\
-
-\textit{Functional real valued vector labels}
-In addition to encoding the functional information as a discrete label we add a richer description by computing the similarity vector w.r.t.
-to each cluster.
-In this way we can fully exploit the latent description of the genes in terms of the different functional groups captured by the clustering procedure.
-Formally, given a vector $v \in I\!R^{26501}$ we compute a similarity vector $S(v)= {s_1, s_2, \ldots s_K}$ with entries $s_i = \frac{1}{1+ \ell(v,c_i)}$ where $\ell(v,c_i)$ is the Euclidean distance of $v$ from the center of the $i^{th}$ cluster $c_i = \frac{1}{|C_i|}\sum_{x \in C_i} x$ computed as the geometric mean of the elements in the cluster $C_i$.
+Given an input labeled graph $G=(V,E, \mathcal{L}_1, \mathcal{L}_2)$, our kernel consists of two phases.
+In the first phase, a network decomposition procedure is applied to transform the graph into a set of linked sparse connected components.
+In this procedure, we define two different kinds of link: \textit{conjunctive} and \textit{disjunctive} in which we treat them in distinct manners.
+In the second step, the similarity between any node couple ($u$, $v$) is computed by adopting NSPDK on two neighborhood subgraphs rooted as $u$ and $v$ which extracted from the resulted graph after decomposing.
+In the following, we describe each phase in detail.
 
 \subsection{Network Decomposition} 
-In genetic networks it is not uncommon to find nodes with high degrees.
-Unfortunately these cases cannot be effectively processed by a neighborhood based decomposition kernel (see convolution kernels presented in \ref{convolution-kernels}) since these are based on the notion of exact matches and the probability of finding identical neighborhoods decreases exponentially as the degree increases.
+\label{network-decomposition}
+In genetic networks, it is not uncommon to find nodes with high degrees.
+Unfortunately these cases cannot be effectively processed by a neighborhood based decomposition kernel (see \ref{graph-kernels}) since these are based on the notion of exact matches.
+Neighborhood subgraphs rooted at high degree nodes are relatively big due to a high number of neighbors and edges.
+It leads to a low likelihood of finding neighborhood subgraphs rooted at high degree nodes which are isomorphic.
+Thus, the probability of having identical neighborhoods decreases exponentially as the degree increases.
 This means that in a finite network it quickly becomes impossible to find any match and hence learn or generalize at all.
-Therefore, we propose a procedure to ``sparsify'' the network that is observed by the neighborhood kernel.
+
+We propose a procedure to ``sparsify'' the network that is observed by the neighborhood kernel.
 In practice, we mostly keep the same the cardinality of the edge set.
@@
 
-\textit{Iterative k-core decomposition} \cite{alvarez2005k}: The node set is partitioned in two groups on the basis of the degree of each node w.r.t.
+\textit{Iterative k-core decomposition} \cite{alvarez2005k}: the node set is partitioned in two groups on the basis of the degree of each node w.r.t.
 a threshold degree $D$, the first part contains all nodes with degree smaller than or equal to $D$ and the second part the remaining ones.
@@
 
-\textit{Clique decomposition} \cite{tarjan1985decomposition}: To model the notion that nodes in a clique are tightly related, we summarize the whole clique with a new ``representative'' node.
+\textit{Clique decomposition} \cite{tarjan1985decomposition}: to model the notion that nodes in a clique are tightly related, we summarize the whole clique with a new ``representative'' node.
 All the cliques (completely connected subgraphs) with a number of nodes greater than or equal to a given threshold size $C$ are identified.
@@
 \centering
-\includegraphics[width=.9\textwidth]{clique}
+\includegraphics[width=.9\textwidth]{cliques}
 \caption{Clique decomposition with threshold $C = 4$}
@@
 \subsection{The Conjunctive Disjunctive Node Kernel}
-We start from the Neighborhood Subgraph Pairwise Distance Kernel (NSPDK) \cite{costa2010fast} and adapt it to express the similarity between nodes in a single network.
+%We start from the Neighborhood Subgraph Pairwise Distance Kernel (NSPDK) \cite{costa2010fast} and adapt it to express the similarity between nodes in a single network.
 The key idea in NSPDK is to decompose graphs in small fragments and count how many pairs of fragments are shared between two instances.
@@
 \includegraphics[width=.9\textwidth]{cdnk}
-\caption{Pairwise neighborhood subgraphs for the ``red'' node with $r=1$ and $d=3$ using ``conjuction'' and ``disjuctive'' edges}
+\caption{Pairwise neighborhood subgraphs for the ``red'' node with $r=1$ and $d=3$ using ``\textit{conjuction}'' and ``\textit{disjuctive}'' edges}
 \label{fig:cdnk}
@@
 
-\subsection{Real valued node information}
-In order to integrate the information of real vectors we proceed as follows.
+In order to integrate the information of real vector labels, we proceed as follows.
 We compute a sparse vector representation for the neighborhood graph rooted in node $v$ following \cite{costa2010fast}: for each neighborhood subgraph we calculate the quasi-isomorphism certificate hash code; we then combine the hashes for the pair of neighborhoods and use the resulting integer as a feature indicator.
-This yields a direct sparse vector representation (associated to node $u$ in graph $G$) $f: G_u \longmapsto I\!R^L$ where $L \approx 10K-1M$.
-Given the real valued vector information (associated to node $u$ in graph $G$) $g: G_u \longmapsto I\!R^K$ computed as the multi-class similarity to the $K$ clusters (c.f.r.
+This yields a direct sparse vector representation (associated to node $u$ in graph $G$) $f: G_u \longmapsto I\!R^N$ where $N \approx 10K-1M$.
+Given the real valued vector information (associated to node $u$ in graph $G$) $g: G_u \longmapsto I\!R^P$ computed as the multi-class similarity to the $P$ clusters (c.f.r.
 Section \ref{sec:lab}), we update the computation of CDNK considering the discrete convolution of the discrete information with the real valued information:
@@
 
+\section{Parameter Space}
+Our kernel consists of five parameters: the threshold degree $D$, clique size threshold $C$, maximal radius $r^*$, maximal distance $d^*$ and number of clusters $P$ clusters in which their values are in $\mathbb{N}^*$.
+In order to choose the optimal tuple of parameters for kernel in a specific setting, a model selection procedure is normally adopted from a determined subset of parameter space.
+The parameter space of our kernel seems large due to the relatively high number of parameters.
+However, most parameter values are supposed to be in limited ranges.
+Therefore, it is actually not too big.
+In particular, the clique size threshold $C$ is recommended to be in $\lbrace 4, 5\rbrace$.
+The maximal radius $r^*$ is set with a value smaller than or equal to $3$ since it will lead to big neighborhood subgraphs (as discussed in \ref{network-decomposition}).
+The maximal distance $d^*$ is assigned with values less than or equal to $4$ because the value of the maximal shortest distance between nodes in a connected component is often not too high.
+The values of the threshold degree $D$ should neither be too high and too low.
+If it is high, we have to face with high degree node problem, and if it is too low, the obtained graph consists of too sparse connected components.
+Therefore, we suggest the value for $D$ is in $\left[6, 20 \right]$.
+
 \section{Empirical Evaluation}
-In this section, we desire to evaluate the performance of CDNK and compare it with the state of the art graph node kernels.
-In order to do that, we employ disease gene prioritization problem where the aim is to build a learning system which allows to prioritize candidate genes based on their probabilities of being associated to a given genetic disease.
+\label{empirical-evaluation-cdnk}
+In this section, we desire to evaluate the performance of CDNK and other graph node kernels to answer for the two following questions:
+\begin{itemize}
+\item Does CDNK show better performance comparing to other graph node kernels?
+\item Does the use of side information (real vector labels) help to improve the performance of CDNK?
+\end{itemize}
 
-Given a genetic graph and a list of training genes for a genetic disease (training set consists both positive and negative genes), we first apply  a graph node kernel to compute kernel matrix.
-This kernel matrix together with training gene set are used as input of a learning algorithm to construct a system.
-The trained system is used to prioritize for candidate genes.
+\subsection{Experimental Settings and Evalution Method}
+We carry out experiments in the context of disease gene prioritization where the aim is to build a learning system which allows to prioritize candidate genes based on their probabilities of being associated to a given genetic disease.
 
-To compare the performance of graph node kernels, we fix the learning algorithm and substitute graph node kernels.
-The performance of the system obtained in different cases reflect the performance of graph node kernels.
+ Given a genetic graph and a list of training genes known to cause a genetic disease.
+We first apply  a graph node kernel to compute kernel matrix.
+This kernel matrix together with training gene set are used as the input of a learning algorithm to construct a model.
+The obtained model is then used to prioritize for candidate genes.
 
-Datasets: We carry out in two separate networks derived from BioGPS and Pathways datasets. Following are the short description of these networks.
-We carry out experiments on two separate datasets and we follow experiment's procedure carried out in \cite{chen2014disease} where 
-12 diseases \cite{goh2007human} are used in each disease is associated to at least 30 positive genes.
+The experiments are performed on two separate networks derived from BioGPS and Pathways datasets in which we follow the experiment procedure in \cite{chen2014disease} where 12 diseases \cite{goh2007human} are used in which each disease is associated to at least 30 positive genes (see table \ref{tab:cdnk-diseases} for the list of genetic diseases and the number of positive genes in each disease).
 For each disease, we construct a positive set $\mathcal{P}$ with all positive (confirmed) disease genes, and a negative set $\mathcal{N}$ which contains random genes associated at least to one disease class which is not related to the class that is defining the positive set.
 In \cite{chen2014disease} the ratio between the dataset sizes is chosen as $\vert \mathcal{N} \vert = \frac{1}{2} \vert \mathcal{P} \vert$.
-The predictive performance of each method is evaluated via a leave-one-out cross validation: one gene is kept out in turn and the rest are used to train an SVM model.
-We compute a decision score $q_i$ for the test gene $g_i$ as the top percentage value of score $s_i$ among all candidate gene scores.
-We collect all decision scores for every gene in the training set to form a global decision score list on which we compute the AUC-ROC.
-
+This is due to the fact that genes known to be related to at least a genetic disease, but not to the considered one are well studied, so they have low probability to be associated to the current disease.
 \begin{itemize}
@@
 \end{itemize}
+To compare the performance of graph node kernels, we fix the learning algorithm and sequentially substitute differnt graph node kernels.
+The performance of the system obtained in different cases are used to compare the performance of the employed graph node kernels: LEDK, MEDK, MDK, RLK and CDNK.
 
-\textbf{Model Selection:}
-The hyper parameters of the various methods are tuned using a k-fold strategy.
+The predictive performance of the system using each kernel is evaluated via a leave-one-out cross validation: one gene is kept out in turn and the rest are used to train an SVM model.
+We compute a decision score $q_i$ for the test gene $g_i$ as the top percentage value of score $s_i$ among all candidate gene scores.
+We collect all decision scores for every gene in the test set to form a global decision score list on which we compute the AUC-ROC.
+
+\subsection{Node Labeling} 
+\label{sec:lab}
+The graph node kernel takes the labled graphs as its input in which labels can be discrete and/or real valued vectors.
+Therefore, to carry out experiments on graphs derived from BioGPS and Pathways, we need ways to label for these genetic graphs.
+Following are our proposed node labeling methods for genetic networks.
+Note that we allow the user to define and use any node labeling functions which are informative.
+
+\textit{Discrete labels:} We cast two different approaches to associate genes with dicrete labels.
+
+The first approach, we use a same label for every node of graphs.
+In this case, the configuration of nodes expressed by the pairwise neighborhood subgraphs in CDNK now turn to pairwise neighborhoods.
+Besides, since labels are identical, the cardinality of neighborhoods are took into account.
+
+The second approach aims to use nodes labels to encode abstract information about the genes.
+In this way downstream machine learning algorithms can generalize from similar examples and allow the identification of overlooked but related genes.
+We employ a gene ontology \cite{gene2004gene} to construct binary vectors representing a bag-of-words encoding for each gene, i.e.
+each element of a binary vector is equal to 1 if its corresponding GO-term is associated to the gene, and is equal to 0 otherwise.).
+The resulting vectors are then clustered using the k-means algorithm into a user defined number of classes, $P$, so that genes with similar description profiles receive the same class identifier as label.
+
+\textit{Real vector labels:} In addition to encoding the functional information as a discrete label we add a richer description by computing the similarity vector w.r.t.
+to each cluster.
+In this way we can fully exploit the latent description of the genes in terms of the different functional groups captured by the clustering procedure.
+Formally, given a vector $v \in I\!R^{26501}$ we compute a similarity vector $S(v)= {s_1, s_2, \ldots s_P}$ with entries $s_i = \frac{1}{1+ \ell(v,c_i)}$ where $\ell(v,c_i)$ is the Euclidean distance of $v$ from the center of the $i^{th}$ cluster $c_i = \frac{1}{|C_i|}\sum_{x \in C_i} x$ computed as the geometric mean of the elements in the cluster $C_i$.
+
+\subsection{Model Selection}
+The hyper parameters of the various methods are tuned using a k-fold cross validation.
 However due to the non i.i.d.
 nature of the problem, we employ a stronger setup to ensure no information leakage.
-The dataset on which we are validating the performance is never subsequently used in the predictive performance estimate.
-The values for diffusion parameter in DK and MED are sampled in $\lbrace 10^{-3}, 10^{-3}, 10^{-2}, 10^{-1} \rbrace$, time steps in MD in $\lbrace 1, 10, 100 \rbrace$ and RL parameter in $\lbrace 1, 4, 7 \rbrace$.
-For CDNK, the degree threshold values are sampled in $\lbrace 10,\ 15,\ 20 \rbrace$, clique size threshold in $\lbrace 4,\ 5 \rbrace$, maximum radius in $\lbrace 1,\ 2 \rbrace$, maximum distance in $\lbrace 2,\ 3,\ 4 \rbrace$, number of clusters $K$ in $\lbrace 5,\ 7 \rbrace$.
+The dataset on which we are validating the performance is never subsequently used in the predictive performance estimation.
+The values for diffusion parameter $\beta$ in LEDK and MEDK are sampled in $\lbrace 10^{-3}, 10^{-3}, 10^{-2}, 10^{-1} \rbrace$, time steps $t$ in MDK in $\lbrace 1, 10, 100 \rbrace$ and RLK parameter $\beta$ in $\lbrace 1, 4, 7 \rbrace$.
+For CDNK, the degree threshold values, $D$, are sampled in $\lbrace 10,\ 15,\ 20 \rbrace$, clique size threshold, $C$, in $\lbrace 4,\ 5 \rbrace$, maximum radius, $r^*$, in $\lbrace 1,\ 2 \rbrace$, maximum distance, $d^*$, in $\lbrace 2,\ 3,\ 4 \rbrace$, number of clusters $P$ in $\lbrace 5,\ 7 \rbrace$.
 Finally, the regularization trade off parameter $C$ for the SVM is sampled in $\lbrace 10^{-5},  \ 10^{-4}, \ 10^{-3},\ 10^{-2},\ 10^{-1}, 1,\ 10,\ 10^2 \rbrace$.
 
-Table \ref{table_biogps} and \ref{table_pathways} show the AUC performance of the system corresponding to different cases of employing graph node kernels.
-Predictive models based on CDNK variations are ranked in top positions in all cases on BioGPS dataset and 9 out of 11 cases on Pathways when compared to diffusion based kernels.
-CDNK is ranking first when considering both the average AUC-ROC and the average rank with a difference, w.r.t.
-the best diffusion kernel, ranging from 5.4$\%$ to 10$\%$ and from 1.2$\%$ to 3.2$\%$ on BioGPS and Pathways, respectively.
-Regarding the variations of CDNK, the integration of real valued vector information improves the performance in most cases on using discrete labels only.
-The improvement in average ranges from 1.3$\%$ to 4.6$\%$.
-Note that also in the case where only topological information is used to induce the discrete label (CDNK3 in the Table \ref{table_biogps} and \ref{table_pathways}), the proposed approach improves on state-of-the-art.
+\subsection{Results and Discussion}
+Table \ref{fig:cdnk-biogps} and \ref{fig:cdnk-pathways} show the AUC-ROC performance of different models using state of the art graph node kernels and different variations of CDNK on BioGPS and Pathways networks, respectively.
+In what follows, we analyze the results to answer for the two questions which are previously raised.
 
+Concering the performance of different graph node kernels, CDNK variations outperform diffusion-based graph node kernels in all cases on BioGPS dataset and 8 out of 11 cases on Pathways when compared to diffusion-based kernels.
+CDNK is ranked first when considering both the average AUC-ROC and the average rank with a difference, w.r.t.
+the best diffusion-based kernel, ranging from 5.4$\%$ to 10$\%$ and from 1.2$\%$ to 3.4$\%$ on BioGPS and Pathways, respectively.
+As a consequence, we can conclude that CDNK show state of the art performance in graph node kernels.
+
+Regarding the variations of CDNK, the integration of real valued vectors improves the performance in most cases.
+In particular, considering the use of discrete labels based on ontology, using side information helps to increase the performance in all diseases on BioGPS and in 9 out of 11 diseases on Pathways.
+Similarly, by employing auxiliary information associated to graph nodes, the performance of CNDK in case of using uniform discrete labels is also improved in 9 and 8 out of 11 on BioGPS and Pathways, respectively.
+%==================================================
+\section{Conclusion and Future Work}
+We have shown how decomposing a network in a set of connected sparse graphs allows us to take advantage of the discriminative power of CDNK, a novel decomposition kernel, to achieve state-of-the-art results.
+Moreover, we have also introduced the way to integrate ``side'' information in form of real valued vectors when it is available on graph to get even better performance of CDNK.
+In future work we will investigate how to \textit{i}) decompose networks in a data driven way and \textit{ii}) extend the CDNK approach to gene-disease association problems exploiting multiple heterogeneous information sources in a joint way.
+%======================================================
+%==================================================
+%==================================================
 \begin{table}
@@
 We report the AUC ROC and the rank for each kernel method.
-CDNK1 = ontology for discrete labels, CDNK2 = ontology for both discrete and vector labels, CDNK3 = node degree for discrete labels and CDNK4 = node degree for discrete labels and ontology for vector label.}}
-\vspace*{5pt}
+CDNK1 = ontology for discrete labels, CDNK2 = ontology for both discrete and vector labels, CDNK3 = uniform discrete labels and CDNK4 = uniform discrete labels and ontology for vector label.}} \vspace{1em}
+%\vspace*{5pt}
 \centering
@@
  \hline
-1 & 51.9/8 & 57.4/7 & 59.0/6 & 59.2/5 & 65.1/4 & 69.5/2 & 69.3/3 & \textbf{70.3/1} \\
-2 & 81.7/5 & 78.5/6 & 75.2/7 & 75.0/8 & 88.3/2 & \textbf{88.8/1} & 85.1/4 & 86.8/3 \\
-3 & 64.3/7 & 59.6/8 & 71.6/3 & 71.8/2 & 65.5/5 & \textbf{72.5/1} & 64.7/6 & 66.4/4 \\
-4 & 65.3/7 & 58.2/8 & 67.8/6 & 67.8/5 & 71.9/4 & \textbf{78.7/1} & 73.9/3 & 77.0/2 \\
-5 & 64.0/8 & 64.1/7 & 66.5/5 & 66.2/6 & 75.9/4 & 76.2/3 & 76.9/2 & \textbf{77.4/1} \\
-6 & 74.6/5 & 70.2/8 & 71.0/7 & 71.2/6 & 79.3/2 & \textbf{83.7/1} & 76.7/4 & 79.0/3 \\
-7 & 73.0/5 & 66.7/8 & 75.4/3 & 75.6/2 & 68.8/6 & 73.9/4 & 67.3/7 & \textbf{76.9/1} \\
-8 & 74.4/8 & 76.8/3 & 76.2/5 & 76.4/4 & 74.7/7 & \textbf{77.7/1} & 76.0/6 & 76.8/2 \\
-9 & 71.5/2 & 65.6/8 & 67.7/5 & 69.9/3 & 66.8/7 & \textbf{71.7/1} & 67.1/6 & 68.1/4 \\
-10 & 54.0/6 & 50.3/8 & 56.1/5 & 51.1/7 & 77.6/4 & \textbf{82.7/1} & 80.5/2 & 80.0/3 \\
-11 & 58.2/7 & 51.3/8 & 59.3/6 & 59.3/5 & 71.8/4 & \textbf{80.2/1} & 75.3/3 & 77.1/2 \\
+1 & 51.9/8 & 57.4/7 & 59.0/6 & 59.2/5 & 65.1/4 & 69.5/3 & 72.0/2 & \textbf{72.3/1} \\
+
+2 & 81.7/5 & 78.5/6 & 75.2/7 & 75.0/8 & 88.3/1 & \textbf{88.8/1} & 85.0/3 & 84.1/4 \\
+
+3 & 64.3/6 & 59.6/8 & 71.6/3 & 71.8/2 & 65.5/5 & \textbf{72.5/1} & 61.4/7 & 67.9/4 \\
+
+4 & 65.3/7 & 58.2/8 & 67.8/6 & 67.8/5 & 71.9/4 & \textbf{78.7/1} & 73.8/3 & 76.7/2 \\
+
+5 & 64.0/8 & 64.1/7 & 66.5/5 & 66.2/6 & 75.9/4 & 76.2/3 & 77.8/2 & \textbf{78.2/1} \\
+
+6 & 74.6/5 & 70.2/8 & 71.0/7 & 71.2/6 & 79.3/3 & \textbf{83.7/1} & 77.2/4 & 80.0/2 \\
+
+7 & 73.0/5 & 66.7/7 & 75.4/3 & 75.6/2 & 68.8/6 & 73.9/4 & 66.1/8 & \textbf{77.4/1} \\
+
+8 & 74.4/8 & 76.8/3 & 76.2/5 & 76.4/4 & 74.7/7 & 77.7/1 & 76.1/6 & 76.9/2 \\
+
+9 & 71.5/2 & 65.6/8 & 67.7/6 & 69.9/3 & 66.8/7 & \textbf{71.7/1} & 67.9/5 & 68.1/4 \\
+
+10 & 54.0/6 & 50.3/8 & 56.1/5 & 51.1/7 & 77.6/4 & \textbf{82.7/1} & 77.7/3 & 78.3/2 \\
+
+11 & 58.2/7 & 51.3/8 & 59.3/6 & 59.3/5 & 71.8/4 & \textbf{80.2/1} & 74.8/2 & 74.0/3 \\
+
 \hline
-$\overline{AUC}$ & 66.6 & 63.5 & 67.8 & 67.6 & 73.2 & \textbf{77.8} & 73.9 & 76.0 \\
-$\overline{Rank}$ & 	6.18 & 7.18 & 5.27 & 4.82 & 4.45 & \textbf{1.55} & 4.18 & 2.36 \\
+$\overline{AUC}$ & 66.6	& 63.5 & 67.8 &	67.6 & 73.2 & \textbf{77.8} & 73.6 & 75.8 \\
+
+$\overline{Rank}$ & 6.09 & 7.09 &  5.36 & 4.82 & 4.45 & \textbf{1.64} & 4.09 & 2.36 \\
  \hline 
 \end{tabular}
-\label{table_biogps}
+\label{fig:cdnk-biogps}
 \end{table}
-%==================================================
+
+%=====================================
+%=====================================
 \begin{table}
@@
  \hline
-1 & 74.7/8 & 76.4/7 & 78.7/6 & 78.8/5 & 80.2/3 & \textbf{82.4/1} & 80.6/2 & 79.4/4 \\
-2 & 55.1/8 & 64.9/7 & 76.6/6 & 76.6/5 & 81.1/2 & 80.3/3 & 79.8/4 & \textbf{82.2/1} \\
-3 & 55.0/8 & 62.7/7 & 64.1/5 & 65.6/4 & 67.1/3 & 63.6/6 & 68.5/2 & \textbf{71.5/1} \\
-4 & 54.3/8 & 65.2/7 & \textbf{73.7/1} & 73.7/2 & 66.1/5 & 68.1/3 & 67.5/4 & 65.6/6 \\
-5 & 52.9/8 & 55.7/7 & 62.7/5 & 62.7/6 & 68.3/2 & \textbf{69.6/1} & 66.3/4 & 66.8/3 \\
-6 & 83.4/8 & 92.7/7 & 96.5/2 & \textbf{96.5/1} & 93.0/6 & 94.1/5 & 94.4/4 & 95.5/3 \\
-7 & 84.5/8 & 88.3/7 & 89.4/3 & 89.5/2 & 88.5/6 & 88.5/5 & 88.7/4 & \textbf{90.5/1} \\
-8 & 53.7/8 & 65.6/7 & 72.0/6 & 72.3/5 & 72.5/3 & 72.5/2 & 72.3/4 & \textbf{76.5/1} \\
-9 & 52.5/8 & 64.9/5 & 64.2/7 & 64.2/6 & \textbf{81.3/1} & 81.0/2 & 79.6/3 & 78.8/4 \\
-10 & 68.8/6 & 65.4/8 & 74.4/5 & 74.4/4 & 66.9/7 & 76.8/2 & 76.1/3 & \textbf{79.5/1} \\
-11 & 53.7/8 & 69.2/7 & 74.6/5 & 74.1/6 & 77.0/2 & \textbf{78.7/1} & 75.4/4 & 77.0/3 \\
+1 & 74.7/8 & 76.4/7 & 78.7/6 & 78.8/5 & 80.2/4 & 82.4/2 & \textbf{82.8/1} & 82.3/3 \\
+
+2 & 55.1/8 & 64.9/7 & 76.6/6 & 76.6/5 & 81.1/2 & 80.3/4 & 80.1/3 & \textbf{81.9/1} \\
+
+3 & 55.0/8 & 62.7/7 & 64.1/5 & 65.6/4 & 67.1/3 & 63.6/6 & 69.7/2 & \textbf{71.4/1} \\
+
+4 & 54.3/8 & 65.2/6 & \textbf{73.7/1} & 73.7/2 & 66.1/5 & 68.1/3 & 64.0/7 & 67.3/4 \\
+
+5 & 52.9/8 & 55.7/7 & 62.7/5 & 62.7/6 & 68.3/2 & \textbf{69.6/1} & 66.2/4 & 68.1/3 \\
+
+6 & 83.4/8 & 92.7/7 & 96.5/2 & \textbf{96.5/1} & 93.0/6 & 94.1/5 & 94.5/4 & 94.9/3 \\
+
+7 & 84.5/7 & 88.3/8 & 89.4/2 & \textbf{89.5/1} & 88.5/6 & 88.5/5 & 89.0/3 & 88.7/4 \\
+
+8 & 53.7/8 & 65.6/7 & 72.0/6 & 72.3/5 & 72.5/4 & 72.5/3 & 74.9/2 & \textbf{75.3/1} \\
+
+9 & 52.5/8 & 64.9/5 & 64.2/7 & 64.2/6 & \textbf{81.3/1} & 81.0/3 & 80.1/2 & 79.8/4 \\
+
+10 & 68.8/6 & 65.4/8 & 74.4/5 & 74.4/4 & 66.9/7 & 76.8/2 & 75.2/3 & \textbf{78.7/1} \\
+
+11 & 53.7/8 & 69.2/7 & 74.6/5 & 74.1/6 & 77.0/3 & \textbf{78.7/1} & 75.1/4 & 77.3/2 \\
+
 \hline
-$\overline{AUC}$ & 62.6 & 70.1 & 75.2 & 75.3 & 76.5 & 77.8 & 77.2 & \textbf{78.5} \\
-$\overline{Rank}$ & 7.82 & 6.91 & 4.64 & 4.18 & 3.64 & 2.82 & 3.45 & \textbf{2.55} \\
+$\overline{AUC}$ & 62.6	& 70.1 & 75.2 & 75.3 & 76.5 & 77.8 & 77.4 &	\textbf{78.7}
+ \\
+$\overline{Rank}$ & 7.73 & 6.82 & 4.55 & 4.09 & 3.91 & \textbf{2.27} & 3.18 & 2.45 \\
  \hline 
 \end{tabular}
-\label{table_pathways}
+\label{fig:cdnk-pathways}
 \end{table}
-
 \section{Conclusions}

### 2017-10-28 neurocomputing.tex
F

@@
 \begin{abstract}
-Biological relations encoded in data sources are naturally represented in the form of graphs.
-As a consequence, a high number of graph-based biological learning systems have been proposed and shown promissing results.
-In graph-based learning systems, one of the key points which determines their performance is the denifition of graph node similarity measure.
+Due to advances in data acquisition technologies (such as next generation sequencing), large amounts of information on biological systems are being encoded in relational data bases at an increasing rate.
+State of the art systems capable to deal with such structured data at scale,  include graph approaches based on the notion of diffusion kernels.
+In these systems, information is propagated in a discounted way according to the topology of the data network.
+A disadvantage of these approaches is that the specific configuration of the neighborhood of each node is ignored.
+Moreover it would be beneficial to integrate, when available, side information, encoded as arbitrary feature vectors associated to nodes or edges.
+%Biological relations encoded in data sources are naturally represented in the form of graphs.
+As a consequence, a high number of graph-based biological learning systems have been proposed and shown promising results.
+In graph-based learning systems, one of the key points which determines their performance is the definition of graph node similarity measure.
 Node similarities are normally measured by graph node kernels.
@@
 Second they are not able to proceed the auxiliary information associated to graph nodes.
-
-In this paper, we propose an efficient graph node kernel that not only effectively exploits nodes' context, but also process with additional information which available on the graph nodes.
-Empirical evaluation on disease gene prioritization show that our proposed graph node kernel shows the state of the art concerning graph node kernels.
+In this paper, we propose a graph decomposition approach that can exploit the information present in the local topology and that can efficiently incorporate vectorial side information.
+We consider the challenging problem of disease gene prioritization and show that we can achieve  state-of-the-art results.
 
@@
 \begin{keyword}
-Graph node kernels, graph decomposition, disease gene prioritization
+Graph node kernels, disease gene association
 \end{keyword}
@@
 \section{Introduction}
+
 The release of advanced technologies is one of the main reasons for the
-revolution in various scientific research fields. In Biological and Medical
-domain, modern technologies are making it not only easier but also more
-economical than ever to undertake experiments and creating applications.
-As a consequence, a vast amount of biological data in terms of volume
-and type is generated through scientific experiments, published literature,
-high-throughput experiment technology, and computational analysis. This
-huge quantity of data are saved as biological datasets and made discoverable
-through web browsers, application programming interfaces, scalable search
-technology and extensive cross-referencing between databases. Biological
-databases normally contain information about gene function, structure, lo-
-calization, clinical effects of mutations and similarities of biological sequences and structures.
+revolution in various scientific research fields.
+In Biological and Medical domain, modern technologies are making it not only easier but also more economical than ever to undertake experiments and creating applications.
+As a consequence, a vast amount of biological data in terms of volume and type is generated through scientific experiments, published literature, high-throughput experiment technology, and computational analysis.
+This huge quantity of data are saved as biological datasets and made discoverable through web browsers, application programming interfaces, scalable search technology and extensive cross-referencing between databases.
+Biological databases normally contain information about gene function, structure, localization, clinical effects of mutations and similarities of biological sequences and structures.
 
@@
 
+\subsection{Parameter Impact}
+In order to evaluate the stability of kernel w.r.t the change of each parameter values, we perform experiments with disease-gene association 2 and on the BioGPS network.
+For each parameter, first fix the values of all other parameters, but allow the value of this papameter to change.
+More precisely, we use the optimal tuple $C = 5, d^* = 1, D = 20, P = 15$ for the fixed of parameters' values.
+We let the changes of $C$ in $\lbrace 4, 5, 6, 7, 8, 9\rbrace$, $d^*$ in $\lbrace 1, 2, 3, 4\rbrace$, $D$ in $\lbrace 8, 10,  15, 20, 30, 40, 50\rbrace$ and real vector size, $P$ in $\lbrace 8, 10,  15, 20, 30, 50, 70, 90\rbrace$
+
 \subsection{Results and Discussion}
@@
 
-2 & 81.7/5 & 78.5/6 & 75.2/7 & 75.0/8 & 88.3/1 & \textbf{88.8/1} & 85.0/3 & 84.1/4 \\
+2 & 81.7/5 & 78.5/6 & 75.2/7 & 75.0/8 & 88.3/2 & \textbf{88.8/1} & 83.2/4 & 84.8/3 \\
 
@@
 \hline
-$\overline{AUC}$ & 66.6	& 63.5 & 67.8 &	67.6 & 73.2 & \textbf{77.8} & 73.6 & 75.8 \\
+$\overline{AUC}$ & 66.6	& 63.5 & 67.8 &	67.6 & 73.2 & \textbf{77.8} & 73.5 & 75.9 \\
 
-$\overline{Rank}$ & 6.09 & 7.09 &  5.36 & 4.82 & 4.45 & \textbf{1.64} & 4.09 & 2.36 \\
+$\overline{Rank}$ & 6.09 & 7.09 &  5.36 & 4.82 & 4.45 & \textbf{1.64} & 4.18 & 2.27 \\
  \hline 
@@
 \end{table}
+
+\newpage
+\begin{figure}[!tbp]
+  \centering
+  \begin{minipage}[b]{0.47\textwidth}
+    \includegraphics[width=\textwidth]{cliques.png}
+    \caption{Performance with C changes}
+    \label{fig:c-change}
+  \end{minipage}
+  \hfill
+  \begin{minipage}[b]{0.47\textwidth}
+    \includegraphics[width=\textwidth]{distance.png}
+    \caption{Performance with d* changes}
+    \label{fig:d-change}
+  \end{minipage}
+\end{figure}
+
+\begin{figure}[!tbp]
+\centering
+  \begin{minipage}[b]{0.47\textwidth}
+    \includegraphics[width=\textwidth]{degree.png}
+    \label{fig:D-change}
+    \caption{Performance with D changes}
+  \end{minipage}
+  \begin{minipage}[b]{0.47\textwidth}
+    \includegraphics[width=\textwidth]{vec_size.png}
+    \label{fig:vs-change}
+    \caption{Performance with P changes}
+  \end{minipage}  
+\end{figure}
+
 \section{Conclusions}

### 2017-11-09 neurocomputing.tex
modify pics

@@
 \usepackage{fixme}
-\fxsetup{
-    status=draft,
-    author=,
-    layout=inline,
-    theme=color
-}
+\usepackage{subfig}
+\usepackage{float}
+%\usepackage[caption = false]{subfig}
+%\fxsetup{
+%    status=draft,
+%    author=,
+%    layout=inline,
+%    theme=color
+%}
 
@@
 In this section, we first introduce definitions and notations that are used to define our proposed method. We then describe the state-of-the-art concerning graph node kernels.
-\subsection{Definitions and Notations}
+\subsection{Definitions and notations}
 A graph is a structure $G=(\mathbb{V},\mathbb{E}, \mathcal{L}_1, \mathcal{L}_2)$ where $\mathbb{V}$, $\mathbb{E}, \mathcal{L}_1, \mathcal{L}_2$ are the vertex (node) set, link (edge) set, discrete labeling fuction and real vector labeling function, respectively.
@@
 Examples include Weisfeiler-Lehman kernels \cite{shervashidze2009fast, shervashidze2011weisfeiler}, The Neighborhood Subgraph Pairwise Distance Kernel (NSPDK) \cite{costa2010fast}.
-Following we describe NSPDK since it is later adopted to develop our graph node kernel (presented in \ref{method}).
+Following we describe NSPDK since it is later adopted to develop our graph node kernel (presented in Section \ref{method}).
 
 The NSPDK \cite{costa2010fast} is an instance of convolution kernel \cite{haussler1999convolution} where given a graph $G \in \mathcal{G}$ and two rooted graphs $A_u, B_v$, the relation $R_{r,d}(A_u, B_v, G)$ is true {\em iff} $A_u \cong \mathcal{N}_r^u$ is (up to isomorphism $\cong$) a neighborhood subgraph of radius $r$ of $G$ and so is $B_v \cong  \mathcal{N}_r^v$, with roots at distance $\mathcal{D}(u,v)= d$.
-Figure \ref{sec:ndpdk} illustrates a pairwise neighborhood subgraph.
+Figure \ref{fig:nspdk} illustrates a pairwise neighborhood subgraph.
 We denote $R^{-1}$ as the inverse relation that returns all pairs of neighborhoods of radius $r$ at distance $d$ in $G$, $R^{-1}_{r,d}(G) = \lbrace A_u, B_v | R_{r,d}(A_u,B_v,G)=true\rbrace$.
@@
 \cite{fouss2006experimental} and exploits the idea of diffusion distance, which is a measure of how similar the pattern of heat diffusion is among a pair of initialized nodes.
-In other words, it expresses how much nodes "influence" each other in a similar fashion.
+In other words, it expresses how much nodes ``influence" each other in a similar fashion.
 If their diffusion ways are alike, the similarity will be high and, vice versa, it will be low if they diffuse differently.
@@
 
-\subsection{Network Decomposition} 
+\subsection{Network decomposition} 
 \label{network-decomposition}
 In genetic networks, it is not uncommon to find nodes with high degrees.
-Unfortunately these cases cannot be effectively processed by a neighborhood based decomposition kernel (see \ref{graph-kernels}) since these are based on the notion of exact matches.
+Unfortunately these cases cannot be effectively processed by a neighborhood based decomposition kernel (see Section \ref{graph-kernels}) since these are based on the notion of exact matches.
 Neighborhood subgraphs rooted at high degree nodes are relatively big due to a high number of neighbors and edges.
@@
 \centering
-\includegraphics[width=.9\textwidth]{cliques}
+\includegraphics[width=.9\textwidth]{cliques.pdf}
 \caption{Clique decomposition with threshold $C = 4$}
@@
 
-\subsection{The Conjunctive Disjunctive Node Kernel}
+\subsection{The conjunctive disjunctive node kernel}
 %We start from the Neighborhood Subgraph Pairwise Distance Kernel (NSPDK) \cite{costa2010fast} and adapt it to express the similarity between nodes in a single network.
@@
 \includegraphics[width=.9\textwidth]{cdnk}
-\caption{Pairwise neighborhood subgraphs for the ``red'' node with $r=1$ and $d=3$ using ``\textit{conjuction}'' and ``\textit{disjuctive}'' edges}
+\caption{Pairwise neighborhood subgraphs for the ``red'' node with $r=1$ and $d=3$ using ``\textit{conjunctive}'' and ``\textit{disjunctive}'' edges}
 \label{fig:cdnk}
@@
 
-\section{Parameter Space}
+\section{Parameter space}
 Our kernel consists of five parameters: the threshold degree $D$, clique size threshold $C$, maximal radius $r^*$, maximal distance $d^*$ and number of clusters $P$ clusters in which their values are in $\mathbb{N}^*$.
@@
 In particular, the clique size threshold $C$ is recommended to be in $\lbrace 4, 5\rbrace$.
-The maximal radius $r^*$ is set with a value smaller than or equal to $3$ since it will lead to big neighborhood subgraphs (as discussed in \ref{network-decomposition}).
+The maximal radius $r^*$ is set with a value smaller than or equal to $3$ since it will lead to big neighborhood subgraphs (as discussed in Section \ref{network-decomposition}).
 The maximal distance $d^*$ is assigned with values less than or equal to $4$ because the value of the maximal shortest distance between nodes in a connected component is often not too high.
@@
 
-\section{Empirical Evaluation}
+\section{Empirical evaluation}
 \label{empirical-evaluation-cdnk}
@@
 
-\subsection{Experimental Settings and Evalution Method}
+\subsection{Experimental settings and evalution methods}
 We carry out experiments in the context of disease gene prioritization where the aim is to build a learning system which allows to prioritize candidate genes based on their probabilities of being associated to a given genetic disease.
@@
 
-The experiments are performed on two separate networks derived from BioGPS and Pathways datasets in which we follow the experiment procedure in \cite{chen2014disease} where 12 diseases \cite{goh2007human} are used in which each disease is associated to at least 30 positive genes (see table \ref{tab:cdnk-diseases} for the list of genetic diseases and the number of positive genes in each disease).
+The experiments are performed on two separate networks derived from BioGPS and Pathways datasets in which we follow the experiment procedure in \cite{chen2014disease} where 12 diseases \cite{goh2007human} are used in which each disease is associated to at least 30 positive genes (see \textit{Appendix} \footnote{https://github.com/dinhinfotech/Neurocoputing/blob/master/appendix.pdf} for the list of genetic diseases and the number of positive genes in each disease).
 For each disease, we construct a positive set $\mathcal{P}$ with all positive (confirmed) disease genes, and a negative set $\mathcal{N}$ which contains random genes associated at least to one disease class which is not related to the class that is defining the positive set.
@@
 
-\subsection{Node Labeling} 
+\subsection{Node labeling} 
 \label{sec:lab}
@@
 
-\subsection{Model Selection}
+\subsection{Model selection}
 The hyper parameters of the various methods are tuned using a k-fold cross validation.
@@
 
-\subsection{Parameter Impact}
+\subsection{Parameter impact}
 In order to evaluate the stability of kernel w.r.t the change of each parameter values, we perform experiments with disease-gene association 2 and on the BioGPS network.
 For each parameter, first fix the values of all other parameters, but allow the value of this papameter to change.
-More precisely, we use the optimal tuple $C = 5, d^* = 1, D = 20, P = 15$ for the fixed of parameters' values.
-We let the changes of $C$ in $\lbrace 4, 5, 6, 7, 8, 9\rbrace$, $d^*$ in $\lbrace 1, 2, 3, 4\rbrace$, $D$ in $\lbrace 8, 10,  15, 20, 30, 40, 50\rbrace$ and real vector size, $P$ in $\lbrace 8, 10,  15, 20, 30, 50, 70, 90\rbrace$
+More precisely, we use the optimal tuple $C = 5,\ d^* = 1,\ D = 20,\ P = 15$ for the fixed of parameters' values.
+We let the changes of $C$ in $\lbrace 4,\ 5,\ 6,\ 7,\ 8,\ 9\rbrace$, $d^*$ in $\lbrace 1,\ 2,\ 3,\ 4\rbrace$, $D$ in $\lbrace 8,\ 10,\ 15,\ 20,\ 30,\ 40,\ 50\rbrace$ and real vector size, $P$ in $\lbrace 8,\ 10,\ 15,\ 20,\ 30,\ 50,\ 70,\ 90\rbrace$
 
-\subsection{Results and Discussion}
-Table \ref{fig:cdnk-biogps} and \ref{fig:cdnk-pathways} show the AUC-ROC performance of different models using state of the art graph node kernels and different variations of CDNK on BioGPS and Pathways networks, respectively.
-In what follows, we analyze the results to answer for the two questions which are previously raised.
-
-Concering the performance of different graph node kernels, CDNK variations outperform diffusion-based graph node kernels in all cases on BioGPS dataset and 8 out of 11 cases on Pathways when compared to diffusion-based kernels.
-CDNK is ranked first when considering both the average AUC-ROC and the average rank with a difference, w.r.t.
-the best diffusion-based kernel, ranging from 5.4$\%$ to 10$\%$ and from 1.2$\%$ to 3.4$\%$ on BioGPS and Pathways, respectively.
-As a consequence, we can conclude that CDNK show state of the art performance in graph node kernels.
-
-Regarding the variations of CDNK, the integration of real valued vectors improves the performance in most cases.
-In particular, considering the use of discrete labels based on ontology, using side information helps to increase the performance in all diseases on BioGPS and in 9 out of 11 diseases on Pathways.
-Similarly, by employing auxiliary information associated to graph nodes, the performance of CNDK in case of using uniform discrete labels is also improved in 9 and 8 out of 11 on BioGPS and Pathways, respectively.
-%==================================================
-\section{Conclusion and Future Work}
-We have shown how decomposing a network in a set of connected sparse graphs allows us to take advantage of the discriminative power of CDNK, a novel decomposition kernel, to achieve state-of-the-art results.
-Moreover, we have also introduced the way to integrate ``side'' information in form of real valued vectors when it is available on graph to get even better performance of CDNK.
-In future work we will investigate how to \textit{i}) decompose networks in a data driven way and \textit{ii}) extend the CDNK approach to gene-disease association problems exploiting multiple heterogeneous information sources in a joint way.
+\subsection{Results and discussion}
 %======================================================
@@
 %==================================================
-\begin{table}
+\begin{table}[h]
 \caption{\textit {Predictive performance on 11 gene-disease associations in percentage using network induced by the BioGPS.
@@
 \end{table}
-
 %=====================================
 %=====================================
-\begin{table}
+\begin{table}[h]
 \caption{\textit {Predictive performance on 11 gene-disease associations in percentage using network induced by the Pathways.
@@
 \end{table}
+Table \ref{fig:cdnk-biogps} and \ref{fig:cdnk-pathways} show the AUC-ROC performance of different models using state of the art graph node kernels and different variations of CDNK on BioGPS and Pathways networks, respectively.
+In what follows, we analyze the results to answer for the two questions which are previously raised.
 
-\newpage
-\begin{figure}[!tbp]
-  \centering
-  \begin{minipage}[b]{0.47\textwidth}
-    \includegraphics[width=\textwidth]{cliques.png}
-    \caption{Performance with C changes}
-    \label{fig:c-change}
-  \end{minipage}
-  \hfill
-  \begin{minipage}[b]{0.47\textwidth}
-    \includegraphics[width=\textwidth]{distance.png}
-    \caption{Performance with d* changes}
-    \label{fig:d-change}
-  \end{minipage}
-\end{figure}
+Concering the performance of different graph node kernels, CDNK variations outperform diffusion-based graph node kernels in all cases on BioGPS dataset and 8 out of 11 cases on Pathways when compared to diffusion-based kernels.
+CDNK is ranked first when considering both the average AUC-ROC and the average rank with a difference, w.r.t.
+the best diffusion-based kernel, ranging from 5.4$\%$ to 10$\%$ and from 1.2$\%$ to 3.4$\%$ on BioGPS and Pathways, respectively.
+As a consequence, we can conclude that CDNK show state of the art performance in graph node kernels.
 
-\begin{figure}[!tbp]
-\centering
-  \begin{minipage}[b]{0.47\textwidth}
-    \includegraphics[width=\textwidth]{degree.png}
-    \label{fig:D-change}
-    \caption{Performance with D changes}
-  \end{minipage}
-  \begin{minipage}[b]{0.47\textwidth}
-    \includegraphics[width=\textwidth]{vec_size.png}
-    \label{fig:vs-change}
-    \caption{Performance with P changes}
-  \end{minipage}  
-\end{figure}
-
-\section{Conclusions}
+Regarding the variations of CDNK, the integration of real valued vectors improves the performance in most cases.
+In particular, considering the use of discrete labels based on ontology, using side information helps to increase the performance in all diseases on BioGPS and in 9 out of 11 diseases on Pathways.
+Similarly, by employing auxiliary information associated to graph nodes, the performance of CNDK in case of using uniform discrete labels is also improved in 9 and 8 out of 11 on BioGPS and Pathways, respectively.
+%==================================================
+\section{Conclusion and future work}
 We have shown how decomposing a network in a set of connected sparse graphs allows us to take advantage of the discriminative power of CDNK, a novel decomposition kernel, to achieve state-of-the-art results.
@@
 In future work we will investigate how to \textit{i}) decompose networks in a data driven way and \textit{ii}) extend the CDNK approach to gene-disease association problems exploiting multiple heterogeneous information sources in a joint way.
+
+\begin{figure}[h]
+\centering
+\subfloat[Subfigure 2 list of figures text][]{
+\includegraphics[width=0.4\textwidth]{cliques.png}
+\label{fig:subfig1}}
+\subfloat[Subfigure 3 list of figures text][]{
+\includegraphics[width=0.4\textwidth]{distance.png}
+\label{fig:subfig2}}
+\qquad
+\subfloat[Subfigure 2 list of figures text][]{
+\includegraphics[width=0.4\textwidth]{degree.png}
+\label{fig:subfig3}}
+\subfloat[Subfigure 3 list of figures text][]{
+\includegraphics[width=0.4\textwidth]{vec_size.png}
+\label{fig:subfig4}}
+\caption{The performance of the system w.r.t the change of each hyper-parameter: (a) clique threshold, (b) distance threshold, (c) degree threshold, (d) real vec size threshold}
+\label{fig:globfig}
+\end{figure}
+
 %\section{The Elsevier article class}

### 2017-11-16 Response_neurocomputing.tex
final version

--- 2017-10-11 Response_3.tex
@@
 \usepackage[utf8x]{inputenc}
-\title{Revision notes for paper\\``The Conjunctive Disjunctive Graph Node Kernel''}
+\title{Revision notes for paper\\``The Conjunctive Disjunctive Graph Node Kernel for Disease Gene Prioritization''}
 \author{Dinh Tran Van, Alessandro Sperduti and Fabrizio Costa }
@@
 
-We thank the reviewer for the reviews and helpful comments on our manuscript ``The Conjunctive Disjunctive Graph Node Kernel''. 
-In the following, we provide a description of how each comment (in \emph{italic}) has been addressed in this version of the paper.
+We thank the reviewer for the reviews and helpful comments on our manuscript ``The Conjunctive Disjunctive Graph Node Kernel for
+Disease Gene Prioritization''. In the following, we provide a description of how each comment (in \emph{italic}) has been addressed in this version of the paper.
 
@@
 
-\underline{Answer}:
+\underline{Answer}: we explain in detail in ``Section Graph Node Kernels" of the revised version as follows.\\
+
+In DK, the similarity values between high degree vertices are generally higher compared to those between low degree ones.
+Intuitively, the more paths connect two vertices, the more heat can flow between them.
+This could be problematic since peripheral nodes (i.e., low degree nodes) have unbalanced similarities with respect to central nodes (i.e., high degree nodes).
+In order to make the strength of individual vertices comparable, a modified version of DK is Markov exponential diffusion kernel (MED) and given by the following formula:
+\begin{equation}
+\textbf{K} = e^{-\beta \textbf{M}}.
+\end{equation}
+The difference with respect to the DK is the replacement of $\textbf{L}$ by the matrix $\textbf{M}=(\textbf{D}-\textbf{A}-n\textbf{I})/n$ where $n$ is the total number of vertices in the graph.
+In particular, the introduced modification of the Laplacian is designed to compensate the problem described above via a sort of normalization of the degree of vertices.
+The role of $\beta$ is the same as for DK.
 
 \item \textit{lines 20-25: what does "process in an additive and independent fashion" mean in this context.
-Similarly as before, please justify your claim about the poor discriminative power of these kernels;
-}
+Similarly as before, please justify your claim about the poor discriminative power of these kernels;}
 
-\underline{Answer}:
+\underline{Answer}: we explain in more intuitive way in the new version of the paper.\\
+
+The state-of-the-art graph node kernels used to measure node similarity are based on the notion of information diffusion, which mainly depends on the number of paths connecting two nodes in the graph.
+These graph node kernels often show relatively low discriminative capacity, especially when working with sparse graphs (i.e., graphs with a  low number of  links), because of the following limitations.
+First, they are defined using a heat diffusion dynamics which sums up the contributions of all  paths from one node to another one, disregarding the local topological context of the nodes and considering the single contribution of one path as independent with respect to the contributions of other paths.
+Second, they do not take into account additional information (i.e. properties of a single node) eventually associated to nodes of graphs, when available.
 
@@
 
-\underline{Answer}:
+\underline{Answer}: The existance of high degree nodes will lead to the big neighborhood subgraphs.
+Since the proposed kernel is based on the exact matching neighborhood subgraphs (isomorphic neighborhood subgraphs), there will be a low probability to find any couple of big neighborhood subgraphs which are isomorphic.
+This probabability decrease shapely w.r.t the increase of neighborhood subgraphs size or node degree.
 
@@
 How robust is this clustering to very incomplete information as found in GO databases?
-In a similar fashion, can you provide an evaluation of your
-decomposition on the networks you are working with?}
+In a similar fashion, can you provide an evaluation of your decomposition on the networks you are working with?} \\
 
-\underline{Answer}:
+\underline{Answer}: We agree with the suggestion of the reviewer about the title of the kernel that it should be toward to disease gene prioritization.
+Therefore, we have modified it as "The conjunctive disjunctive graph node kernel for disease gene prioritization" in the new version of the paper.\\
+
+In general, our kernel takes labeled graphs as the input.
+Given a graph, users can apply any proper node labeling function to the graph.
+The labeling functions proposed in the paper are examples to evaluate the performance of the disease gene prioritization system with the use of our kernel.
+Regarding the proposed labeling functions, we consider topological labels as discrete labels.
+The GO datasets do not provide complete information.
+However, they still provide additional information for the kernel.
+We strongly agree that the intensitive evaluation for labeling functions is necessary to improve the performance of the proposed kernel.
+However, in this paper, we would like to have an attempt to see how the CDNK performs compared to other graph node kernels.
 
@@
 
-\underline{Answer}:
+\underline{Answer}: we have re-organized the structure of the paper based on the reviewer's comments.
 
@@
 
-\underline{Answer}:
+\underline{Answer}: we have modified the paper based on the reviewer's comment.
 
 \item \textit{Experiments (since the discussion section is very short, I don't think that it is necessary or even relevant to have it separated from the rest of the paper)}
+
+\underline{Answer}: we have modified the paper based on the reviewer's comment.
+
 \end{enumerate}
-
-\underline{Answer}:
-
 \subsection*{4} \textit{The experimental part is very light.
-I would have expected to see an smart evaluation on a simulated dataset illustrating in which theoretical cases the approach is relevant.
+I would have expected to see a smart evaluation on a simulated dataset illustrating in which theoretical cases the approach is relevant.
 I think that an evaluation of the sensitivity of the method to some parameters is missing.
@@
 (To my own experience, this is a very low threshold that should give a very dense network and is not the standard choices in this area).
-Also, why have twice less negative examples than positive ones (this is very far from being realistic since positive genes are usually very rare in this kind of real-life applications)?}
+Also, why have twice less negative examples than positive ones (this is very far from being realistic since positive genes are usually very rare in this kind of real-life applications)?} \\
 
-\underline{Answer}:
+\underline{Answer}: We would like to compare our node kernel diffusion-based graph node kernels on relatively dense networks, since diffusion-based kernels often show good performance with dense networks.
+That is why we borrow BioGPS and Pathways which are used in [2].
+\\
+
+In disease prioritization, given a genetic disease, we are often given a small set of positive genes and a large set of un-labeled genes.
+It is difficult to have a proper negative set of genes.
+Therefore, in [2] they choose the number of postive genes two times greater than the number of negative ones.
+The negative genes are choosen from genes that are associated with at least one disease class, but are not related to the class that is defining the positive set.
+We assume that they are less likely to be associated to the considered class, since disease genes are generally more studied and a potential association has more chances to be identified.
 
@@
 
-\underline{Answer}:
+\underline{Answer}: we have modified the structure of the paper in order to be easier to follow.
 
@@
 
-\underline{Answer}:
+\underline{Answer}: We have added a more detailed explanation in the new version of the paper based on the reviewer's comment.
 
@@
 
-\underline{Answer}: The large number of parameters is a drawback of our propoased graph node kernel. However, some of them are suggested to have smale range of values to be effective.
+\underline{Answer}: We have added a new section named ``Hyper-parameter" to explain why the parameter space for the proposed kernel is not high as follows.\\
+
+Our kernel consists of five hyper-parameters: the threshold degree $D$, the clique size threshold $C$, the maximal radius $r^*$, the maximal distance $d^*$, and the number of clusters $P$.
+In order to choose the optimal tuple of   values for kernel hyper-parameters in a specific setting, a model selection procedure is normally adopted.
+The hyper-parameter space of our kernel is potentially large due to the relatively high number of hyper-parameters.
+However, all hyper-parameters are integer variables and it makes sense to use only small values for them.
+For example, the value of the threshold degree $D$ should neither be too high nor too low: if it is high, the neighborhood subgraphs are going to be too big and the isomorphism check becomes inefficient; vice versa, if it is too low, the decomposed graph contains connected components which are too sparse to allow learning to be effective.
+Moreover, it is reasonable to keep the clique size threshold $C$ within a narrow range of values in order to avoid the removal of too many edges.
+Concerning the maximal radius $r^*$,  too  high values will lead again to big neighborhood subgraphs, especially in graphs with a high node degree average.
+The maximal distance $d^*$ can be set with small values since the average shortest distance length between nodes of biological networks is relatively low.
+Finally, if the value assigned to $P$ is too low, all nodes basically get the same information attached to them and thus the labels are not informative; vice versa, if a high value is used, there will be very few matches and the kernel will be extremely sparse, making learning ineffective.
 
@@
 \end{itemize}
+
+\bibliography{mybibfile}
 \end{document}

### 2017-11-16 neurocomputingv3.tex
add main files

\documentclass[review]{elsarticle}

\usepackage{lineno,hyperref}
\newtheorem{definition}{Definition}
\usepackage{amssymb,amsmath,array}
\usepackage{xcolor}% provides \colorlet
\usepackage{fixme}
\usepackage{subfig}
\usepackage{float}
\fxsetup{
    status=draft,
    author=,
    layout=inline,
    theme=color
}

\definecolor{fxnote}{rgb}{0.8000,0.0000,0.0000}
% define the background colour:
\colorlet{fxnotebg}{yellow}

% refedine the layout macro:
\makeatletter
\renewcommand*\FXLayoutInline[3]{%
  \@fxdocolon {#3}{%
    \@fxuseface {inline}%
    \colorbox{fx#1bg}{\color {fx#1}\ignorespaces #3\@fxcolon #2}}}
\makeatother

\modulolinenumbers[5]

\journal{Journal of \LaTeX\ Templates}

\newcommand{\IR}{\rm \mathbb{R}}

%%%%%%%%%%%%%%%%%%%%%%%
%% Elsevier bibliography styles
%%%%%%%%%%%%%%%%%%%%%%%
%% To change the style, put a % in front of the second line of the current style and
%% remove the % from the second line of the style you would like to use.
%%%%%%%%%%%%%%%%%%%%%%%

%% Numbered
%\bibliographystyle{model1-num-names}

%% Numbered without titles
%\bibliographystyle{model1a-num-names}

%% Harvard
%\bibliographystyle{model2-names.bst}\biboptions{authoryear}

%% Vancouver numbered
%\usepackage{numcompress}\bibliographystyle{model3-num-names}

%% Vancouver name/year
%\usepackage{numcompress}\bibliographystyle{model4-names}\biboptions{authoryear}

%% APA style
%\bibliographystyle{model5-names}\biboptions{authoryear}

%% AMA style
%\usepackage{numcompress}\bibliographystyle{model6-num-names}

%% `Elsevier LaTeX' style
\bibliographystyle{elsarticle-num}
%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

\begin{frontmatter}

\title{The Conjunctive Disjunctive Graph Node Kernel for Disease Gene Prioritization}
%\tnotetext[mytitlenote]{Fully documented templates are available in the elsarticle package on \href{http://www.ctan.org/tex-archive/macros/latex/contrib/elsarticle}{CTAN}.}

%% Group authors per affiliation:
\author{Dinh Tran Van}
\author{Alessandro Sperduti}
\address{Department of Mathematics, Padova University, Trieste, 63, 35121 Padova, Italy}
\author{Fabrizio Costa}
\address{Department of Computer Science, University of Exeter Exeter EX4 4QF, UK}
%\fntext[myfootnote]{Since 1880.}

%% or include affiliations in footnotes:
%\author[mymainaddress,mysecondaryaddress]{Elsevier Inc}
%\ead[url]{www.elsevier.com}

%\author[mysecondaryaddress]{Global Customer Service\corref{mycorrespondingauthor}}
%\cortext[mycorrespondingauthor]{Corresponding author}
%\ead{support@elsevier.com}

%\address[mymainaddress]{1600 John F Kennedy Boulevard, Philadelphia}
%\address[mysecondaryaddress]{360 Park Avenue South, New York}

\begin{abstract}
Biological relations encoded in data sources are naturally represented in the form of graphs. As a consequence, a high number of graph-based biological learning systems have been proposed and shown promising results. In graph-based learning systems, one of the key points which determines their performance is the definition of the graph node similarity measure. Node similarities are normally measured by graph node kernels. However, most available graph node kernels do not show high discriminative capacity since they share two common limitations. First, they are based on the diffusion phenomenon which does not effectively exploit the nodes' context. Second, they are not able to process the auxiliary information associated to graph nodes.

In this paper, we propose an efficient graph node kernel that not only is able to effectively take into account nodes' context, but also to  exploit additional information  available on graph nodes. Empirical evaluation on disease gene prioritization shows that our proposed graph node kernel is able to reach  state-of-the-art performances when considering graph node kernels.

%Gene-disease associations are inferred on the basis of similarities between the proteins encoded by genes. Biological relationships used to define similarities range from interacting proteins, proteins that participate in pathways and protein expression profiles. Though graph kernel methods have become a prominent approach for association prediction, most solutions are based on a notion of information diffusion that does not capture the specificity of different network parts. Here we propose a graph kernel method that explicitly models the configuration of each gene's context. An empirical evaluation on several biological databases shows that our proposal achieves state-of-the-art results.
\end{abstract}

\begin{keyword}
Graph node kernels, graph decomposition, disease gene prioritization
\end{keyword}
\end{frontmatter}

\linenumbers
\section{Introduction}
The release of advanced technologies is one of the main reasons for the
revolution in various scientific research fields. In Biological and Medical
domains, modern technologies are making it not only easier but also more
economical than ever to undertake experiments and creating applications.
As a consequence, a vast amount of biological data in terms of volume
and type is generated through scientific experiments, published literature,
high-throughput experiment technology, and computational analysis. 
%% DINH!!!!! EARLIER VERSION OF THE SENTENCE BELOW WAS BASICALLY COPIED FROM: The European Bioinformatics Institute in 2016: Data growth and integration
This huge quantity of data is made available in the form of biological databases, accessible via different 
tools, such as web browsers, application programming interfaces, and scalable search
engines.

%% DINH!!!!! EARLIER VERSION OF THE SENTENCE BELOW WAS BASICALLY COPIED FROM: OMICs Technologies: Tools for Food Science
The nature of the data contained in biological databases concerns information about genes, such as their function, structure, and localization, as well as information about clinical effects of mutations and similarities of biological sequences and structures.

In the Biomedical field, disease-gene association recovery is a major goal that has received much attention from many researchers. Despite the fact that a big progress has been made in the last decades, the typical number of genes known to be related to a genetic disease is normally limited. In order to find out the remaining unknown set of genes related to the disease, one way is to search for them  in the whole genome or in specific regions that often contain a large number of suspected genes (candidate genes). This is obviously not a good approach since it is expensive, not only in terms of time consumption, but also from a financial point of view. For this reason, many gene prioritization methods, which automatically predict a list of candidate genes sorted according to the probability to be actually involved in the target disease, have been proposed in literature. Predictive systems for gene-disease association are often based on a notion of similarity between genes inferred from the available knowledge encoded in the biological databases. In addition, many of them exploit machine learning methods to predict the gene-disease associations. Moreover, a common strategy is to encode relations between genes as a network where two genes are connected each other if they turn out to be very similar. Graph based techniques are then used to make useful inferences, as in \cite{mordelet2011prodige, chen2014disease,valentini2014extensive}.

One of the key points that determines the performance of graph-based biological learning systems in general, and graph-based gene-disease  prioritization systems in particular, is the definition of the node similarity measure. The node similarity is often measured by graph node kernels \cite{kondor2002diffusion,chen2014disease,fouss2006experimental,chebotarev2006matrix}. The state-of-the-art graph node kernels used to measure node similarity are based on the notion of information diffusion, which mainly depends on the number of paths connecting two nodes in the graph. These graph node kernels often show relatively low discriminative capacity, especially when working with sparse graphs (i.e., graphs with a  low number of  links), because of the following limitations. First, they are defined using a heat diffusion dynamics which sums up the contributions of all  paths from one node to another one, disregarding the local topological context of the nodes and considering the single contribution of one path as independent with respect to the contributions of other paths.
Second, they do not take into account additional information (i.e. properties of a single node) eventually associated to nodes of graphs, when available. These additional information normally provide a complement to the graph topology, and as such may contribute to improve, if used, the expressiveness of the kernel.

In this paper, we propose an effective convolutional graph node kernel, named \textit{Conjunctive Disjunctive Node Kernel} (CDNK) which is able to: \textit{i}) effectively exploit the nodes' context; \textit{ii}) involve in the computation of the kernel the additional information attached to graph' nodes, when available.
In particular, in CDNK, first, the graph is transformed into a set of linked
connected components in which we distinguish between “conjunctive” links
whose endpoints are in the same connected components and “disjunctive”
links that connect nodes located in different connected components. Then
the similarity between any couple of nodes is measured by employing a
particular graph kernel on two neighborhood subgraphs rooted at each
node. Next, it integrates the side information by applying convolution
of the discrete information with the real valued vectors associated to
graph nodes.

\section{Background}
In this section, we first introduce definitions and notations that are used to define our proposed method. We then describe the state-of-the-art concerning graph node kernels.
\subsection{Definitions and Notations}
A graph is a structure $G=(\mathbb{V},\mathbb{E}, \mathcal{L}_1, \mathcal{L}_2)$ where $\mathbb{V}$, $\mathbb{E}, \mathcal{L}_1, \mathcal{L}_2$ are the vertex (node) set, link (edge) set, discrete labeling fuction and real vector labeling function, respectively. The functions $\mathcal{L}_1, \mathcal{L}_2$ are defined as:
\begin{itemize}
\item $\mathcal{L}_1: \mathbb{V} \longmapsto \mathbb{L}$, where $\mathbb{L}$ is a set of discrete labels. $\mathcal{L}_1$ assigns a single discrete label $\ell \in \mathbb{L}$ for each node $v \in \mathbb{V}$, $\mathcal{L}_1(v) = \ell$. 
\item $\mathcal{L}_2: \mathbb{V} \longmapsto \mathbb{R}^n$. $\mathcal{L}_2$ assigns a single real vector label $(v_1,v_2,\ldots,v_n) \in \mathbb{R}^n$ for each ndoe $v \in \mathbb{V}$, $\mathcal{L}_2(v) = (v_1,v_2,\ldots,v_n)$.
\end{itemize}

We define the length of a shortest path between $u$ and $v$, denoted as $\mathcal{D}(u,v)$, as the number of edges on a shortest path between them. The \textit{neighborhood} of a node $u$ with radius $r$, $N_r(u) = \lbrace v\ |\ \mathcal{D}(u,v) \leq r \rbrace$, is the set of nodes at distance no greater than $r$ from $u$. The corresponding \textit{neighborhood subgraph} $\mathcal{N}_{r}^{u}$ is the  subgraph induced by the neighborhood (i.e. considering all the edges with endpoints in $N_r(u)$). The \textit{degree} of a node $u$, $deg(u) = |\mathcal{N}_{1}^{u}|$, is the cardinality of its neighborhood for $r=1$. The maximum node degree in the graph $G$ is $deg(G)=\max_{v\in \mathbb{V}}deg(v)$.

\begin{definition}{}
\textit{An adjacency matrix $\textbf{A}$ is a symmetric matrix used to characterize the direct links between vertices $v_{i}$ and $v_{j}$ in the graph. Any entry $A_{ij}$ is equal to $w_{ij}\in \mathbb{R}$ when there exists a link connecting $v_{i}$ and $v_{j}$, and is 0 otherwise.}. 
\end{definition}

\begin{definition}{}
\textit{The Laplacian matrix $\textbf{L}$ is defined as $\textbf{L} = \textbf{D}-\textbf{A}$, where $\textbf{D}$ is the diagonal matrix with non-null entries equal to the summation over the corresponding row of the adjacency matrix, i.e. $\textbf{D}_{ii}=\sum_j \textbf{A}_{ij}$.}
\end{definition}

\begin{definition}{}
\textit{The Transition matrix of a graph $G$, denoted as $\textbf{P}$, is a matrix with entries $P_{ij} = A_{ij}/\sum_{i}^{}A_{ij}$. When considering a random walk in $G$, $P_{ij}$ can be interpreted as proportional to the probability of moving from node $v_i$ to  node $v_j$.}
\end{definition}

\subsection{Kernels on Graphs}
Kernel methods have emerged as one of the most powerful framework in machine learning. They have been successfully applied in various domains, due to their modularity, i.e. the definition of kernel functions is independent from the design of the learning algorithm.

A kernel function can be  considered as the similarity measure between input instances, whatever the nature of the instances may be, e.g. vectors, sequences, trees, graphs.
% in the feature space. Interestingly, It can be defined in feature space and with any type of data representation. 
Formally, a kernel $k: \mathbb{X} \times \mathbb{X}\longmapsto \mathbb{R}$, where $\mathbb{X}$ is a set of entities, is a function satisfying the following properties: \textit{i}) $k$ is symmetric, i.e., $k(x_1,x_2) = k(x_2,x_1)$, where $x_1, x_2 \in \mathbb{X}$; \textit{ii}) $k$ is positive semi-definite, that is $\sum_{i=1}^{N}\sum_{j=1}^{N} c_i c_jk(x_i,x_j) \geq 0$ for any $N>0$, $c_i, c_j \in \mathbb{R}$, and  $x_i, x_j \in \mathbb{X}$.

Canonical machine learning methods take vectorial data, i.e. numerical vectors that collect the results of measures on features of the input instances, as their input. However, there are many fields where data is naturally represented by structured forms, such as graphs, one of the most popular representations for structured data. 
Two interesting examples of domains involving graphs are Chemistry and the Web. In Chemistry, chemical compounds are represented via their molecular graphs and typical computational tasks consist in the prediction of their physicochemical properties. Thus, in this domain, the target function to learn is a mapping from one graph to a real value. The Web can be described as a huge graph, where nodes are web pages and edges are links from one page to another one.  A typical task in this context is to automatically predict the topics covered by the textual content of a page on the basis of the characteristics of web pages connected with it. So, the  target function to learn is a mapping from one node of the graph to a set of discrete values.
Therefore, the definition of a kernel function for graphs has to take into account one of the two scenarios described above.
In the first case, we talk about graph kernels, while in the second case we talk about graph node kernels. Both graph kernels and graph node kernels are widely applied to build graph-based learning systems for fields ranging from Social Sciences, to Recommendation Systems, and Biology.

In general, one important contribution with respect to the design of kernels for structured data, and in particular for graphs, has been given by Haussler, who proposed a convolution-based framework for the definition of decompositional kernels \cite{haussler1999convolution}. 
In the following, we shortly describe some of the state-of-the-art graph (node) kernels underpinning our proposed approach.

%Examples are kernels proposed in \cite{shervashidze2011weisfeiler, gartner2003survey, borgwardt2005shortest, costa2010fast}. Following we describe NSPDK \cite{costa2010fast}, a convolutional graph kernel which shows state of performance and it is later on used to define our proposed graph node kernel.
\subsubsection{Graph Kernels}
\label{graph-kernels}
The task of designing efficient and expressive graph kernels plays an important role in the development of graph-based predictive systems. Existing graph kernels are decompositional kernels and can be classified into two categories: sequence-based graph kernels and subgraph-based graph kernels. The sequence-based graph kernels decompose graphs into ``parts" in sequence-based forms, such as paths and walks. Typical examples of sequence-based graph kernels are the product graph kernel \cite{gartner2003survey}, and the shortest path kernel \cite{borgwardt2005shortest}. The subgraph-based graph kernels decompose graphs into subgraphs. Examples include the Weisfeiler-Lehman kernel \cite{shervashidze2009fast, shervashidze2011weisfeiler}, and the Neighborhood Subgraph Pairwise Distance Kernel (NSPDK) \cite{costa2010fast}. 
This latter category of kernels are generally more effective because sub-graphs are more expressive than walks and paths.
Moreover, they can be computed quite efficiently thanks to sparsity of representation that allows an explicit encoding of the graph via hash functions.
In the following, we describe NSPDK, since it is later adopted to develop the proposed graph node kernel (presented in Section \ref{method}).

The NSPDK \cite{costa2010fast} is an instance of convolution kernel \cite{haussler1999convolution} where a given
 graph $G$ is decomposed in features (pairwise neighborhood subgraphs) constituted by couples of subgraphs of radius $r$ rooted at nodes of $G$ which are
 at distance $d$. More formally, given two rooted graphs $A_u, B_v$, where $u$ and $v$ are nodes of $G$, the relation $R_{r,d}(A_u, B_v, G)$ is true {\em iff}  $\mathcal{D}(u,v)= d$ and $A_u \cong \mathcal{N}_r^u$ is (up to isomorphism $\cong$) a neighborhood subgraph of radius $r$ of $G$ as well as $B_v \cong  \mathcal{N}_r^v$. Figure \ref{fig:nspdk} illustrates a pairwise neighborhood subgraph. We denote with $R^{-1}$ the inverse relation that returns all pairs of neighborhoods of radius $r$ at distance $d$ in $G$, $R^{-1}_{r,d}(G) = \lbrace (A_u, B_v) | R_{r,d}(A_u,B_v,G)=true\rbrace$. The kernel $\kappa_{r,d}$ over $\mathcal{G} \times \mathcal{G}$, counts the number of such fragments in common in two input graphs: 
\begin{center}
$\kappa_{r,d}(G,G^{'}) = 
\!\!\!\!\!\!\!\!\!\!\!\! 
\sum\limits_{\substack{(A_u, B_v) \ \in \ R_{r,d}^{-1}(G) \\ 
({A'}_{u^\prime}, {B'}_{v'}) \ \in \ R_{r,d}^{-1}(G')
}} \!\!\!\!\!\!\!\!\!\!\!\!  { { \textbf{1}_{A_{u} \cong A'_{u^\prime}}} \cdot {
\textbf{1}_{B_{v} \cong B'_{v'}}} }$, 
\end{center} 
\noindent where $\textbf{1}_{A \cong B}$ is the \textit{exact matching function} that returns 1 if $A$ is
isomorphic to $B$ and 0 otherwise.  Finally, the NSPDK is defined as $K(G,G') = \sum\limits_{r}{\sum\limits_{d}{\kappa_{r,d}(G,G')}}$, where for efficiency reasons, the values of $r$ and $d$ are upper bounded by  given maximal values $r^*$ and $d^*$, respectively.
\begin{figure}
\centering
\includegraphics[width=.9\textwidth]{nspdk}
\caption{\label{fig:nspdk} Example of pairwise neighborhood subgraphs for nodes $u$ and $v$ with $r=1$ and \mbox{$d=3$}. {\it Left:} input graph where the nodes $u$ and $v$ are connected by a shortest path (highlighted in red) of length 3; the neighborhood subgraphs of radius $1$ rooted at $u$ (highlighted in green) and at $v$ (highlighted in yellow) are shown as well. {\it Right:} representation of the induced pairwise neighborhood subgraphs feature.}
\end{figure}

\subsubsection{Graph Node Kernels}
The state-of-the-art diffusion-based kernels for nodes are based on the same equations used to model the heat diffusion phenomenon. In other words, they take paths connecting two nodes into account when measuring their similarity. In the following, we shortly introduce some of the most used diffusion-based graph node kernels.

%\begin{itemize}
%\item 
{\it Laplacian exponential diffusion kernel.} One of the most well-known graph node kernels is the Laplacian exponential diffusion kernel (DK), as it is widely used for exploiting discrete structures in general and graphs in particular. On the basis of the heat diffusion dynamics, Kondor and Lafferty proposed DK in \cite{kondor2002diffusion}: imagine to initialize each vertex with a given amount of heat and let it flow through the edges until an arbitrary instant of time. The similarity between any vertex couple $v_{i}$, $v_{j}$ is the amount of heat starting from $v_{i}$ and reaching $v_{j}$ within the given time. Therefore, DK can capture the long range relationship between vertices of a graph to define the global similarities. Below is the formula to compute LEDK values:
\begin{equation} 
\label{LEDK-formula}
\textbf{K} = e^{-\beta \textbf{L}} = \textbf{I} - \beta \textbf{L} + \frac{\beta \textbf{L}^{2}}{2!} - ...
\end{equation}
where $\beta$ is the diffusion parameter and is used to control the rate of diffusion, and $\textbf{I}$ is the identity matrix. Choosing a consistent value for $\beta$ is very important: on the one side, if $\beta$ is too small, the local information cannot be diffused effectively and, on the other side, if it is too large, the local information will be lost. DK is positive semi-definite as proved in \cite{kondor2002diffusion}.

%\item 
{\it Exponential diffusion kernel.} In DK, the similarity values between high degree vertices are generally higher compared to those between low degree ones. Intuitively, the more paths connect two vertices, the more heat can flow between them. This could be problematic since peripheral nodes (i.e., low degree nodes) have unbalanced similarities with respect to central nodes (i.e., high degree nodes). In order to make the strength of individual vertices comparable, a modified version of DK is introduced by Chen et al in \cite{chen2014disease}, called Markov exponential diffusion kernel (MED) and given by the following formula:
\begin{equation} \label{MEDK-formula}
\textbf{K} = e^{-\beta \textbf{M}}.
\end{equation}
The difference with respect to the Laplacian diffusion kernel is the replacement of $\textbf{L}$ by the matrix $\textbf{M}=(\textbf{D}-\textbf{A}-n\textbf{I})/n$ where $n$ is the total number of vertices in the graph. In particular, the introduced modification of the Laplacian is designed to compensate the problem described above via a sort of normalization of the degree of vertices. The role of $\beta$ is the same as for DK.

%\item 
{\it Markov diffusion kernel.} The original Markov diffusion kernel (MD) was introduced by Fouss et al. \cite{fouss2006experimental} and exploits the idea of diffusion distance, which is a measure of how similar the pattern of heat diffusion  between a pair of  nodes is. In other words, it expresses how much nodes ``influence'' each other in a similar fashion. If their diffusion patterns are alike, the similarity will be high and, vice versa, it will be low if they diffuse differently. This kernel is computed as:
\begin{equation} 
\label{MDK-formula}
\textbf{K} = \textbf{Z}(t)\textbf{Z}^{\top}(t),
\end{equation}
where $\textbf{Z}(t) = \frac{1}{t}\sum_{\tau=1}^{t}\textbf{P}^{\tau}$, and $\textbf{P}$ is the transition matrix.

%\item 
{\it Regularized Laplacian kernel.} Another popular graph node kernel function used in graph mining is the regularized Laplacian kernel (RL). This kernel function was introduced by Chebotarev and Shamis in \cite{chebotarev2006matrix}, and it implements a normalized version of the random walk with restart model. It is defined as follows:
\begin{equation} 
\label{RLK-formula}
\textbf{K} = \sum_{n=0}^{\infty}\beta^{n}(-\textbf{L})^n = (\textbf{I} + \beta \textbf{L})^{-1},
\end{equation}
where the parameter $\beta$ is again the diffusion parameter. RL counts the paths connecting two nodes on the graph induced by taking $-\textbf{L}$ as the adjacency matrix, regardless of the path length. Thus, a non-zero value is assigned to any couple of nodes as long as they are connected by any indirect path. RL remains a relatedness measure even when the diffusion factor is large, by virtue of the negative weights assigned to self-loops.

%\end{itemize}
\section{A New Graph Node Kernel}
\label{method}
As already argued in the introduction, diffusion-based kernels compute the similarity between two nodes only on the basis of how well they are connected within the graph. Thus, it can be stated that they exploit the global topology of the graph to compute the output. Small values of the diffusion parameter $\beta$ can be used to give more emphasis to shorter paths with respect to longer ones in order to take into account the local topology around a node, however the effect of this setting is homogeneous through the whole set of nodes. 
The kernel we propose in this paper is, on the contrary, based on the exploitation of a more detailed information on the local topology around a node. We obtain this information by using, in a specific way, the similarity notion computed by the neighborhood based decomposition kernel NSPDK \cite{costa2010fast}. The underpinning idea of the new kernel is to capture local topological information by looking at local  subgraphs rooted at the two nodes of interest. This can be done, in principle, by a direct application of the NSPDK kernel to the local subgraphs. In practice, when a node has a high degree, the local subgraph may be quite large and, since NSPDK is based on exact matching, the probability to have an exact match between the two local subgraphs decreases with the increase of  node degree. For this reason, we propose to decompose the original graph into a set of sparse connected components that singularly have lower node degree with respect to the original graph. The above idea is then applied to these set of components, and the novel kernel is computed by summing up all the obtained contributions from these components.  
In addition to that, we also extend the definition of the kernel in such a way to allow  the involvment  of additional information, that may be associated to each node, in the computation. In the following, we give a more detailed description of the proposed kernel.

Given an input undirected labeled graph $G=(\mathbb{V},\mathbb{E}, \mathcal{L}_1, \mathcal{L}_2)$, the computation of our kernel consists of two phases. In the first phase, a network decomposition procedure is applied to transform the graph into a set of linked sparse connected components. In this procedure, we define two different kinds of link: \textit{conjunctive} and \textit{disjunctive}, that we treat in distinct manners.
Nodes linked by conjunctive edges are going to be used jointly to define the notion of context and will be visible to the NSPDK neighborhood graph kernel. Nodes linked by disjunctive edges are instead used to define features based only on the pairwise co-occurrence of the genes at the endpoints and are processed by our novel kernel. In the second step, the similarity between any node couple ($u$, $v$) is computed by applying NSPDK to the two neighborhood subgraphs rooted at nodes $u$ and $v$, extracted from the graph obtained from the first decomposition phase. In the following, we describe each phase in detail.

\subsection{Network Decomposition} 
\label{network-decomposition}
In genetic networks, it is not uncommon to find nodes with high degrees. Unfortunately these cases cannot be effectively processed by a neighborhood based decomposition kernel (see Section \ref{graph-kernels}) since these are based on the notion of exact matches. Neighborhood subgraphs rooted at high degree nodes are relatively big due to a high number of neighbors and edges. Since the number of non-isomorphic graphs grows exponentially with the number of nodes and edges, this leads to a low likelihood of finding neighborhood subgraphs rooted at high degree nodes which are isomorphic. 
%Thus, the probability of having identical neighborhoods decreases exponentially as the degree increases.  
This means that in a finite network it quickly becomes impossible to find any match and hence to learn or to generalize at all.

In order to solve the above problem, we propose a procedure to ``sparsify'' the input network. In practice, we mostly keep the same the cardinality of the edge set. However, we mark the edges with special attributes so that the novel kernel is able to treat them differently during computation. The result is a procedure that decomposes the network in a linked collection of sparse sub-networks where each node has a reduced connectivity when considering the edges of a specific type. However the other edges are still available to connect the various sub-networks.

\textit{Iterative k-core decomposition} \cite{alvarez2005k}: the node set is partitioned in two groups on the basis of the degree of each node w.r.t. a threshold degree $D$, the first part contains all nodes with degree smaller than or equal to $D$ and the second part the remaining ones. The node partition is used to induce the ``conjunctive'' vs ``disjunctive'' notion for the edge partition: edges that have both endpoints in the same part are marked as conjunctive, otherwise they are marked as disjunctive. We apply the k-core decomposition iteratively, where at each iteration we consider only the graph induced by the conjunctive edges. We stop iterating the decomposition after a user defined number of steps. Note that this decomposition does not alter the cardinality of the edge set, it is simply a procedure to mark each edge with the attribute conjunctive or disjunctive.

\begin{figure}
\centering
\includegraphics[width=.9\textwidth]{k_core}
\caption{Example of application of one step of \textit{k-core decomposition} using degree threshold $D = 3$. {\it Left:} the input graph. \textit{Right:} resulting marking of the edges: an edge $(u,v)$ is marked as {\it disjunctive} (represented by dashed line) if either $ deg(u) \leq 3 \wedge  deg(v) > 3 $ or $ deg(u) > 3 \wedge  deg(v) \leq 3 $; all remaining edges are marked as {\it conjunctive} (represented by solid line).}
\label{fig:kcore-decomposition}
\end{figure}

\textit{Clique decomposition} \cite{tarjan1985decomposition}: to model the notion that nodes in a clique are tightly related, we summarize the whole clique with a new ``representative'' node. All the cliques (completely connected subgraphs) with a number of nodes greater than or equal to a given threshold  $C$ are identified. The endpoints of all edges incident on the clique's nodes are moved to the representative node. Disjunctive edges are introduced to connect each node in the clique to the representative. Finally all edges with both endpoints in the clique are removed.

In our work a graph $G$ is transformed by  first applying the iterative k-core decomposition and then the clique decomposition. Let denote with $G^{\wedge\vee}$ the graph obtained as output of this transformation.

\begin{figure}
\centering
\includegraphics[width=.9\textwidth]{cliques.pdf}
\caption{Example of application of the {\it clique decomposition} step using threshold $C = 4$. {\it Left:} the input graph; nodes coloured in blue belong to a clique of size 4. \textit{Right:} the resulting graph after {\it clique decomposition}; a representative node (green color) is introduced for the clique of size 4; edges belonging to the clique are removed, while disjunctive edges are introduced to connect each node in the clique to the representative; in addition, edges not belonging to the clique but with one endpoint in clique's nodes  are moved  to the representative node. }
\label{fig:clique-decomposition}
\end{figure}

\subsection{The Conjunctive Disjunctive Node Kernel}
%We start from the Neighborhood Subgraph Pairwise Distance Kernel (NSPDK) \cite{costa2010fast} and adapt it to express the similarity between nodes in a single network. The key idea in NSPDK is to decompose graphs in small fragments and count how many pairs of fragments are shared between two instances. We introduce two improvements: \textit{i}) we partition the features according to the individual node's neighborhood, and \textit{ii}) we introduce a  distinction between ``disjunctive'' and ``conjunctive'' edges.

We define a node kernel $K(u,u^\prime)$ between  nodes $u$ and $u^\prime$ belonging to $G^{\wedge\vee}$ as follows.
% The idea is to define the features of a node $u$ as the subset of NSPDK features that always have the node $u$ as one of the roots.  THIS SENTENCE IS NOT TRUE!!
%In addition we distinguish between two types of edges, called {\em conjunctive} and {\em disjunctive} edges. THIS IS  A REPETITION
When computing distances to induce neighborhood subgraphs, only conjunctive edges are considered. When choosing the
pair of neighborhoods to form a single feature, we additionally consider roots $u$ and $v$ that are not at distance $d$ but such that $u$ is connected to $w$ via a disjunctive edge and such that $w$ is at distance $d$ from $v$ (Figure \ref{fig:cdnk} is an illustration). In this way disjunctive edges can still allow an {\em information flow} even if their
endpoints are only considered in a pairwise fashion and not jointly.

In order to obtain a formal definition of the proposed kernel, we start by defining new specific notions of distance and neighborhood subgraph, which only depends on conjunctive edges. With $\mathcal{D}^{\wedge}(u,v)$ we denote the  length of a shortest path between $u$ and $v$ (belonging to $G^{\wedge\vee}$) where all edges are conjunctive edges. We can then define $N^{\wedge}_r(u) = \lbrace v\ |\ \mathcal{D}^{\wedge}(u,v) \leq r \rbrace$
and the {\it conjunctive neighborhood subgraph} $\mathcal{N}_{r}^{\wedge u}$ as the  subgraph induced by $N^{\wedge}_r(u)$ only considering conjunctive edges. We can now define 
two relations: the \textit{conjunctive relation} $R^{\wedge}_{r,d,u}(A_w, B_v, G^{\wedge\vee})$, which is true 
{\em iff}  $w = u$ and $\mathcal{D}^{\wedge}(w,v)= d$ and $A_w \cong \mathcal{N}_r^{\wedge w}$ is (up to isomorphism $\cong$) a conjunctive neighborhood subgraph of radius $r$ of $G^{\wedge\vee}$ as well as $B_v \cong  \mathcal{N}_r^{\wedge v}$;
the \textit{disjunctive relation} $R_{r,d,u}^{\vee}(A_w, B_v, G^{\wedge\vee})$ which is true {\em iff} $w = u$ and $A_w \cong \mathcal{N}_r^{\wedge w}$ and $B_v \cong \mathcal{N}_r^{\wedge v}$ are true and $\exists z$ s.t. $\mathcal{D}^{\wedge}(z,v)= d \wedge (w,z)$ is a disjunctive edge.

We define $\kappa^{\wedge\vee}_{r,d}$ on the  inverse relations ${R^{\wedge -1}_{r,d,u}}$ and ${R^{\vee -1}_{r,d,u}}$:
\begin{center}
 $\kappa^{\wedge\vee}_{r,d}(u,u^\prime) = \!\!\!\!\!\!\!\!\!\!\!\!
 \sum\limits_{\substack {A_u,{B}_{v} \in {R_{r,d,u}^{\wedge -1}}(G^{\wedge\vee}) \\ A'_{u^\prime},{B'}_{v'} \in {R_{r,d,u^\prime}^{\wedge -1}}(G^{\wedge\vee}) }} \!\!\!\!\!\!\!\!\!\!\!\!
  { \textbf{1}_{A_u \cong A'_{u^\prime}} \cdot { \textbf{1}_{B_{v} \cong B'_{v'}}}}
+ \!\!\!\!\!\!\!\!\!\!\!\!
 \sum\limits_{\substack {A_u,{B}_{v} \in {R_{r,d,u}^{\vee -1}}(G^{\wedge\vee}) \\
  A'_{u^\prime},{B'}_{v'} \in \ {R_{r,d,u^\prime}^{\vee -1}}(G^{\wedge\vee}) }} \!\!\!\!\!\!\!\!\!\!\!\!
  { \textbf{1}_{A_u \cong A'_{u^\prime}} \cdot { \textbf{1}_{B_{v} \cong B'_{v'}}}}
  $.
\end{center}
The CDNK is finally defined as $K(u,u^\prime) = \sum\limits_{r}{\sum\limits_{d}{\kappa^{\wedge\vee}_{r,d}(u,u^\prime)}}$, where once again for efficiency reasons, the values of $r$ and $d$ are upper bounded to a given maximal $r^*$ and $d^*$.

\begin{figure}
\centering
\includegraphics[width=.9\textwidth]{cdnk}
\caption{\label{fig:cdnk} Examples of two pairwise neighborhood subgraphs for the $u$ node with $r=1$ and $d=3$ using \textit{conjuctive} and \textit{disjuctive} edges. {\it Left:} input graph where nodes $v$ and $v^\prime$ have a {\it conjunctive} shortest path (highlighted in red) of length $3$ with $u$;  {\it conjunctive} neighborhood subgraphs of radius $1$ for $u$ (highlighted in green), $v$ (highlighted in violet), $v^\prime$ (highlighted in yellow)  are shown as well. {\it Right:} representation of the two induced pairwise \textit{conjuctive} neighborhood subgraphs.} 
\end{figure}

\subsection{Additional Node Information}
In order to integrate the information of real vector labels that may be present associated to nodes, we proceed as follows. We compute a sparse vector representation for the conjunctive neighborhood graph rooted at node $u$ following \cite{costa2010fast}: for each conjunctive neighborhood subgraph we calculate the quasi-isomorphism certificate hash code; we then combine the hashes for the pair of neighborhoods and use the resulting integer as a feature indicator. This yields a direct sparse vector representation (associated to node $u$ in graph $G^{\wedge\vee}$) $f: u \longmapsto \mathbb{R}^N$ where $N \approx 10^6$. Given the real valued vector information associated to nodes in graph $G^{\wedge\vee}$, we also cluster\footnote{Any clustering algorithm that returns cluster representations provided of a function to measure the similarity of a vector with respect to a given cluster will do the job.} them in $P$ clusters (c.f.r. Section \ref{sec:lab}) and define a function $g: u \longmapsto \mathbb{R}^P$ which returns the multi-class similarity w.r.t. each of the $P$ cluster (prototypes).

A new version of CDNK considering the discrete convolution of the discrete information with the real valued information can now be computed as:   
\begin{equation}
K(u,u^\prime) = \left< f(u) \bigotimes g(u), f(u^\prime) \bigotimes g(u^\prime) \right>,
\end{equation}
where the discrete convolution is defined as: 
$(f \bigotimes g)[n]=\sum _{m=0}^{K-1}f[n-m]g[m])$.
In words, we place a scaled copy of the real valued vector at the position indicated by each feature computed on the basis of the discrete information. Intuitively, when both the real valued and the discrete information match, the kernel computes a large similarity, but if there is a discrepancy in either one of the sources of information, the similarity will be penalized.

\subsection{Hyper-parameter Space}\label{hyper}
Our kernel consists of five hyper-parameters: the threshold degree $D$, the clique size threshold $C$, the maximal radius $r^*$, the maximal distance $d^*$, and the number of clusters $P$. In order to choose the optimal tuple of   values for kernel hyper-parameters in a specific setting, a model selection procedure is normally adopted. The hyper-parameter space of our kernel is potentially large due to the relatively high number of hyper-parameters. However, all hyper-parameters are integer variables and it makes sense to use only small values for them. For example, the value of the threshold degree $D$ should neither be too high nor too low: if it is high, the neighborhood subgraphs are going to be too big and the isomorphism check becomes inefficient; vice versa, if it is too low, the decomposed graph contains connected components which are too sparse to allow learning to be effective. Moreover, it is reasonable to keep the clique size threshold $C$ 
within a narrow range of values in order to avoid the removal of too many edges. Concerning the maximal radius $r^*$,  too  high values will lead again to big neighborhood subgraphs, especially in graphs with a high node degree average. The maximal distance $d^*$ can be set with small values since the average shortest distance length between nodes of biological networks is relatively low. Finally, if the value assigned to $P$ is too low, all nodes basically get the same information attached to them and thus the labels are not informative; vice versa, if a high value is used, there will be very few matches and the kernel will be extremely sparse, making learning ineffective.

%Our kernel consists of five parameters: the threshold degree $D$, clique size threshold $C$, maximal radius $r^*$, maximal distance $d^*$ and number of clusters $P$ clusters in which their values are in $\mathbb{N}^*$. In order to choose the optimal tuple of parameters for kernel in a specific setting, a model selection procedure is normally adopted from a determined subset of parameter space. The parameter space of our kernel seems large due to the relatively high number of parameters. However, most parameter values are supposed to be in limited ranges. Therefore, it is actually not too big. In particular, the clique size threshold $C$ is recommended to be in $\lbrace 4, 5\rbrace$. The maximal radius $r^*$ is set with a value smaller than or equal to $3$ since it will lead to big neighborhood subgraphs (as discussed in \ref{network-decomposition}). The maximal distance $d^*$ is assigned with values less than or equal to $4$ because the value of the maximal shortest distance between nodes in a connected component is often not too high. The values of the threshold degree $D$ should neither be too high and too low. If it is high, we have to face with high degree node problem, and if it is too low, the obtained graph consists of too sparse connected components. Therefore, we suggest the value for $D$ is in $\left[6, 20 \right]$.

\section{Empirical Evaluation}
\label{empirical-evaluation-cdnk}
In this section, we evaluate the performance of CDNK versus other graph node kernels. Specifically, we focus on the following two questions:
\begin{itemize}
\item {\it Q1}: Does CDNK show better performance with respect to other graph node kernels?
\item {\it Q2}: Does the use of local topology (as envisioned in the proposed kernel) and/or additional information (real vector labels) help to  improve the performance of CDNK?
\end{itemize}

\subsection{Experimental Settings and Evaluation Methods}
We carry out experiments in the context of disease gene prioritization where the aim is to build a learning system which allows to prioritize candidate genes based on their probabilities of being associated to a given genetic disease.

The learning pipeline typically used for this kind of tasks is described in the following. In input the system takes: a genetic graph, a set of training genes known to cause a specific genetic disease, a set of candidate genes which are suspected to be involved as well in the disease. The selected graph node kernel is applied to the genetic graph, so to obtain the corresponding kernel matrix. The kernel matrix and the training genes are given in input to a kernel machine to construct a model.  The output of the obtained model for the candidate genes is then used to prioritize them. The top ranked genes are the ones supposed to have the highest likelihood of concurring in causing the disease.

The experiments are performed on two separate networks derived from the {\it BioGPS} and the {\it Pathways} datasets:
\begin{itemize}
\item \textit{BioGPS:} A gene co-expression network involving 79 tissues, measured with the Affymetrix U133A array. Edges are inserted when the pairwise Pearson correlation coefficient (PCC) between genes is larger than 0.5.

\item \textit{Pathways:} Pathway information is retrieved from KEGG, Reactome, PharmGKB and the Pathway Interaction Database. If a couple of proteins co-participate in any pathway, the two corresponding genes are linked.  
\end{itemize}

In Figure~\ref{fig:histograms} we have plotted some statistics on the distribution of the outdegree of nodes in the two networks. It can be observed that there are many nodes with a very high outdegree, especially for the Pathways network, so the decomposition procedure is really important to drastically reduce it without loosing too much information about the network topology.

\begin{figure}[h]
\centering
\subfloat[Subfigure 2 list of figures text][]{
\includegraphics[width=0.5\textwidth]{degree_histogram_biogps.pdf}
\label{fig:subfig1}}
\subfloat[Subfigure 3 list of figures text][]{
\includegraphics[width=0.5\textwidth]{degree_histogram_pathways.pdf}
\label{fig:subfig2}}
\caption{\label{fig:histograms} Outdegree distribution for nodes in the BioGPS network (a), and the Pathways network (b). The number associated to the bin is the threshold value used for the outdegree.}
\end{figure}

We followed the experimental procedure described in \cite{chen2014disease}, where $12$ diseases \cite{goh2007human} are used. Each disease is associated to at least 30 positive genes. See Table~\ref{tab:diseases} for the list of genetic diseases and the number of positive genes in each disease. For each disease, we construct a positive set $\mathcal{P}$ with all positive (confirmed) disease genes, and a negative set $\mathcal{N}$ which contains random genes associated at least to one disease class which is not related to the class that is defining the positive set. In \cite{chen2014disease} the ratio between the dataset sizes is chosen as $\vert \mathcal{N} \vert = \frac{1}{2} \vert \mathcal{P} \vert$. This is due to the fact that genes known to be related to at least one genetic disease, but not to the considered one, are well studied, so they have a low probability of being associated to the current disease.

%{\renewcommand{\arraystretch}{1.3}
%\vspace{-1.5cm}
\begin{table}
%\vspace*{1cm}
\small
\centering
\caption{\label{tab:diseases} Indices and gene number of genetic disease-gene associations.} \vspace{1em}
\label{tab:cdnk-diseases}
\setlength{\tabcolsep}{1cm}
\begin{tabular}{ccc}
\hline
Index & Disease & Number\\
\hline
0 & Connective & 35 \\
1 & Cardiovascular & 75 \\
2 & Dermatological & 54 \\
3 & Developmental & 37 \\
4 & Endocrine & 62 \\
5 & Hematological & 106 \\
6 & Immunological & 94 \\
7 & Metabolic & 159 \\
8 & Muscular & 55 \\
9 & Ophthalmological & 61 \\
10 & Renal & 42 \\
11 & Skeletal & 35 \\
\hline
\end{tabular}
\end{table}%}

In order to compare the performance of different graph node kernels, i.e. LEDK, MEDK, MDK, RLK, and CDNK., the learning algorithm is fixed. 
The predictive performance of the system using each kernel is evaluated via a leave-one-out cross validation: one gene is kept out in turn and the rest are used to train an SVM model. 
We computed a decision score $q_i$ for the test gene $g_i$ as the top percentage value of score $s_i$ among all candidate gene scores, i.e. $q_i = \frac{ |\lbrace g_j | s_i \geq s_j\rbrace |}{N}$ where $N$ is the number of candidate genes. We collected all decision scores for every gene in the test set to form a global decision score list on which we computed the AUC-ROC.

\subsection{Node Labeling} 
\label{sec:lab}
The proposed graph node kernel is able to exploit labels attached to nodes.  Labels can be discrete and/or real valued vectors. Here we discuss how to define these labels for the graphs derived from BioGPS and Pathways in such a way to enrich the amount of information that the kernel can exploit to perform its task.
Notice that the definition of the kernel allows the user to define and use any node labeling function, not only the ones we describe in the following.

\textit{Discrete labels:} we designed two different approaches to associate genes with discrete labels.

A first baseline approach is to use the same label for every node of the graph. In this case, the underpinning information for the computation of the kernel is just the topology of the neighborhood subgraphs. Topology is the information that is used by the diffusion-based kernels. Thus, this choice allows us to test whether putting the focus of attention on local topological information, as encoded by the proposed kernel, is useful or not.

The second approach aims to encode abstract information about genes into the labels. The idea, in this case, is to allow  downstream machine learning algorithms to generalize from similar examples and to allow the identification of overlooked but related genes. To this aim, we employ a gene ontology (GO) \cite{gene2004gene} to construct binary vectors representing a bag-of-terms encoding for each gene, i.e. an element of the binary vector is equal to 1 if its corresponding GO-term is associated to the gene, and is equal to 0 otherwise. The resulting vectors are then clustered using the k-means algorithm into a user defined number of classes, $P$, so that genes with similar description profiles receive the same class identifier as label.

\textit{Real vector labels:} In addition to encoding the functional information as a discrete label, we add a richer description by computing the similarity vector w.r.t. to each cluster. In this way, we can fully exploit the latent description of the genes in terms of the different functional groups captured by the clustering procedure. Formally, given a vector $v \in \mathbb{R}^{26501}$ we compute a similarity vector $S(v)= {s_1, s_2, \ldots, s_P}$ with entries $s_i = \frac{1}{1+ \ell(v,c_i)}$ where $\ell(v,c_i)$ is the Euclidean distance of $v$ from the center of the $i^{th}$ cluster $c_i = \frac{1}{|C_i|}\sum_{x \in C_i} x$ computed as the geometric mean of the elements in the cluster $C_i$.

\subsection{Model Selection}
The hyper-parameters of the various methods were tuned using a 10-fold cross validation. However, due to the non i.i.d. nature of the problem, we emploied a stronger setup to ensure no information leakage. The dataset $0$ (Connective) on which we  validated the performance was never subsequently used in the predictive performance estimation. The values for the diffusion parameter $\beta$ in LEDK and MEDK were sampled in $\lbrace 10^{-3}, 10^{-3}, 10^{-2}, 10^{-1} \rbrace$, time steps $t$ in MDK in $\lbrace 1, 10, 100 \rbrace$ and RLK parameter $\beta$ in $\lbrace 1, 4, 7 \rbrace$. For CDNK, the degree threshold values, $D$, were sampled in $\lbrace 10,\ 15,\ 20 \rbrace$, clique size threshold, $C$, in $\lbrace 4,\ 5 \rbrace$, maximum radius, $r^*$, in $\lbrace 1,\ 2 \rbrace$, maximum distance, $d^*$, in $\lbrace 2,\ 3,\ 4 \rbrace$, number of clusters $P$ in $\lbrace 5,\ 7 \rbrace$. Finally, the regularization trade off parameter in the SVM was sampled in $\lbrace 10^{-5},  \ 10^{-4}, \ 10^{-3},\ 10^{-2},\ 10^{-1}, 1,\ 10,\ 10^2 \rbrace$.
%\subsection{Parameter Impact}
%In order to evaluate the stability of kernel w.r.t the change of each parameter values, we perform experiments with disease-gene association 2 on the BioGPS network. For each parameter, first we fix the values of all other parameters, but allow the value of considered papameter to be changed. More precisely, we use the optimal tuple $C = 5,\ d^* = 1,\ D = 20,\ P = 15$ for the fixed of parameters' values. We let the changes of $C$ in $\lbrace 4,\ 5,\ 6,\ 7,\ 8,\ 9\rbrace$, $d^*$ in $\lbrace 1,\ 2,\ 3,\ 4\rbrace$, $D$ in $\lbrace 8,\ 10,\  15,\ 20,\ 30,\ 40,\ 50\rbrace$ and real vector size, $P$ in $\lbrace 8,\ 10,\ 15,\ 20,\ 30,\ 50,\ 70, 90\rbrace$.

\subsection{Results and Discussion}
%==================================================
%==================================================
%==================================================
\begin{table}[h]
\caption{Predictive performance on 11 gene-disease associations in percentage using network induced by the BioGPS. Best results in bold. We report the AUC ROC and after the ``/'' the rank for each kernel method (DK = LEDK; MD = MDK; MED = MEDK; RL = RLK; CDNK1 = ontology for discrete labels; CDNK2 = ontology for both discrete and vector labels; CDNK3 = uniform discrete labels;  CDNK4 = uniform discrete labels and ontology for vector labels.) The last two rows report the averages of AUC ROC and rank over the $11$ diseases.} \vspace{1em}
%\vspace*{5pt}
\centering
\setlength{\tabcolsep}{1mm}
\begin{tabular}{|c|c|c|c|c|c|c|c|c|}
\hline
         & \multicolumn{8}{c|}{\textbf{BioGPS}}\\
 \hline
Disease & DK & MD & MED & RL & CDNK1 & CDNK2 & CDNK3 & CDNK4\\

 \hline
1 & 51.9/8 & 57.4/7 & 59.0/6 & 59.2/5 & 65.1/4 & 69.5/3 & 72.0/2 & \textbf{72.3/1} \\

2 & 81.7/5 & 78.5/6 & 75.2/7 & 75.0/8 & 88.3/2 & \textbf{88.8/1} & 83.2/4 & 84.8/3 \\

3 & 64.3/6 & 59.6/8 & 71.6/3 & 71.8/2 & 65.5/5 & \textbf{72.5/1} & 61.4/7 & 67.9/4 \\

4 & 65.3/7 & 58.2/8 & 67.8/6 & 67.8/5 & 71.9/4 & \textbf{78.7/1} & 73.8/3 & 76.7/2 \\

5 & 64.0/8 & 64.1/7 & 66.5/5 & 66.2/6 & 75.9/4 & 76.2/3 & 77.8/2 & \textbf{78.2/1} \\

6 & 74.6/5 & 70.2/8 & 71.0/7 & 71.2/6 & 79.3/3 & \textbf{83.7/1} & 77.2/4 & 80.0/2 \\

7 & 73.0/5 & 66.7/7 & 75.4/3 & 75.6/2 & 68.8/6 & 73.9/4 & 66.1/8 & \textbf{77.4/1} \\

8 & 74.4/8 & 76.8/3 & 76.2/5 & 76.4/4 & 74.7/7 & 77.7/1 & 76.1/6 & 76.9/2 \\

9 & 71.5/2 & 65.6/8 & 67.7/6 & 69.9/3 & 66.8/7 & \textbf{71.7/1} & 67.9/5 & 68.1/4 \\

10 & 54.0/6 & 50.3/8 & 56.1/5 & 51.1/7 & 77.6/4 & \textbf{82.7/1} & 77.7/3 & 78.3/2 \\

11 & 58.2/7 & 51.3/8 & 59.3/6 & 59.3/5 & 71.8/4 & \textbf{80.2/1} & 74.8/2 & 74.0/3 \\

\hline
$\overline{AUC}$ & 66.6	& 63.5 & 67.8 &	67.6 & 73.2 & \textbf{77.8} & 73.5 & 75.9 \\

$\overline{Rank}$ & 6.09 & 7.09 &  5.36 & 4.82 & 4.45 & \textbf{1.64} & 4.18 & 2.27 \\
 \hline 
\end{tabular}
\label{tab:cdnk-biogps}
\end{table}
%=====================================
%=====================================
\begin{table}[h]
\caption{Predictive performance on 11 gene-disease associations in percentage using network induced by the Pathways. Best results in bold. We report the AUC ROC and after the ``/'' the rank for each kernel method (DK = LEDK; MD = MDK; MED = MEDK; RL = RLK; CDNK1 = ontology for discrete labels; CDNK2 = ontology for both discrete and vector labels; CDNK3 = uniform discrete labels;  CDNK4 = uniform discrete labels and ontology for vector labels.) The last two rows report the averages of AUC ROC and rank over the $11$ diseases.} \vspace{1em}
%\vspace*{5pt}
\centering
\setlength{\tabcolsep}{1mm}
\begin{tabular}{|c|c|c|c|c|c|c|c|c|}
\hline
         & \multicolumn{8}{c|}{\textbf{Pathways}} \\
 \hline
Disease & DK & MD & MED & RL & CDNK1 & CDNK2 & CDNK3 & CDNK4 \\

 \hline
1 & 74.7/8 & 76.4/7 & 78.7/6 & 78.8/5 & 80.2/4 & 82.4/2 & \textbf{82.8/1} & 82.3/3 \\

2 & 55.1/8 & 64.9/7 & 76.6/6 & 76.6/5 & 81.1/2 & 80.3/4 & 80.1/3 & \textbf{81.9/1} \\

3 & 55.0/8 & 62.7/7 & 64.1/5 & 65.6/4 & 67.1/3 & 63.6/6 & 69.7/2 & \textbf{71.4/1} \\

4 & 54.3/8 & 65.2/6 & \textbf{73.7/1} & 73.7/2 & 66.1/5 & 68.1/3 & 64.0/7 & 67.3/4 \\

5 & 52.9/8 & 55.7/7 & 62.7/5 & 62.7/6 & 68.3/2 & \textbf{69.6/1} & 66.2/4 & 68.1/3 \\

6 & 83.4/8 & 92.7/7 & 96.5/2 & \textbf{96.5/1} & 93.0/6 & 94.1/5 & 94.5/4 & 94.9/3 \\

7 & 84.5/7 & 88.3/8 & 89.4/2 & \textbf{89.5/1} & 88.5/6 & 88.5/5 & 89.0/3 & 88.7/4 \\

8 & 53.7/8 & 65.6/7 & 72.0/6 & 72.3/5 & 72.5/4 & 72.5/3 & 74.9/2 & \textbf{75.3/1} \\

9 & 52.5/8 & 64.9/5 & 64.2/7 & 64.2/6 & \textbf{81.3/1} & 81.0/3 & 80.1/2 & 79.8/4 \\

10 & 68.8/6 & 65.4/8 & 74.4/5 & 74.4/4 & 66.9/7 & 76.8/2 & 75.2/3 & \textbf{78.7/1} \\

11 & 53.7/8 & 69.2/7 & 74.6/5 & 74.1/6 & 77.0/3 & \textbf{78.7/1} & 75.1/4 & 77.3/2 \\

\hline
$\overline{AUC}$ & 62.6	& 70.1 & 75.2 & 75.3 & 76.5 & 77.8 & 77.4 &	\textbf{78.7}
 \\
$\overline{Rank}$ & 7.73 & 6.82 & 4.55 & 4.09 & 3.91 & \textbf{2.27} & 3.18 & 2.45 \\
 \hline 
\end{tabular}
\label{tab:cdnk-pathways}
\end{table}
Tables \ref{tab:cdnk-biogps} and \ref{tab:cdnk-pathways} report the AUC-ROC performance of the different diffusion-based graph node kernels and the different variations of CDNK on BioGPS and Pathways networks, respectively. In what follows, we analyze the results to answer to the two research questions we posed at the beginning of the section.

Concerning {\it Q1}, i.e. improvement on the performance of diffusion-based graph node kernels, CDNK variations outperformed diffusion-based graph node kernels in almost all cases on BioGPS network. For the moment, we postpone the discuss of these cases. When looking at the Pathways network, CDNK variations were outperformed by one of the diffusion-based graph node kernels in only $3$ out of $11$ cases. 
When looking at the average results, it is clear that
CDNK variations are ranked first when considering both the average AUC-ROC and the average rank with a difference, w.r.t. the best diffusion-based kernel, ranging from $5.4\%$ to $10\%$ and from $1.2\%$ to $3.4\%$ on BioGPS and Pathways, respectively. As a consequence, we can conclude that CDNK variations show state-of-the art performance in graph node kernels for this task. Thus, the answer to question {\it Q1} is positive.

Let now turn our attention to question {\it Q2}, i.e. whether the possibility to use local topology and/or real valued information is useful to improve performance.
The first part of the question, i.e. usefulness of local topology, can be answered by looking at results of CDNK3, where only discrete labels defined by the baseline approach are used. If we look at results obtained on BioGPS network, we see that CDNK3 outperforms diffusion-based kernels with the 
 exceptions of datasets $3$, $7$, $8$, and $9$, thus $7$ times over $11$. It is important to stress, however, that on the average CDNK3 outperformed the diffusion-based kernels of $10.19\%$, while it was outperformed on the average of $6.05\%$. An even better scenario is observed for the Pathways network, where only in one case (dataset $4$) the performance of CDNK3 was inferior of almost all diffusion-based kernels.
Thus, it seems that exploiting local topological information is actually important in many cases. 
Regarding the integration of real valued vectors, we have to look at variations CDNK2 and CDNK4, that exploit real valued labels based on GO.
It is readily evident that CDNK2 and CDNK4 improve on the performance of diffusion-based kernels in most cases. Specifically, for BioGPS network,
CDNK2 is the best performer in absolute in $7$ out of $11$ datasets. Moreover, its worst performance places it above all diffusion-based kernels.
In turn, CDNK4 is first in $3$ datasets and second in $4$ datasets, and again its worst performance places it above all diffusion-based kernels.
When looking at the results for the Pathways network, CDNK4 is the best performer among all kernels, while CNDK2 places in second position since
even if it is first on two datasets as well as RLK, RLK has a worst ranking score on most of the other datasets.
We can conclude that also the exploitation of real valued information seems to improve the performance. Thus, also for question  {\it Q2} we can give a positive answer.
%In particular, considering the use of discrete labels based on ontology, using side information helps to increase the performance in all diseases on BioGPS and in 9 out of 11 diseases on Pathways. Similarly, by employing auxiliary information associated to graph nodes, the performance of CNDK in case of using uniform discrete labels is also improved in 9 and 8 out of 11 on BioGPS and Pathways, respectively.

\subsection{Sensitivity Analysis for Hyper-parameters}
In order to support our claims in Section~\ref{hyper}, we performed the following experiment with kernel CDNK4. For the BioGPS network we selected disease 2 (Dermatological)
and, starting from the configuration of the values for the hyper-parameters selected by the model selection procedure, we in turn moved the value of each hyper-parameter, monitoring the change in performance. We left unchanged the maximum radius ($r^*$) for computational reasons: having too big neighborhood subgraphs makes computation of isomorphism too heavy and significantly reduces the likelihood of finding matchings. The obtained results are shown in Figure~\ref{fig:globfig}.
It can be observed that the range of values where the performance is maximal is located around small values for the hyper-parameters, thus confirming that too low or too high values for the hyper-parameters do not bring good performances.

%==================================================
\begin{figure}[h]
\centering
\subfloat[Subfigure 2 list of figures text][]{
\includegraphics[width=0.4\textwidth]{cliques.png}
\label{fig:subfig1}}
\subfloat[Subfigure 3 list of figures text][]{
\includegraphics[width=0.4\textwidth]{distance.png}
\label{fig:subfig2}}
\qquad
\subfloat[Subfigure 2 list of figures text][]{
\includegraphics[width=0.4\textwidth]{degree.png}
\label{fig:subfig3}}
\subfloat[Subfigure 3 list of figures text][]{
\includegraphics[width=0.4\textwidth]{vec_size.png}
\label{fig:subfig4}}
\caption{\label{fig:globfig} Performance of CDNK4 on the BioGPS network for disease 2 (Dermatological) w.r.t the change of most of the hyper-parameters: (a) clique threshold ($C$); (b) distance threshold ($d^*$); (c) degree threshold ($D$); (d) real vec size threshold ($P$). The radius ($r^*$) was not changed because of computational reasons.}
\end{figure}

\section{Conclusion and Future Work}
We have shown how decomposing a network in a set of connected sparse graphs allows us to take advantage of the discriminative power of CDNK, a novel decomposition kernel, to achieve state-of-the-art results for a disease gene prioritization task.
Moreover, we have also introduced the possibility to integrate in the kernel computation  additional information in the form of real valued vectors in the case in which it is available for graph nodes. We experimentally showed that this possibility actually leads to even better performances. Future work we will investigate on how to: \textit{i}) decompose networks in a data driven way; \textit{ii}) extend the CDNK approach to gene-disease association problems exploiting multiple heterogeneous information sources in a joint way.

\section*{Funding}
This work was supported by the University of Padova, Strategic Project BIOINFOGEN.
\section*{References}

\bibliography{mybibfile}
\end{document}

### 2017-11-16 Response_neurocomputing.tex
update response letter

@@
 
-\underline{Answer}: We would like to compare our node kernel diffusion-based graph node kernels on relatively dense networks, since diffusion-based kernels often show good performance with dense networks.
+\underline{Answer}: We have added an evaluation of the sensitivity of the proposed kernel to some parameters in section ``Sensitivity Analysis for Hyper-parameters" of the new version.
+
+In the evaluation, we would like to compare node kernel diffusion-based graph node kernels on relatively dense networks, since diffusion-based kernels often show good performance with dense networks.
 That is why we borrow BioGPS and Pathways which are used in [2].

### 2017-11-16 Response_neurocomputing.tex
F

@@
 
-\underline{Answer}: The existance of high degree nodes will lead to the big neighborhood subgraphs.
+\underline{Answer}: The existence of high degree nodes will lead to big neighborhood subgraphs.
 Since the proposed kernel is based on the exact matching neighborhood subgraphs (isomorphic neighborhood subgraphs), there will be a low probability to find any couple of big neighborhood subgraphs which are isomorphic.
-This probabability decrease shapely w.r.t the increase of neighborhood subgraphs size or node degree.
+This probability decrease sharply w.r.t the increase of neighborhood subgraphs size or node degree as a result of the combinatorial increase in the number of possible labeled graphs w.r.t.
+the graph size.
 
@@
 
-In general, our kernel takes labeled graphs as the input.
-Given a graph, users can apply any proper node labeling function to the graph.
-The labeling functions proposed in the paper are examples to evaluate the performance of the disease gene prioritization system with the use of our kernel.
-Regarding the proposed labeling functions, we consider topological labels as discrete labels.
-The GO datasets do not provide complete information.
-However, they still provide additional information for the kernel.
-We strongly agree that the intensitive evaluation for labeling functions is necessary to improve the performance of the proposed kernel.
-However, in this paper, we would like to have an attempt to see how the CDNK performs compared to other graph node kernels.
+The proposed kernel takes labeled graphs in input.
+The appropriate node labeling function is domain dependent and can be freely defined by the user.
+In this paper we propose a couple of labeling functions to evaluate the performance of the disease gene prioritization system with the use of our kernel.
+We consider a trivial constant label to show the capacity of the approach to exploit topological information only.
+In addition we consider a domain dependent source of information derived by the GO hierarchy.
+We agree that a more extensive evaluation for labeling functions would improve the performance on a specific task.
+The scope of the evaluation presented in this paper is however only to show that CDNK can in principle improve results by exploiting domain knowledge.
 
@@
 
-In disease prioritization, given a genetic disease, we are often given a small set of positive genes and a large set of un-labeled genes.
+In disease prioritization, given a genetic disease, we are often given a small set of positive genes and a large set of unlabeled genes.
 It is difficult to have a proper negative set of genes.
-Therefore, in [2] they choose the number of postive genes two times greater than the number of negative ones.
-The negative genes are choosen from genes that are associated with at least one disease class, but are not related to the class that is defining the positive set.
+Therefore, in [2] they choose the number of positive genes two times greater than the number of negative ones.
+The negative genes are chosen from genes that are associated with at least one disease class, but are not related to the class that is defining the positive set.
 We assume that they are less likely to be associated to the considered class, since disease genes are generally more studied and a potential association has more chances to be identified.

### 2017-11-16 neurocomputingv3.tex

@@
 In this way, we can fully exploit the latent description of the genes in terms of the different functional groups captured by the clustering procedure.
-Formally, given a vector $v \in \mathbb{R}^{26501}$ we compute a similarity vector $S(v)= {s_1, s_2, \ldots, s_P}$ with entries $s_i = \frac{1}{1+ \ell(v,c_i)}$ where $\ell(v,c_i)$ is the Euclidean distance of $v$ from the center of the $i^{th}$ cluster $c_i = \frac{1}{|C_i|}\sum_{x \in C_i} x$ computed as the geometric mean of the elements in the cluster $C_i$.
+Formally, given a vector $v \in \mathbb{R}^{M}$ \footnote{M is the number of GO-terms, M = 26,501} we compute a similarity vector $S(v)= {s_1, s_2, \ldots, s_P}$ with entries $s_i = \frac{1}{1+ \ell(v,c_i)}$ where $\ell(v,c_i)$ is the Euclidean distance of $v$ from the center of the $i^{th}$ cluster $c_i = \frac{1}{|C_i|}\sum_{x \in C_i} x$ computed as the geometric mean of the elements in the cluster $C_i$.
 
@@
 The values for the diffusion parameter $\beta$ in LEDK and MEDK were sampled in $\lbrace 10^{-3}, 10^{-3}, 10^{-2}, 10^{-1} \rbrace$, time steps $t$ in MDK in $\lbrace 1, 10, 100 \rbrace$ and RLK parameter $\beta$ in $\lbrace 1, 4, 7 \rbrace$.
-For CDNK, the degree threshold values, $D$, were sampled in $\lbrace 10,\ 15,\ 20 \rbrace$, clique size threshold, $C$, in $\lbrace 4,\ 5 \rbrace$, maximum radius, $r^*$, in $\lbrace 1,\ 2 \rbrace$, maximum distance, $d^*$, in $\lbrace 2,\ 3,\ 4 \rbrace$, number of clusters $P$ in $\lbrace 5,\ 7 \rbrace$.
+For CDNK, the degree threshold values, $D$, were sampled in $\lbrace 10,\ 15,\ 20 \rbrace$, clique size threshold, $C$, in $\lbrace 4,\ 5 \rbrace$, maximum radius, $r^*$, in $\lbrace 1,\ 2 \rbrace$, maximum distance, $d^*$, in $\lbrace 1, 2,\ 3,\ 4 \rbrace$, number of clusters $P$ in $\lbrace 5,\ 7 \rbrace$.
 Finally, the regularization trade off parameter in the SVM was sampled in $\lbrace 10^{-5},  \ 10^{-4}, \ 10^{-3},\ 10^{-2},\ 10^{-1}, 1,\ 10,\ 10^2 \rbrace$.
+The optimal parameter values of CDNK obtained by validation process are $D = 1$, $C=5$, $r^*=2$, $d^*=1$, $P=15$ and the trade-off parameter of SVM is $10^{-2}$.
 %\subsection{Parameter Impact}

### 2017-11-16 neurocomputingv3.tex
corrected file

@@
 Finally, the regularization trade off parameter in the SVM was sampled in $\lbrace 10^{-5},  \ 10^{-4}, \ 10^{-3},\ 10^{-2},\ 10^{-1}, 1,\ 10,\ 10^2 \rbrace$.
-The optimal parameter values of CDNK obtained by validation process are $D = 1$, $C=5$, $r^*=2$, $d^*=1$, $P=15$ and the trade-off parameter of SVM is $10^{-2}$.
+The optimal parameter values of CDNK obtained by validation process are $D = 20$, $C=5$, $r^*=2$, $d^*=1$, $P=15$ and the trade-off parameter of SVM is $10^{-2}$.
 %\subsection{Parameter Impact}
@@
 It is important to stress, however, that on the average CDNK3 outperformed the diffusion-based kernels of $10.19\%$, while it was outperformed on the average of $6.05\%$.
-An even better scenario is observed for the Pathways network, where only in one case (dataset $4$) the performance of CDNK3 was inferior of almost all diffusion-based kernels.
+An even better scenario is observed for the Pathways network, where only in one case (disease gene association $4$) the performance of CDNK3 was inferior of almost all diffusion-based kernels.
 Thus, it seems that exploiting local topological information is actually important in many cases. 

### 2017-11-16 Response_neurocomputing.tex
F

@@
 
-\underline{Answer}: we have added in the new version of the paper.
+\underline{Answer}: we have added the discussion in the new version of the paper.
 
@@
 
-\underline{Answer}: We have added a new section named ``Hyper-parameter" to explain why the parameter space for the proposed kernel is not high as follows.\\
+\underline{Answer}: We have added the following section named ``Hyper-parameter" to explain why the parameter space for the proposed kernel is not high.\\
 

### 2017-11-16 neurocomputingv3.tex
F

--- 2017-11-16 neurocomputingv3.tex
@@
 In order to integrate the information of real vector labels that may be present associated to nodes, we proceed as follows.
-We compute a sparse vector representation for the conjunctive neighborhood graph rooted at node $u$ following \cite{costa2010fast}: for each conjunctive neighborhood subgraph we calculate the quasi-isomorphism certificate hash code; we then combine the hashes for the pair of neighborhoods and use the resulting integer as a feature indicator.
+We compute a sparse vector representation for the conjunctive neighborhood graph rooted at node $u$ following \cite{costa2010fast}: for each conjunctive neighborhood subgraph we calculate a quasi-isomorphism certificate hash code\footnote{The graph hashing is obtained by  hashing the sorted sequence of edge extended information which includes an invariant encoding of the endpoint labels (see \cite{costa2010fast}).}; we then combine the hashes for the pair of neighborhoods and use the resulting integer as a feature indicator.
 This yields a direct sparse vector representation (associated to node $u$ in graph $G^{\wedge\vee}$) $f: u \longmapsto \mathbb{R}^N$ where $N \approx 10^6$.
@@
 In this way, we can fully exploit the latent description of the genes in terms of the different functional groups captured by the clustering procedure.
-Formally, given a vector $v \in \mathbb{R}^{M}$ \footnote{M is the number of GO-terms, M = 26,501} we compute a similarity vector $S(v)= {s_1, s_2, \ldots, s_P}$ with entries $s_i = \frac{1}{1+ \ell(v,c_i)}$ where $\ell(v,c_i)$ is the Euclidean distance of $v$ from the center of the $i^{th}$ cluster $c_i = \frac{1}{|C_i|}\sum_{x \in C_i} x$ computed as the geometric mean of the elements in the cluster $C_i$.
+Formally, given a vector $v \in \mathbb{R}^{M}$ \footnote{M is the number of GO-terms, in our experiments M = 26,501.} we compute a similarity vector $S(v)= {s_1, s_2, \ldots, s_P}$ with entries $s_i = \frac{1}{1+ \ell(v,c_i)}$ where $\ell(v,c_i)$ is the Euclidean distance of $v$ from the center of the $i^{th}$ cluster $c_i = \frac{1}{|C_i|}\sum_{x \in C_i} x$ computed as the geometric mean of the elements in the cluster $C_i$.
+While other notions of similarity can be used (e.g.
+the cosine similarity or a similarity derived from a more flexible  Mahalanobis distance), here we choose the Euclidean distance given the prior knowledge on the domain: the cosine similarity is often used as an effective way to normalize bag of words representations that encode frequencies or counts (while here we have a 0-1 encoding), and Mahalanobis distances are useful when unaccounted correlations and collinearities between features are expected (while in our case the ontology terms form a controlled non redundant vocabulary).
 
@@
 Finally, the regularization trade off parameter in the SVM was sampled in $\lbrace 10^{-5},  \ 10^{-4}, \ 10^{-3},\ 10^{-2},\ 10^{-1}, 1,\ 10,\ 10^2 \rbrace$.
-The optimal parameter values of CDNK obtained by validation process are $D = 1$, $C=5$, $r^*=2$, $d^*=1$, $P=15$ and the trade-off parameter of SVM is $10^{-2}$.
+The optimal parameter values of CDNK obtained by validation process are $D = 10$, $C=5$, $r^*=2$, $d^*=1$, $P=15$ and the trade-off parameter of SVM is $10^{-2}$.
 %\subsection{Parameter Impact}
@@
 We have shown how decomposing a network in a set of connected sparse graphs allows us to take advantage of the discriminative power of CDNK, a novel decomposition kernel, to achieve state-of-the-art results for a disease gene prioritization task.
-Moreover, we have also introduced the possibility to integrate in the kernel computation  additional information in the form of real valued vectors in the case in which it is available for graph nodes.
-We experimentally showed that this possibility actually leads to even better performances.
-Future work we will investigate on how to: \textit{i}) decompose networks in a data driven way; \textit{ii}) extend the CDNK approach to gene-disease association problems exploiting multiple heterogeneous information sources in a joint way.
+Moreover, we have also introduced the possibility to integrate in the kernel computation  additional information in the form of real valued vectors when such information is available for each graph node.
+We experimentally showed that this actually leads to improved performances.
+In future work we will investigate how to \textit{i}) decompose networks in a data driven way and \textit{ii}) extend the CDNK approach to gene-disease association problems exploiting multiple heterogeneous information sources.
 
@@
 
-\bibliography{mybibfile}
+%\bibliography{mybibfile}
+\begin{thebibliography}{10}
+\expandafter\ifx\csname url\endcsname\relax
+  \def\url#1{\texttt{#1}}\fi
+\expandafter\ifx\csname urlprefix\endcsname\relax\def\urlprefix{URL }\fi
+\expandafter\ifx\csname href\endcsname\relax
+  \def\href#1#2{#2} \def\path#1{#1}\fi
+
+\bibitem{mordelet2011prodige}
+F.~Mordelet, J.-P. Vert, Prodige: Prioritization of disease genes with
+  multitask machine learning from positive and unlabeled examples, BMC
+  bioinformatics 12~(1) (2011) 389.
+
+\bibitem{chen2014disease}
+B.~Chen, M.~Li, J.~Wang, F.-X. Wu, Disease gene identification by using graph
+  kernels and markov random fields, Science China. Life Sciences 57~(11) (2014)
+  1054.
+
+\bibitem{valentini2014extensive}
+G.~Valentini, A.~Paccanaro, H.~Caniza, A.~E. Romero, M.~Re, An extensive
+  analysis of disease-gene associations using network integration and fast
+  kernel-based gene prioritization methods, Artificial Intelligence in Medicine
+  61~(2) (2014) 63--78.
+
+\bibitem{kondor2002diffusion}
+R.~I. Kondor, J.~Lafferty, Diffusion kernels on graphs and other discrete input
+  spaces, in: ICML, Vol.~2, 2002, pp. 315--322.
+
+\bibitem{fouss2006experimental}
+F.~Fouss, L.~Yen, A.~Pirotte, M.~Saerens, An experimental investigation of
+  graph kernels on a collaborative recommendation task, in: Data Mining, 2006.
+  ICDM'06. Sixth International Conference on, IEEE, 2006, pp. 863--868.
+
+\bibitem{chebotarev2006matrix}
+P.~Chebotarev, E.~Shamis, The matrix-forest theorem and measuring relations in
+  small social groups, arXiv preprint math/0602070.
+
+\bibitem{haussler1999convolution}
+D.~Haussler, Convolution kernels on discrete structures, Tech. rep., Technical
+  report, Department of Computer Science, University of California at Santa
+  Cruz (1999).
+
+\bibitem{gartner2003survey}
+T.~G{\"a}rtner, A survey of kernels for structured data, ACM SIGKDD
+  Explorations Newsletter 5~(1) (2003) 49--58.
+
+\bibitem{borgwardt2005shortest}
+K.~M. Borgwardt, H.-P. Kriegel, Shortest-path kernels on graphs, in: Data
+  Mining, Fifth IEEE International Conference on, IEEE, 2005, pp. 8--pp.
+
+\bibitem{shervashidze2009fast}
+N.~Shervashidze, K.~M. Borgwardt, Fast subtree kernels on graphs, in: Advances
+  in neural information processing systems, 2009, pp. 1660--1668.
+
+\bibitem{shervashidze2011weisfeiler}
+N.~Shervashidze, P.~Schweitzer, E.~J.~v. Leeuwen, K.~Mehlhorn, K.~M. Borgwardt,
+  Weisfeiler-lehman graph kernels, Journal of Machine Learning Research
+  12~(Sep) (2011) 2539--2561.
+
+\bibitem{costa2010fast}
+F.~Costa, K.~De~Grave, Fast neighborhood subgraph pairwise distance kernel, in:
+  Proceedings of the 26th International Conference on Machine Learning,
+  Omnipress, 2010, pp. 255--262.
+
+\bibitem{alvarez2005k}
+J.~I. Alvarez-Hamelin, L.~Dall'Asta, A.~Barrat, A.~Vespignani, k-core
+  decomposition: A tool for the visualization of large scale networks, arXiv
+  preprint cs/0504107.
+
+\bibitem{tarjan1985decomposition}
+R.~E. Tarjan, Decomposition by clique separators, Discrete mathematics 55~(2)
+  (1985) 221--232.
+
+\bibitem{goh2007human}
+K.-I. Goh, M.~E. Cusick, D.~Valle, B.~Childs, M.~Vidal, A.-L. Barab{\'a}si, The
+  human disease network, Proceedings of the National Academy of Sciences
+  104~(21) (2007) 8685--8690.
+
+\bibitem{gene2004gene}
+G.~O. Consortium, et~al., The gene ontology (go) database and informatics
+  resource, Nucleic acids research 32~(suppl 1) (2004) D258--D261.
+
+\end{thebibliography}
+
 \end{document}

### 2017-11-16 neurocomputingv3.tex

@@
 Finally, the regularization trade off parameter in the SVM was sampled in $\lbrace 10^{-5},  \ 10^{-4}, \ 10^{-3},\ 10^{-2},\ 10^{-1}, 1,\ 10,\ 10^2 \rbrace$.
-The optimal parameter values of CDNK obtained by validation process are $D = 10$, $C=5$, $r^*=2$, $d^*=1$, $P=15$ and the trade-off parameter of SVM is $10^{-2}$.
+The optimal parameter values of CDNK obtained by validation process are $D = 20$, $C=5$, $r^*=2$, $d^*=1$, $P=15$ and the trade-off parameter of SVM is $10^{-2}$.
+
 %\subsection{Parameter Impact}
@@
 \subsection{Results and Discussion}
-%==================================================
-%==================================================
-%==================================================
+
 \begin{table}[h]
@@
 \end{table}
-%=====================================
-%=====================================
+
 \begin{table}[h]
@@
 
-%==================================================
 \begin{figure}[h]
